Female-germline specific protein Sakura interacts with Otu and is crucial for germline stem cell renewal and differentiation and oogenesis

Version 1

Abstract

Abstract

During oogenesis, self-renewal and differentiation of germline stem cells (GSCs) must be tightly regulated. The Drosophila female germline serves as an excellent model for studying these regulatory mechanisms. Here, we report that a previously uncharacterized gene CG14545, which we named sakura, is essential for oogenesis and female fertility in Drosophila. Sakura is predominantly expressed in the ovaries, particularly in the germline cells, including GSCs. sakura null mutant female flies display rudimentary ovaries with germline-less and tumorous phenotypes, fail to produce eggs, and are completely sterile. The germline-specific depletion of sakura impairs Dpp/BMP signaling, leading to aberrant bag-of-marbles (bam) expression, resulting in faulty differentiation and loss of GSCs. Additionally, sakura is necessary for normal piwi-interacting RNAs (piRNAs) levels and for proper localization of Oo18 RNA-binding protein (Orb) in developing oocytes. We identified Ovarian Tumor (Otu) as protein binding partner of Sakura, and we found that loss of otu phenocopies loss of sakura in ovaries. Thus, we identified Sakura as a crucial factor for GSC renewal and differentiation and oogenesis, and propose that Sakura and Otu function together in these processes.

Introduction

Oogenesis is a complex process in which germline stem cells (GSCs) develop into mature female gametes or oocytes. This process involves multiple layers of regulations and numerous genes, some of which remain to be discovered and characterized. Drosophila melanogaster, a genetically tractable model organism, has long been at the forefront of oogenesis research.

Adult Drosophila females possess a pair of ovaries, each composed of 12-16 ovarioles. At the anterior tip of each ovariole is a structure called the germarium, which houses two to three GSCs (Fig S1). GSCs reside in the most anterior region of the germarium, where they directly contact cap cells and escort cells, forming a niche crucial for GSC self-renewal 1–3. GSCs undergo asymmetric mitotic division, producing two distinct cells: one GSC and one cystoblast1,4. The cystoblast, destined to differentiate, undergoes four mitotic divisions with incomplete cytokinesis, resulting in interconnected cysts of 2, 4, 8, and 16 cells 5,6. Of the 16 cells, one will differentiate into an oocyte, while the remaining 15 develop into polyploid nurse cells, which synthesize proteins and RNAs for deposition into the oocyte7.

GSC proliferation is essential for proper oogenesis. GSCs must be tightly regulated to maintain their undifferentiated state while continuing to divide and differentiate. Failure in GSC self-renewal or inappropriate differentiation leads to stem cell loss, disrupting oocyte production8,9. Conversely, uncontrolled GSC division or defective differentiation results in an overabundance of GSC-like cells (a tumorous phenotype) and fewer differentiated cells, ultimately inhibiting oocyte production10,11. Thus, a precise balance between GSC self-renewal and cyst differentiation within the germarium is essential for proper oogenesis and is tightly regulated.

Bone Morphogenetic Protein (BMP) signaling plays a critical role in regulating GSC self-renewal and differentiation during Drosophila oogenesis. In the stem cell niche, GSCs are physically anchored to cap cells via adherens and gap junctions12,13. Cap cells secrete BMP ligands, such as Decapentaplegic (Dpp) and Glass bottom boat (Gbb), which are recognized by transducing receptors like Thickvein (Tkv) and Saxophone (Sax) on the GSCs14,15 (Fig S1). Upon activation, these receptors phosphorylate a transcription factor Mad. The phosphorylated Mad (pMad) translocates into the nucleus, where it represses the transcription of bag-of-marbles (bam), a key differentiation factor 16,17.

Bam, a ubiquitin-associated protein, is essential for cystoblast differentiation. Dpp/BMP signaling in the niche represses bam expression, preventing GSCs from differentiating and enabling self-renew. However, as the daughter cells (cytoblasts) exit the niche, they derepress bam and initiate the differentiation program (Fig S1)7. Tight regulation of Dpp/BMP signaling and bam expression ensures that only cells within the niche adopt GSC fate, while adjacent daughter cells differentiate into cystoblasts. Loss of bam leads to blocked differentiation and the accumulation of GSC-like cells, whereas ectopic bam expression forces GSCs differentiation, depleting the stem cell pool 18–20.

One of the proteins that interacts with Bam is Ovarian Tumor (Otu). Otu forms a deubiquitinase complex with Bam, which deubiquitinates Cyclin A (CycA), stabilizing CycA and promoting GSC differentiation21. Otu is crucial for oogenesis and female fertility22,23. Mutations in the otu gene cause various ovarian defects, including tumorous growths, germ cell loss, and abnormalities in oocyte determination and nurse cell dumping, indicating Otu’s importance in multiple stages of oogenesis22–25. However, the precise molecular mechanisms by which Otu functions in these processes remain largely unknown.

In this study, we aimed to identify a novel gene required for oogenesis. Through a search for uncharacterized genes expressed exclusively in Drosophila ovaries, we discovered CG14545, a gene encoding a 114 amino acids (aa) long protein without any known functional domain (Fig 1A). CG14545 is expressed specifically in female germline cells, including GSCs. Our genetic analysis revealed that CG14545 is required for oogenesis and female fertility and plays an important intrinsic role in regulating GSCs renewal and differentiation. Additionally, CG14545 is important for the piRNA pathway. We identified Otu as a binding partner of the CG14545 protein, suggesting that both proteins may participate in the same molecular pathways to regulate GSC fate and differentiation. We named the CG14545 gene sakura, meaning “cherry blossom” in Japanese, symbolizing birth and renewal.

Sakura expression pattern and mutant allele

(A) Drosophila Sakura protein (Sakura/CG14545) and its null mutant allele generated in this study. (B) sakura mRNA expression pattern. Data obtained from http://flybase.org/reports/FBgn0040602.htm. (C) Western blot of dissected fly tissues. (D) Western blot of ovary lysates. (E) Confocal images of the germarium from the sakura-EGFP transgenic fly. Sakura-EGFP (green), Hts (red), and DAPI (blue). Scale bar: 10 µm. (F) Confocal images of the egg chambers from the sakura-EGFP transgenic fly. Sakura-EGFP (green), Vasa (red), and DAPI (blue). Sakura-EGFP is expressed in nurse cells and enriched in the developing oocyte (yellow arrow). Scale bar: 20 µm.

Results

Sakura is exclusively expressed in the ovaries

Based on high-throughput expression data in Flybase, sakura mRNA is highly and exclusively expressed in adult ovaries (Fig 1B). To investigate Sakura protein expression, we generated a polyclonal anti-Sakura antibody against a recombinant full-length Sakura protein. Using this antibody, we examine Sakura protein expression at different stages of Drosophila development and across tissues. We found that Sakura protein is specifically expressed in the ovary, particularly in stage 1-11 oocytes (Fig 1C).

sakura mutant flies

To explore the biological and molecular functions of Sakura in vivo, we generated a sakura mutant allele (sakuranull) with a one nucleotide (nt) insertion replacing 33 nts within the Sakura coding region using the CRISPR/Cas9 genome editing system (Fig 1A)26. This mutation causes a frameshift and a premature stop codon, resulting in a 33-amino acid (aa) N-terminal fragment of Sakura followed by a 22-aa segment produced by frameshifted translation (Fig 1A). This truncated fragment is unlikely to be functional, and we were unable to detect stable protein expression, as shown below. Thus, we consider this allele a null mutation. sakuranull/null flies were viable, revealing that Sakura is not essential for survival.

To validate the sakuranull strain, we performed Western blots using ovary lysates from sakura mutant flies using the anti-Sakura antibody. As expected, full-length Sakura protein was detected in the ovaries of wild-type (sakura+/+) and heterozygous controls (sakuranull/+), but not in those of homozygous mutants (sakuranull/null) (Fig 1D). No smaller protein corresponding to the truncated Sakura fragment was detected in either sakuranull/+ or sakuranull/null ovaries thus, suggesting that the fragment is unstable or not expressed.

Sakura is cytoplasmic and expressed specifically in germ cells

To determine the localization of Sakura protein within the ovaries, we generated transgenic flies carrying a sakura coding sequence fused with EGFP at the C-terminus (sakura-EGFP) under the control of the sakura promoter. Western blot analysis of ovary lysates showed that Sakura-EGFP is expressed at levels comparable to endogenous Sakura (Fig 1D). Confocal imaging revealed that Sakura-EGFP is specifically expressed in the cytoplasm of germ cells, including GSCs, cyst cells, nurse cells, and developing oocytes (Fig 1E and 1F). Sakura-EGFP was not expressed in somatic cells such as terminal filament, cap cells, escort cells, or follicle cells. (Fig 1E). In the egg chamber, Sakura-EGFP was detected in the cytoplasm of nurse cells and was enriched in developing oocytes (Fig 1F).

sakura is essential for female fertility

Given its ovary-specific expression at both mRNA and protein levels (Fig 1B and 1C), we hypothesized that Sakura plays a critical role in ovarian function. Fertility assays revealed that sakuranull/null females laid no eggs when mated with wild-type males, while control female flies (sakura+/+ and sakuranull/+) laid numerous eggs (Fig 2A). Dissection of sakuranull/nullovaries revealed that they were rudimentary compared to normal ovaries in controls (Fig 2C). This suggests that the inability of sakuranull/null female to lay eggs is due to their underdeveloped ovaries. In contrast, sakuranull/null males were fertile and showed no significant differences compared to controls (sakura+/+ and sakuranull/+) (Fig S2).

sakura mutant flies are female-sterile and have rudimentary ovaries.

(A-B) Female fertility assays. (A) The number of eggs laid by test females crossed with OregonR wild-type males and (B) hatching rates of the eggs. Mean ± SD (n = 3). P-value < 0.05 (Student’s t-test, unpaired, two-tailed) is indicated by *. (C) Stereomicroscope images of dissected whole ovaries. Scale bar: 1 mm.

To confirm that the observed female sterility was caused by the loss of sakura, we generated sakura rescue flies expressing Sakura-EGFP in the sakuranull/null background (sakuranull/null; sakura-EGFP). Western blotting confirmed that these flies expressed Sakura-EGFP, but not endogenous Sakura (Fig 1D). The sakura-EGFP transgene fully rescued both female fertility and ovary morphology in the sakuranull/null background (Fig 2). sakura-EGFP rescue females showed no significant fertility differences compared to controls (Fig 2A and 2B), and their ovaries were normal in shape and size (Fig 2C). We concluded that Sakura is essential for female fertility and normal ovary morphology but dispensable for male fertility, consistent with the observation that Sakura is predominantly expressed in ovarian germline cells (Fig 1).

sakuranull/null ovaries are germless and tumorous

Given the rudimentary appearance of sakuranull/null ovaries, we questioned whether they contained germ cells. We used flies carrying vasa-EGFP knocked-in reporter in the sakuranull/null background, as Vasa is a known germ cell marker7. We observed that some sakuranull/null ovarioles were devoid of germ cells ("germless"), while others retained germ cells (Fig 3A), suggesting that loss of sakura leads to germ cell depletion. Additionally, sakuranull/null ovaries exhibited significantly higher levels of cleaved Caspase-3 staining, indicative of apoptosis, compared to controls (sakuranull/+) (Fig 3B).

sakuranull ovaries are germless and tumorous

(A) Confocal images of the ovaries from control (sakuranull/+) and sakuranull/null harboring vasa-EGFP. Vasa-EGFP (green) and DAPI (blue). Yellow stars label the anterior tip of the ovaries. Scale bar: 100 µm. (B) Confocal images of sakuranull/null ovaries stained with anti-cleaved Caspase-3 antibody. Cleaved caspase-3 (green) and DAPI (blue). Control ovary is sakuranull/+. Yellow stars label the anterior tip of the ovaries. Scale bars: 20 µm for control and 100 µm for sakuranull/null. (C) Confocal images of control (sakuranull/+) and sakuranull/null ovaries stained with anti-Hts antibody to label spectrosomes and fusomes. Hts (red) and DAPI (blue). Yellow star labels the anterior tip in the control ovary. Red and green stars label the anterior tip of sakuranull/null tumorous and germless ovarioles, respectively. Scale bars: 20 µm for control and 50 µm for sakuranull/null. (D) Ratio (%) of normal, germless, and tumorous ovaries of indicated genotypes (n=33, 25, 74, 29 respectively). (E) Quantification of GSC-like cell number in germaria of indicated genotypes (n = 33, 30, 35, and 31 respectively).

Oogenesis begins in the germarium, which contains 2-3 GSCs identified by round, unbranched spectrosomes that contact cap cells7 (Fig S1). In contrast, cysts have branched fusomes. Immunostaining with hu-li tai shao (HTS) antibody, which marks spectrosomes and fusomes, revealed an excess number of GSC-like cells with round spectrosomes in sakuranull/null ovarioles containing germ cells (Fig 3C, red stars), indicative of a "tumorous" phenotype as previously described27–29. Additionally, we observed an excess number of cyst cells with branched fusomes that persisted throughout the ovarioles (Fig 3C), suggesting abnormal cyst cell differentiation and division.

Within the same ovary, some sakuranull/null ovarioles exhibited the tumorous phenotype, while others were germless, without any spectrosome or fusome staining (Fig 3C, green stars). Quantification revealed that 35% of sakuranull/null ovarioles were tumorous while the remaining 65% were germless (n=74) (Fig 3D). In contrast, no tumorous or germless phenotypes were observed in control (sakura+/+ and sakuranull/+) or sakura-EGFP rescue flies (Fig 3D). Furthermore, ∼95% of sakuranull/null ovarioles with germ cells contained excess GSC-like cells (>5) (Fig 3E and Fig S3). The mean number of GSCs or GSC-like cells for sakura+/+, sakuranull/+, and sakura-EGFP rescue was 2.2 ± 0.6, 2.2 ± 0.6, and 2.1 ± 0.7, respectively while that for sakuranull/null was 26.5 ± 17.5 (p-value <0.05). These results suggest that Sakura is essential for regulating GSC division and differentiation. Overall, we conclude that Sakura is required for germ cell survival, division, and differentiation.

Loss of sakura results in reduced germline piRNA

Piwi-interacting RNAs (piRNAs) are produced in germline cells and somatic follicle cells and transcriptionally and post-transcriptionally silence transposon expression through sequence-complementarity. Disruption of the piRNA pathway can result in oogenesis arrest, germ cell loss, rudimentary ovaries, and sterility9,30. Loss of piRNAs in germ cells leads to the upregulation (desilencing) of transposons, increased DNA damage, and apoptotic cell death31. We investigated whether the loss of sakura would result in loss of piRNA and transposon upregulation. We performed high-throughput sequencing (small RNA-seq and polyA+ RNA-seq) on ovary RNA samples from sakura+/+, sakuranull/+, sakuranull/null, and sakura-EGFP rescue (sakuranull/null with sakura-EGFP) flies.

piRNA levels in sakuranull/null ovaries were reduced compared to controls (sakura+/+), while sakuranull/+ and sakura-EGFP rescue showed no such reduction (Fig 4A). Consistent with this, transposon RNA levels in sakuranull/null ovaries were upregulated compared to sakura+/+, while sakuranull/+ and sakura-EGFP rescue showed no such upregulation (Fig 4B). Given that Sakura is specifically expressed in germline cells, but not in somatic cells (Fig 1E and 1F), we hypothesized that loss of sakura interrupts the piRNA pathway in germ cells, leading to the desilencing of germline transposons. A well-characterized germline transposon in Drosophila is Burdock30,32. We found that Burdock piRNA levels were significantly lower in sakuranull/nullcompared to controls and rescue flies (Fig 4C).

Loss of sakura results in lower piRNA levels and loss of piRNA-mediated transposon silencing

(A) Scatter plots of normalized number of transposon piRNA reads from small RNA-seq of indicated genotypes compared with sakura+/+. Means of three biological replicates are plotted. Downregulated (fold-change < 0.5) and upregulated (fold-change > 2) transposon piRNAs are shown in cyan and magenta, respectively. (B) Volcano plots of transposon RNAs from RNA-seq of indicated genotypes compared with sakura+/+. Three biological replicates per genotype were analyzed. Downregulated (adjusted p-value < 0.001 and log2(fold-change) < -2) and upregulated (adjusted p-value < 0.001 and log2(fold-change) > 2) transposons are shown in cyan and magenta, respectively. (C) Normalized number of reads for Burdock piRNAs from small RNA-seq. Mean ± SD (n = 3.) P-value < 0.05 (Student’s t-test, unpaired, two-tailed) is indicated by *. (D) The Burdock sensor harbors a nanos promoter, a nuclear localization signal (NLS) appended to GFP and β-gal coding sequences, and a target sequence for Burdock piRNAs in the 3’UTR. Confocal images of ovaries from control (yRNAi) and sakuraRNAi flies harboring the Burdock sensor, where RNAi knockdown was specifically driven in the female germline with UAS-Dcr2 and NGT-Gal4. GFP (green), β-gal (red), and DAPI (blue). Scale bar: 50 µm.

We employed the Burdock sensor, a transposon reporter tool to monitor germline piRNA activity30. The reporter expresses nuclear GFP and β-gal under the control of the nanos promoter for germline expression, with a target sequence for Burdock piRNAs in the 3′ UTR30 (Fig 4D). Using a sakura RNAi line driven by UAS-Dcr2 and NGT-Gal4 to knock down sakura specifically in the germline, we observed that GFP and β-gal expression were highly elevated in sakura knockdown germlines compared to control RNAi (yRNAi) knockdowns (Fig 4D). This confirmed that loss of sakura leads to a loss of piRNA-mediated transposon silencing in the germline. We conclude that Sakura is essential for proper piRNA levels and piRNA-based transposon silencing in the germline.

sakura is important for oogenesis in germline cells beyond GSCs and germline cysts

Both sakuranull/null and sakura germline RNAi knockdown driven by UAS-Dcr-2 and NGT-Gal4, which initiates RNAi in germline from GSCs, resulted in rudimentary ovaries (Fig 2C and 5A). These ovaries lacked later-stage germline cells and egg chambers, making it difficult to assess the role of Sakura beyond the germarium. However, because Sakura-EGFP expression is not limited to GSCs and germline cysts (Fig 1E and 1F), we speculated that Sakura might function in later-stage germline cells as well.

sakura is important for oogenesis in later-stage egg chambers

(A) Stereomicroscope images of dissected whole ovaries. Scale bar: 500 µm. (B) Western blot of dissected ovary lysates. sakuraRNAi #1 and sakuraRNAi #2 are two independent RNAi lines. (C) Number of eggs laid by sakura RNAi knockdown driven by UAS-Dcr2 and TOsk-Gal4. Mean ± SD (n = 5). P-value < 0.05 (Student’s t-test, unpaired, two-tailed) is indicated by *. (D) Number of stage 14 oocytes produced by sakura RNAi knockdown flies driven by UAS-Dcr2 and TOsk-Gal4. P-value < 0.05 (Student’s t-test, unpaired, two-tailed) is indicated by *. (E) Confocal images of the ovaries from sakura RNAi knockdown flies driven by UAS-Dcr2 and TOsk-Gal4, stained with anti-Orb and anti-Vasa antibodies. Orb (green), Vasa (red), and DAPI (blue). Yellow arrows label the localization and enrichment of Orb in the developing oocytes. Scale bar: 100 µm. In A-E, y-RNAi was used as a control.

To explore this possibility, we used TOsk-Gal4 (a combination of osk-Gal4 and αTub67C-Gal4) to drive sakura RNAi knockdown in germline cells after the germline cyst stage (from germarium region 2b onward. Fig S1A), sparing GSCs and germline cysts33. Unlike NGT-Gal4, TOsk-Gal4-driven sakura RNAi knockdown did not drastically affect ovary morphology (Fig 5A), allowing us to study the effects of sakura loss in egg chambers. We confirmed sakura RNAi (sakura RNAi #1 and #2) knockdown efficiency by Western blot, showing effective depletion of Sakura in ovaries from both NGT-Gal4 and TOsk-Gal4-driven RNAi lines (Fig 5B).

Interestingly, TOsk-Gal4-driven sakura RNAi knockdown severely reduced the numbers of eggs laid (Fig 5C) and stage 14 oocytes in ovaries compared to control RNAi (yRNAi) (Fig 5D), suggesting that sakura is important for oogenesis beyond the germline cyst stage. Additionally, we observed mislocalized and dispersed Oo18 RNA-binding protein (Orb. A marker for oocyte identity) in TOsk-Gal4-driven sakura RNAi knockdowns (Fig 5E). Proper Orb localization is critical for oocyte specification and maintenance34, and its mislocalization likely underlies the reduced ability to produce stage 14 oocytes and eggs sakura knockdown flies (Fig 5C and 5D). We conclude that Sakura is crucial for proper Orb localization and in germline cells beyond GSCs and germline cysts. In subsequent sakura RNAi experiments, we used the sakura RNAi #2 line.

Sakura is required intrinsically for GSC establishment, maintenance, and differentiation

To explore whether sakura functions autonomously in the germline, we conducted mosaic analysis using the FLP-FRT system35. First, we assessed its role in GSC establishment by marking sakuranull clones in primordial germ cells (PGCs)) before the early third instar larval stage and tracking their development into adult GSCs, as described in a previous study36. In control adult flies,12.3% (20/163) of GSCs were marked, while only 1.9% (4/213) were marked in sakuranullmutants, demonstrating that sakura is autonomously important for GSC establishment.

Next, we assessed GSC maintenance by generating marked sakuranullGSCs in adult flies using the FLP-FRT system and measuring their loss rate over time 14,36. The percentage of marked control GSC clones and marked sakuranull GSC clones 4 days after clone induction was 27.8% and 22.1%, respectively (Fig 6A). These numbers were considered the initial percentage. The percentage of marked control GSCs decreased to 25.8% and 16.8% on days 7 and 14, respectively (Fig 6A), resulting in a loss rate of 39.4% over the 10-day testing period (from day 4 to day 14). In contrast, the percentage of marked sakuranull GSC clones decreased to 16.3% and 5.4% on days 7 and 14, respectively. Thus, the loss rate of sakuranull GSCs is 75.6% over the 10-day period, which is higher than controls. We conclude that sakura is intrinsically required for GSC maintenance.

Germline clonal analysis of sakuranull

(A) Percentage of germaria with marked GSC clones indicated by the absence of GFP at 4, 7, and 14 days after clone induction at the adult stage. Arrows indicate the percent decrease of marked GSC clones compared to day 4. The genotype, actual percentage, and total number of germaria examined are shown in the adjacent table. (B) Confocal images of germ cell clones. sakuranull and control clones were marked by the absence of GFP. GFP (green), Hts (red), and DAPI (blue). Scale bar: 20 µm. (C) Number of marked (GFP-negative) GSC-like cells in germaria with marked GSCs of the indicated genotypes at 4, 7, and 14 days after clone induction. GSC-like cells containing round spectrosome were identified through immunostaining with anti-Hts antibody. P-value < 0.05 (Student’s t-test, unpaired, two-tailed) is indicated by *.

Germline depletion of sakura caused germless and tumorous phenotypes (Fig 3). To determine whether sakura is intrinsically required in the germ line cells for GSC differentiation, we quantified the number of marked (GFP-negative) and unmarked (GFP-positive) GSC-like cells (germline cells with a round spectrosome) in the germaria carrying marked GSC clones at 4, 7, and 14 days after clone induction. We found that the number of marked sakuranullGSC-like cells was higher than that of marked control GSC-like cells at all time points (Fig 6C). The number of marked sakuranullGSC-like cells increased over the 10-day period, while the number of marked control cells did not (Fig 6C). In contrast, the number of unmarked GSC-like cells in germaria containing sakuranull GSC clones did not differ significantly from those in germaria containing control GSC clones, and the number did not increase (Fig S4). These results indicate that sakura is intrinsically important in germline cells, including GSCs, to regulate their proper division and differentiation. Taken together, we conclude that sakura is required intrinsically for GSC establishment, maintenance, and differentiation.

Loss of sakura inhibits Dpp/BMP signaling

Dpp/BMP signaling is a key pathway regulating GSC self-renewal and differentiation (Fig S1)7,37. We hypothesized that the germless and tumorous phenotypes observed in sakura loss-of-function ovaries could be due to Dpp/BMP signaling misregulation. To test this, we knocked down sakura in the germline driven by UAS-Dcr2 and NGT-Gal4 in the presence of bam-GFP reporter38. In controls (yRNAi), Bam-GFP expression was restricted to 8-cell cysts and disappeared in 16-cell cysts and onward (Fig 7A)38. However, in sakura knockdowns, Bam-GFP expression persisted throughout the ovariole, with excess Bam-GFP positive cells (Fig 7A). This suggests that Bam expression is not repressed by Dpp/BMP signaling, leading to GSC loss in sakura knockdowns.

Loss of sakura inhibits Dpp/BMP signaling

(A) Confocal images of ovaries from control (yRNAi) and sakuraRNAi flies harboring the bam-GFP reporter, where RNAi knockdown was specifically driven in the female germline with UAS-Dcr2 and NGT-Gal4. Bam-GFP (green), Hts (red), and DAPI (blue). Scale bar: 20 µm. (B, C) Confocal images of germaria with germline clones of sakuranull stained with (B) anti-Bam antibody and (C) anti-pMad antibody. GFP (green), Bam or pMad (red), and DAPI (blue). Scale bar: 10 µm. (D) Mean pMad intensity in the germline clones of the indicated genotypes. Mean ± SD (n = 7). P-value < 0.05 (Student’s t-test, unpaired, two-tailed) is indicated by * (E) Confocal images of germaria with germline clones of sakuranull stained with anti-CycA antibody. GFP (green), CycA (red), and DAPI (blue). Scale bar: 10 µm. (F) Mean CycA intensity in the germline clones of the indicated genotypes. Mean ± SD (n = 5 and 10 for FRT82B and FRT82B, sakuranull respectively). P-value < 0.05 (Student’s t-test, unpaired, two-tailed) is indicated by *. In B, C, and E, clones were marked with the absence of GFP.

To further confirm this finding, we used the FLP-FRT technique to induce sakuranull clones in the germaria and stained them with anti-Bam antibody20. All sakuranull clones expressed Bam throughout the germarium, whereas in control clones, Bam was not expressed in the GSCs and was restricted to cyst cells (Fig 7B). In GSCs, pMad translocates into the nucleus to repress Bam expression 16,17(Fig S1). We stained ovaries with anti-pMad antibodies and found that pMad intensity was significantly reduced in sakuranull GSC clones compared to control GSCs (Fig 7C and D). This indicates that the dysregulated Bam expression observed in sakuranullwas due to reduced pMad levels in GSCs.

Previous studies have shown that ectopic expression of a stable form of CyA leads to germ cell loss39. This germ cell loss phenotype is also observed upon ectopic bam expression in GSCs14,40,41. It was reported that Bam associates with Ovarian tumor (Otu) to promote deubiquitination and stabilization of CycA21. Since we observed derepressed bam expression in sakuranull cells, we investigated whether CycA levels were affected. Staining ovaries with anti-CycA antibodies revealed higher CycA intensity in sakuranull clones compared to neighboring wild-type cells in the same germarium and control clones in control germarium (Fig 7E). sakuranull clones exhibited significantly higher mean CycA intensity compared to control clones (Fig 7F). The elevated CycA levels in sakuranullclones likely results from Bam misexpression and Bam-mediated stabilization.

Finally, we tested whether we could rescue the germless phenotype caused by sakura loss-of-function by knocking down either bam or cycA in the germline. Double RNAi knockdown of sakura and bam in the germline using UAS-Dcr-2 and NGT-Gal4 partially rescued the germless phenotype (Fig 8). In contrast, knocking down sakura and cycA together failed to rescue the germless phenotype (Fig 8). This suggests that the germless phenotype of sakura is partly due to bam misexpression, not cycA.

The ratio of ovary germless phenotype from double RNAi knockdown of sakura with bam, cycA, or otu

Mean ratio (%) of ovaries with the germless phenotype observed in double RNAi knockdown of sakura with bam, cycA, or otu. UAS-Dcr2 and NGT-Gal4 was used to drive RNAi knockdown in the germline. n is the total number of ovaries examined for the indicated genotypes. Three independent crosses (biological replicates) were set up to generate the flies with genotypes of interest, and the ratio of ovaries with the germless phenotype in each replicate was calculated. Mean ± SD (three biological replicates). P-value < 0.05 (Student’s t-test, unpaired, two-tailed) is indicated by *.

Sakura binds Otu

To explore the molecular function of Sakura in oogenesis, we aimed to identify Sakura-interacting proteins. We conducted a co-immunoprecipitation experiment using anti-GFP magnetic beads on the ovary lysates of flies expressing Sakura-EGFP, followed by mass spectrometry to identify co-immunoprecipitated proteins. Ovary lysates from flies without the sakura-EGFP transgene (w1118) served as negative controls. Mass spectrometric analysis revealed several proteins with peptide signals present in all three biological replicates of the Sakura-EGFP samples without any signals in any of the three biological replicates of the negative control (Table 1). Among these, the most promising candidate, exhibiting the highest unique peptide signal in each of three replicates of the Sakura-EGFP samples was Otu.

To confirm this finding, we generated a polyclonal anti-Otu antibody. Using Western blot analysis, we detected endogenous Otu protein in the Sakura-EGFP immunoprecipitants from ovaries, but not in the negative control (Fig 9A), validating the mass spectrometry results. Next, using wild-type (without Sakura-EGFP) ovary lysates, we immunoprecipitated endogenous Sakura with anti-Sakura antibody and detected endogenous Otu in the immunoprecipitant by Western blot (Fig 9B), demonstrating that endogenous Sakura interacts with endogenous Otu. Furthermore, we found that ectopically expressed Flag-tagged Otu co-immunoprecipitated with ectopically expressed HA-tagged Sakura in S2 cells (Fig 9C).

Sakura interacts with Otu.

(A) Co-immunoprecipitation using anti-GFP magnetic beads followed by Western blotting. Ovary lysates expressing Sakura-EGFP and those from w1118negative control were tested. (B) Co-immunoprecipitation using beads bound with rabbit anti-Sakura followed by Western blotting. Ovary lysates from w1118 flies were used. Rabbit IgG was used as control IP. (C) Co-immunoprecipitation using beads bound with anti-FLAG antibody followed by Western blotting. S2 cell lysates expressing 3xFLAG-Otu and Sakura-mCherry-3xHA or mCherry-3xHA (negative control) were used.

The otu gene encodes two annotated protein isoforms, a predominant 98 kDa isoform and a less abundant 104 kDa isoform resulting from alternative splicing42. The 98kDa isoform lacks Tudor domain encoded in an alternatively spliced 126-bp exon43. It has been reported that the 104 kDa Otu isoform can perform all functions, while the 98 kDa Otu isoform functions in later stages of oogenesis, such as nurse cell regression and oocyte maturation44. The major Otu isoform observed in the inputs and immunoprecipitates from ovaries in Figure 9 corresponds to the 98 kDa isoform.

Next, we sought to identify which regions of Sakura and Otu are important for their interaction. We predicted the structure of the Sakura-Otu protein complex using AlphaFold45. AlphaFold suggests that the N-terminal region of Sakura (1-49aa) and the N-terminal region of Otu (1-405aa) are highly likely to form interactions, while the C-terminal region of Sakura (100-144aa) does not interact with Otu (Fig 10A-C). The Tudor domain of Otu is not directly involved in this interaction. To experimentally validate the predictions made by AlphaFold, we generated a series of truncated Otu fragments, including Otu(ΔTudor) (1-336, 389-853aa, corresponding to the endogenous 98 kDa isoform), Otu(N1) (1-405aa), Otu(N2) (1-339aa), and Otu(C) (405-853aa) (Fig 10B). We then performed co-immunoprecipitation assays to test which fragments could interact with full-length Sakura. All fragments except Otu(C) were associated with full-length Sakura protein in S2 cells (Fig 10D), demonstrating that the C-terminal low-complexity region of Otu is dispensable for the interaction. This assay also showed that the Tudor domain is not critical and the N-terminal fragment (1-339aa) of Otu is sufficient for interaction with Sakura. These findings align well with the predictions made by AlphaFold.

N-terminal regions of Sakura and Otu are important for interaction.

(A) Predicted structure of the Sakura and Otu protein complex made by AlphaFold. Full-length Sakura and N-terminal Otu fragment (N1. 1-405aa) were used for prediction. (B) Full-length Otu and Otu fragments tested in co-immunoprecipitation assays. (C) Full-length Sakura and Sakura fragments tested in co-immunoprecipitation assays (N: N-terminal, M: middle, C: C-terminal). (D, E, F) Co-immunoprecipitation using anti-HA magnetic beads followed by Western blotting. S2 cell lysates expressing HA-tagged mCherry were used as negative controls.

Subsequently, we created several truncated Sakura fragments and conducted co-immunoprecipitation assays in S2 cells to determine which Sakura fragments can interact with full-length Otu. The fragments generated included Sakura(NM) (1-100aa), Sakura(NC) (1-49, 100-144aa), Sakura(MC) (49-144aa), Sakura(N) (1-49aa), Sakura(M) (49-100aa), and Sakura(C) (100-144aa) (Fig 10C). We found that Sakura(NM), Sakura(NC), Sakura(N), and Sakura(M) were associated with Otu (Fig 10E). Compared to Sakura(NM) and Sakura(N), the interaction of Sakura(M) with Otu is relatively weaker, as indicated by a fainter band in the Western blot (Fig 10E). Sakura-NC exhibited a weak interaction with Otu, while Sakura(MC) and Sakura(C) showed no interaction (Fig 10E). These assays demonstrated that the N-terminal region (1-49aa) of Sakura is sufficient for interaction with Otu. Moreover, the results suggest that the C-terminal region of Sakura (100-144aa) does not interact with Otu and that adjoining it with the N-terminal (1-49aa) or the middle (M) region (49-100aa) weakens the interaction with Otu. These findings support the AlphaFold predictions.

Finally, to test whether the interaction between Sakura and Otu can be achieved solely through their N-terminal regions, we performed co-immunoprecipitation assay with Sakura(N) and Otu(N2). We discovered that they interact each other (Fig 10F). A reciprocal co-immunoprecipitation assay also confirmed their interaction (Fig S5), revealing that the interaction between Sakura and Otu can be established with just the 1-49aa of Sakura and the 1-339aa of Otu. Therefore, we conclude that Sakura interacts with Otu in vivo, and the N-terminal regions of both proteins are sufficient for their interaction.

Loss of otu phenocopies loss of sakura

Having established the interaction between Sakura and Otu, we were interested in whether they function together during oogenesis. We speculated that for them to interact and function together, Sakura and Otu must exhibit similar expression and localization patterns in ovaries. We generated two transgenic flies carrying full-length otu-EGFP or otu(ΔTudor)-EGFP transgenes; both constructs are C-terminally fused with EGFP and under the control of the otu promoter. We found that both Otu-EGFP and Otu(ΔTudor)-EGFP display similar expression and localization patterns to Sakura-EGFP (Fig 1E, 1F and Fig 11A). Like Sakura-EGFP, Otu-EGFP and Otu(ΔTudor)-EGFP are specifically expressed in germ cells, but not in follicle cells, and are localized to the cytoplasm, being enriched in the developing oocytes (Fig 11A). We did not observe any difference in localization patterns between Otu-EGFP and Otu(ΔTudor)-EGFP; suggesting that the Tudor domain is not crucial for Otu protein localization. Notably, the localization pattern exhibited by the transgenic Otu-EGFP generated in this study is consistent with a previous report 23.

Loss of otu phenocopies loss of sakura

(A) Confocal images of ovaries from otu-EGFP and otu(ΔTudor)-EGFP transgenic flies. Otu-EGFP and Otu(ΔTudor)-EGFP (green), DAPI (blue). Yellow arrows show the enrichment of Otu-EGFP and Otu(ΔTudor)-EGFP in developing oocytes. Scale bar: 20 µm. (B) Stereomicroscope images of dissected whole ovaries from wRNAi (control) and otuRNAi flies where RNAi knockdown was driven in the female germline with nos-Gal4-VP16 or TOsk-Gal4. Scale bar: 500 µm. (C) Number of eggs laid by TOsk-Gal4 > wRNAi and TOsk-Gal4 > otuRNAi flies. Mean ± SD (n = 5). P-value < 0.05 (Student’s t-test, unpaired, two-tailed) are indicated by *. (D) Violin plots of the number of stage 14 oocytes produced in TOsk-Gal4 > wRNAi and TOsk-Gal4 > otuRNAi flies. n = 30. P-value < 0.05 (Student’s t-test, unpaired, two-tailed) is indicated by *. (E) Confocal images of the ovaries from wRNAi (control) and otuRNAiflies harboring the Burdock sensor where RNAi knockdown was driven in the female germline with UAS-Dcr2 and NGT-Gal4. GFP (green), β-gal (red), and DAPI (blue). Scale bar: 50 µm. (F) Confocal images of the ovaries from nos-Gal4-VP16 > wRNAi and nos-Gal4-VP16 > otuRNAi flies stained with anti-cleaved Caspase 3 antibody. Cleaved Caspase-3 (green) and DAPI (blue). Scale bar: 20 µm

Previous studies have shown that mutations in otu lead to severe defects in germ cell division and differentiation, resulting in a tumorous phenotype22–24,42. To study the role of otu specifically in the germline, we performed RNAi knockdown of otu in the germline driven by nos-Gal4-VP16. Interestingly, we found that germline depletion of otu results in rudimentary ovaries, similar to the loss of sakura (Fig 2D, Fig 5A, and Fig 11B). Furthermore, similar to sakura RNAi (Fig 5), TOsk-Gal4-driven otu RNAi knockdown did not affect ovary morphology (Fig 11B), but it did result in a reduced number of eggs laid and stage 14 oocytes compared with control RNAi (wRNAi) (Figure 11C and 11D).

When sakura and otu were independently RNAi-knockdowned in the germline driven by UAS-Dcr2 and NGT-Gal4, the percentages of germless ovaries for otu-RNAi and sakura-RNAi were similar (∼70%) (Fig 8). However, when otu and sakura were knocked down together, the percentage of germless ovaries increased significantly to 79% (Fig 8), suggesting that the double knockdown of otu and sakura enhances the germless phenotype in the ovaries.

Next, we used the Burdock sensor to test if loss of otu leads to loss of piRNA-mediated transposon silencing in the germline. Similar to sakura (Fig 4D), germline depletion of otu resulted in elevated GFP and β-Gal produced by the Burdock sensor (Fig 11E), indicating loss of piRNA-mediated transposon silencing. To determine if germ cell loss caused by otu-RNAi is a result of apoptosis, we stained the rudimentary ovaries of otu-RNAi for cleaved Caspase-3. Similar to sakuranull/null (Fig 3B), cleaved Caspase-3 signals were higher in otu-RNAi ovaries than in controls, suggesting that the observed germ cell loss is due to apoptosis (Fig 11F).

Loss of sakura leads to misexpression of bam, likely due to low levels of pMad in the germ cells (Fig 7A-C). We were curious whether a similar occurrence could be observed in otu-RNAi. In the otu germline RNAi knockdown ovaries, we found that Bam-GFP expression is not restricted to the germarium but persists throughout the ovariole, resulting in an excess number of Bam-GFP-positive cells (Fig 12A). We also observed lower pMad levels in otu14/14 mutant GSCs (Fig 12B). These data suggest that, similar to sakura, loss of otu results in bam derepression, likely due to low levels of pMad in the germ cells.

Loss of otu results in low levels of pMad and derepression of bam in the germaria

(A) Confocal images of the ovaries from control (wRNAi) and otuRNAi flies harboring the bam-GFP reporter, where RNAi knockdown was driven in the female germline with nos-Gal4-VP16. Bam-GFP (green), Hts (red), and DAPI (blue). Scale bar: 20 µm. (B) Confocal images of germaria from control (otu14/+) and otu14/14 mutant flies stained with anti-pMad (green) and anti-Hts (red) antibodies. Scale bar: 10 µm.

A previous study has shown that Otu possesses deubiquitinase activity21. We asked whether Sakura regulates the Otu’s deubiquitinase activity. To test this, we performed an in vitro deubiquitination assay using Ub-Rhodamine 110 as a model substrate. Consistent with previous studies21, we detected a deubiquitinase activity in Otu that was ectopically expressed and purified from S2 cells (Fig S6). The presence of Sakura, purified from E.coli, did not affect Otu’s deubiquitinase activity under these assay conditions. Additionally, Sakura itself did not exhibit deubiquitinase activity.

Taken together, these findings show that otu and sakura phenocopy each other in ovaries, specifically in regulating germ cell maintenance and differentiation, raising the possibility that they function together. Similar to sakura, loss of otu inhibits Dpp/BMP signaling and results in the de-repression of bam, leading to aberrant germline differentiation.

Discussion

In this study, we identified Sakura, encoded by a previously uncharacterized gene CG14545, as an essential factor for oogenesis and female fertility. We demonstrated that Sakura is specifically expressed in germline cells in the ovary, including GSCs, is localized to the cytoplasm, and is enriched in the developing oocytes. sakura homozygous null mutant flies are viable but completely female-sterile and male-fertile. Loss of sakura, either through null mutation or germline RNAi including in GSCs, results in rudimentary ovaries exhibiting germless and tumorous phenotypes.

The tumorous phenotype associated with the loss of sakura is characterized by an excess of GSC-like cells, which feature round spectrosomes, as well as cyst cells with branched fusomes (Fig 3C). The increased number of GSC-like cells in loss of sakura suggests dysregulation of GSC self-renewal and differentiation. Meanwhile, the presence of excess cyst cells with branched fusomes indicates abnormal differentiation and division of cysts. Differentiation of cystoblasts typically involves four mitotic divisions with incomplete cytokinesis, leading to 16 cyst cells interconnected by branched fusomes7,46. The degree of fusome branching serves as a marker for the stages of cyst cell division, with increased branching from 2-cell to 16-cell cysts47–49. Notably, fusomes begin to degenerate and disappear after 16-cell cysts as the germline cyst enters the meiotic zone or region 2 of the germarium50,51. The persistence of cyst cells with branched fusomes in sakuranull/null suggests that sakura is crucial for the proper division and differentiation of cysts, and we speculate that it is required for germline cysts to enter meiotic division.

Mosaic analysis of sakuranull indicates that sakura is intrinsically required for establishing GSCs in the ovaries. A significantly reduced number of sakuranull marked GSCs were observed at the adult stage when mutant clones were induced in PGC stages, suggesting that many sakura mutant PGCs die before developing into GSCs. Furthermore, after inducing sakuranull clones in adult ovaries, the number of marked sakuranull GSCs decreased at a higher rate than controls, indicating a rapid loss of sakuranull GSCs (Fig 6A). Thus, sakura is also intrinsically necessary for maintaining GSCs. Additionally, germaria containing sakuranull GSCs became increasingly tumorous over time, with a growing number of sakuranull GSC-like cells (Fig 6B and 6C). This suggests that sakuranull GSCs undergo uncontrolled division, becoming more tumorous over time, which eventually triggers an intrinsic cell death mechanism, as evidenced by elevated levels of cleaved Caspase-3 in sakuranull ovaries (Fig 3B).

Dpp/BMP signaling governs GSC self-renewal and cystoblast differentiation by repressing bam in GSCs and de-repressing it in daughter cystoblasts (Fig S1). This process is mediated by the transcription factor Mad, which, when phosphorylated (pMad), translocates into the nucleus to repress bam transcription7,46,52. Our findings indicate that Bam expression is not limited to 8-cell cysts but continues throughout the germarium in ovaries lacking sakura function (Fig 7A and 7B). The persistent expression of Bam is likely due to low levels of pMad in GSCs (Fig 7C, and 7D). Given the low pMad levels observed upon loss of sakura, we believe that the misregulation of Bam expression occurs primarily at the transcriptional level, although it has also been reported that Bam expression can be regulated post-transcriptionally53.

Transposons are mobile genetic elements that, if not silenced, can generate DNA damage and genomic instability54. piRNAs derived from transposons and other repeats can target and silence transposon RNAs to preserve genome integrity in germ cells31. Loss of piRNAs leads to transposon derepression, resulting in increased DNA damage, which subsequently triggers cell death55,56. Genetic damage in germ cells can cause developmental defects and diseases that may be inherited by the next generation. Thus, elimination of defective germ cells is crucial for maintaining germline integrity of a species57,58. We found that loss of sakura results in reduced piRNA levels and loss of piRNA-mediated transposon silencing in the germline (Fig 4). The observed apoptosis, indicated by elevated cleaved Caspase-3 levels in sakura mutant ovaries (Fig 3B), suggest that desilencing of transposon due to reduced piRNA levels likely results in increased DNA damage, triggering cell death. We speculate that the germless phenotype in sakura mutants may partly arise from an apoptotic germline elimination program activated to maintain germline integrity. Loss of piwi, a key component of the piRNA pathway, similarly causes GSC-like tumors, as well as reduced bam expression and elevated pMad expression29. While downregulation of bam and upregulation of pMad in piwi mutant is opposite to upregulation of bam and downregulation of pMad in sakura mutant, defects in the piRNA pathway might be a shared underlying cause for the dysregulation of bam and pMad levels in both cases, leading to tumorous phenotype.

In addition to its expression in GSCs and cysts in the germarium, Sakura is also expressed in germline cells in later-stage egg chambers (Fig 1E-F) and is required at this stage for proper oogenesis. When Sakura is depleted from region 2b of the germarium onward, while leaving GSCs and early cyst cells intact, female flies lay significantly fewer eggs and have fewer stage 14 oocytes (Fig 5). The failure to produce stage 14 oocytes likely results from the Orb mislocalization in developing oocytes (Fig 5E). Orb is an RNA-binding protein crucial for egg chamber formation, oocyte specification, and polarity establishment during oogenesis by regulating mRNA localization and translation, such as that of oskar, which is vital for establishing oocyte anterior-posterior polarity34,59. Interestingly, piRNA pathway mutants exhibit Orb mislocalization in developing oocytes60. Therefore, reduced piRNA levels observed (Fig 4) may underlie the Orb mislocalization in loss of sakura (Fig 5E).

Sakura does not possess any known protein domains, complicating investigations into its molecular function. To infer its function, we identified interacting proteins through immunoprecipitation followed by mass spectrometry. We identified Otu, a protein known to be crucial for oogenesis, as a protein partner of Sakura. Mutations in otu gene lead to a range of ovarian phenotypes, including germ cell loss, tumorous egg chambers filled with undifferentiated germ cells, abnormalities in nurse cell chromosome structure, and defects in oocyte determination22,23. We showed that germline depletion of Otu via RNAi phenocopies loss of sakura. Similar to Sakura, loss of otu inhibits Dpp/BMP signaling, resulting in low pMad levels in GSCs and Bam de-repression (Fig 12A-B). Additionally, loss of otu also results in the loss of piRNA-mediated silencing, paralleling the effects of sakura loss. Furthermore, germline knockdown of otu yielded a similar ratio of germless ovaries as seen in sakura-RNAi, and simultaneous knockdown of both otu and sakura exacerbated the germless ovary phenotype (Fig 8). These observations raise the possibility that Sakura and Otu function together to regulate germ cell maintenance in the ovaries and are involved in Dpp/BMP signaling to balance stem cell renewal and differentiation.

Otu possesses deubiquitinase activity, catalyzed by its N-terminal Otu domain21,61. In germ cells, Otu interacts with Bam, forming a deubiquitinase complex that deubiquitinates and thereby stabilizes CycA, promoting GSC differentiation21. The predicted structure of Sakura and Otu complex suggests that the Mid domain of Sakura directly contacts the Otu domain of Otu (Fig 10A). We found that Sakura lacks a deubiquitinase activity and it does not directly affect Otu’s deubiquitinase rate in our in vitro assays using Ub-Rhodamine 110 as a substrate (Fig S6). This does not preclude the possibility that Sakura may still influence Otu’s deubiquitinase activity, potentially guiding Otu to its substrate and determining substrate specificity. Our in vitro assays may not have detected such specificity changes. Otu is also an RNA-binding protein, and its deubiquitinase activity can be enhanced by bound RNAs7. Future studies should explore whether Sakura regulates Otu’s RNA-binding or other RNA-related functions. Notably, while Otu is expressed in various tissues, including testes and gut42,61, Sakura is exclusively expressed in ovaries, particularly in germline cells such as GSCs, Therefore, Sakura may enhance Otu’s functions specific to female germline cells. We believe that identifying Otu’s deubiquitinase substrates and its RNA partners, along with investigating Sakura’s role in these processes, will deepen our understanding of Sakura’s and Otu’s molecular functions in oogenesis. One speculative hypothesis is that Sakura may regulate Otu’s deubiquitinase activity concerning piRNA pathway proteins crucial for proper piRNA production and function.

In summary, this study identifies and characterizes the previously unknown gene sakura, which is specifically expressed in female germ cells and is essential for oogenesis and female fertility. Together with its protein partner Otu, Sakura likely regulates germline cell fate, maintenance, and differentiation

Materials and methods

Fly strains

We generated the sakuranull strain by introducing indels within the Sakura coding region using the CRISPR/Cas9 genome editing system, as we previously reported26,62–65. The transgenic sakura-EGFP, otu-EGFP, and otu(Δ-Tudor)-EGFP strains were created following previously published methods65,66. A fragment containing Sakura cDNA flanked by ∼1 kbp upstream and ∼1 kbp downstream genomic sequences was cloned. Fragments containing Otu and Otu(Δ-Tudor) cDNAs, flanked by ∼2.5 kbp upstream and ∼1 kbp downstream genomic sequences, were also cloned. The EGFP gene was fused in-frame to the C-terminus of the Sakura and Otu coding sequences within these fragments, which were then inserted into a pattB plasmid vector. The sakura-EGFP plasmid was integrated into the 51C1 site within the fly genome using the BDSC:24482 fly strain and the PhiC31 system, while the otu-EGFP and otu(Δ-Tudor)-EGFP plasmids were integrated into the 25C6 site of the genome using the attP40 fly strain and the PhiC31 system.

The sakura RNAi #1 (VDRC: v39727) and #2 (VDRC: v103660), y-RNAi (v106068), TOsk-Gal4 (v 314033), Burdock sensor [UAS-Dcr2; NGT-Gal4; nosGal4-VP16, nos>NLS_GFP_lacZ_vas-3’UTR_burdock-target] (v 313217) strains were from Vienna Drosophila Resource Center. w-RNAi (BDSC: 35573), otu-RNAi (BDSC: 34065), bam-RNAi (BDSC: 33631), cycA-RNAi (BDSC: 29313), UAS-Dcr-2; NGT-Gal4 (BDSC: 25751), FRT82B/TM6C, Sb (BDSC: 86313) and otu14 (BDSC: 6025) were obtained from the Bloomington Stock Center. The bam-GFP reporter (DGRC: 118177) and vasa-EGFP knocked-in fly (DGRC: 118616) were from Kyoto Drosophila Stock Center.

Fertility assay

The female fertility assay was performed as previously described64,65,67,68. Briefly, five test virgin females of each test genotype were mated with three wild-type (OregonR) males in a cage with a 6-cm grape juice agar plate supplemented with wet yeast paste. The agar plates were replaced daily, and the number of eggs laid on the third plate (on day 3) was recorded. After incubation at 25°C for an additional day, the number of hatched eggs was counted. At least three cages per genotype were tested.

For the male fertility assay, a single test male was mated with five wild-type (OregonR) virgin females in each vial, following previously published methods64,65,67. After 3 days, the females were transferred to a new vial (vial 1). Every two days, they were transferred to a new vial until a total of four vials were obtained. After 2 days in the fourth vial, the females were removed, and the total number of progenies emerging from these four vials was counted. At least five males per genotype were tested.

Sakura and Otu antibodies

We expressed a recombinant full-length Sakura protein as an N-terminal 6xHis-tagged protein in E. coli using a modified pET vector and purified it using Ni-sepharose (GE Healthcare) and HiTrapQ HP (GE Healthcare) columns, as we previously performed for other proteins64,65. Recombinant Otu fragments (145-405aa and 406-853aa) were similarly expressed as N-terminally 6xHis-MBP-fusion proteins in E. coli and purified using Ni-sepharose. These purified proteins were used as antigens to generate polyclonal anti-Sakura and anti-Otu sera in rabbits (Pocono Rabbit Farm & Laboratory, Inc.). The rabbit polyclonal anti-Sakura antibodies were affinity purified using His-MBP-Sakura recombinant protein and Affigel-15 (Bio-rad), following the manufacturer’s instructions. Rabbit anti-Otu sera were first pre-cleared with purified His-MBP protein bound to Affigel-15, and the unbound fraction was affinity purified using purified His-MBP-Otu fragments (145-405aa and 406-853aa) and Affigel-15.

Germline clonal analysis

We generated sakuranull mutant clones using FLP/FRT-mediated recombination69. To create sakuranullGSC clones, 3-day-old female flies with the genotype hs-flp; +; FRT82B, ubi-GFP/FRT82B, sakuranull were heat-shocked at 37°C for 1 hour, twice daily, with an 8-hour interval between heat shocks. Female flies with the genotype hs-flp; +; FRT82B, ubi-GFP/FRT82B were used as controls. Flies were dissected and stained 4, 7, and 14 days after clone induction. To induce primordial germ cell (PGC) clones, early third-instar larvae were heat-shocked similarly, allowed to develop, and dissected and analyzed at 3 to 4 days post-eclosion.

Immunostaining

Stereomicroscope images of dissected ovaries were taken using Leica M125 stereomicrocsope64. Ovaries from 2- to 5-day-old, yeast-fed females were hand-dissected in 1X PBS (137 mM NaCl, 2.7 mM KCl, 10 mM Na2HPO4, 1.8 mM KH2PO4, pH 7.4) at room temperature. The dissected ovaries were fixed in a fixative buffer (4% formaldehyde, 15 mM PIPES (pH 7.0), 80 mM KCl, 20 mM NaCl2, 2 mM EDTA, and 0.5 mM EGTA), incubated for 30 minutes at room temperature with gentle rocking. After fixation, the ovaries were rinsed three times with PBX (0.1% Triton X-100 in 1X PBS) and then incubated in a blocking buffer (2% donkey serum in 3% BSA [w/v], 0.02% NAN3 [w/v] in PBX) for 1 hour at room temperature. Then, the ovaries were incubated with primary antibodies diluted in the blocking buffer overnight at 4°C. The following day, the ovaries were rinsed three times with PBX and incubated with Alexa Fluor-conjugated secondary antibodies for 2 hours. The ovaries were rinsed three times with PBX and then mounted in VECTASHIELD® PLUS antifade mounting medium with DAPI (H-2000, Vector lab). Confocal images were acquired on a Zeiss LSM700 confocal microscope at the Johns Hopkins University School of Medicine Microscope Facility.

The primary antibodies used for immunostaining were mouse anti-HTS (1B1) (DSHB, AB_528070, 1/100), mouse anti-Bam (DSHB, AB_10570327, 1/20), mouse anti-CycA (DSHB, AB_528188, 1/100), rat anti-Vasa (DSHB, AB_760351, 1/100), rabbit anti-pMad (Phospho-SMAD1/5 (Ser463/465) mAb #9516, dilution: 1/200, Cell Signaling), and rabbit anti-cleaved caspase-3 (Cleaved Caspase-3 (Asp175) #9661, dilution: 1/200, Cell Signaling). Secondary antibodies used were Alexa Fluor 488 Donkey anti-Mouse Igg (ThermoFisher A21202, 1/100), Alexa Fluor 594 Donkey anti-Rat Igg (ThermoFisher A21209, 1/100), Alexa Fluor 594 Donkey anti-Mouse Igg (ThermoFisher A21203, 1/100), and Alexa Fluor 594 Donkey anti-Rabbit Igg ThermoFisher A21207, 1/100).

Western blot

Lysates of hand-dissected ovaries and tissues were prepared by homogenizing in RIPA buffer (50 mM Tris-HCl [pH 7.4], 150 mM NaCl, 1% [v/v] IGEPAL CA-630, 0.1% [w/v] sodium dodecyl sulfate (SDS), 0.5% [w/v] sodium deoxycholate, 1 mM ethylenediaminetetraacetic acid (EDTA), 5 mM dithiothreitol, and 0.5 mM phenylmethylsulfonyl fluoride (PMSF))63–65,67. The homogenates were centrifuged at 21,000g at 4°C for 10 min, and the protein concentration of the supernatant was determined using the BCA protein assay kit (Pierce) as needed. Fifteen µg of total protein was loaded per lane for Western blot. The sources and dilutions of the primary antibodies were as below. Rabbit anti-Sakura (1/10000, generated in this study), rabbit anti-Otu (1/10000, generated in this study), rabbit anti-alpha-Tubulin [EP1332Y] (1/10000, Abcam, ab52866), mouse anti-alpha-Tubulin [12G10] (1/10000, DSHB, AB_1157911), mouse anti-FLAG (1/10000, Sigma, F1804), mouse anti-HA (1/10000, Sigma, H3663), and mouse anti-GFP [GF28R] (1/3000, Invitrogen, 14-6674-82). IRDye 800CW goat anti-mouse IgG, IRDye 800CW goat anti-rabbit IgG, IRDye 680RD goat anti-mouse, and IgG IRDye 680RD goat anti-rabbit were used as secondary antibodies. The membranes were scanned using the Li-Cor Odyssey CLx Imaging System.

Mass spectrometry

Immunoprecipitation of Sakura-EGFP protein was performed using the GFP-Trap Magnetic Agarose Kit (Proteintech, gtmak-20) on dissected ovaries from flies harboring sakura-EGFP transgene, with w1118 flies as controls. Ovaries were homogenized in 200 µL ice-cold lysis buffer (10 mM Tris-HCl [pH 7.5], 150 mM NaCl, 0.5 mM EDTA, 0.05% [v/v] IGEPAL CA-630) containing 1× protease inhibitor cocktail (100× protease inhibitor cocktail contains 120 mg/ml 1 mM 4-(2-aminoethyl) benzene sulfonyl fluoride hydrochloride (AEBSF), 1 mg/ml aprotinin, 7 mg/ml bestatin, 1.8 mg/ml E-64, and 2.4 mg/ml leupeptin)70. After homogenization, the tubes were placed on ice for 30 minutes, and the homogenates were extensively pipetted every 10 minutes. The lysates were then centrifuged at 17,000x g for 10 minutes at 4°C. The supernatants were transferred to pre-chilled tubes, and 300 µL dilution buffer (10 mM Tris/Cl pH 7.5, 150 mM NaCl, 0.5 mM EDTA) supplemented with 1x protease inhibitor cocktail were added. The diluted lysates were then added to the GFP-trap magnetic beads in 1.5 mL tubes and rotated for 1 hour at 4°C. After separating the beads with a magnetic tube rack, the beads were washed three times with 500 µL wash buffer (10 mM Tris/Cl pH 7.5, 150 mM NaCl, 0.05 % [v/v] IGEPAL CA-630). Proteins were eluted with 40 µL acidic elution buffer (200 mM glycine pH 2.5) followed by immediate neutralization with 5 uL neutralization buffer (1 M Tris pH 10.4).

As a quality control before mass spectrometry, ∼5 µL of the samples were mixed with an equal volume of 2× SDS-PAGE loading buffer (80 mM Tris-HCl [pH 6.8], 2% [w/v] SDS, 10% [v/v] glycerol, 0.0006% [w/v] bromophenol blue, 2% [v/v] 2-mercaptoethanol), heated at 95 °C for 3 min, and run on 4–20% Mini-PROTEAN® TGX™ Precast Protein Gels (Biorad, #4561094). Silver staining was then performed by using Pierce™ Silver Stain Kit (ThermoFisher, 24612) to assess the quality of the immunoprecipitated protein samples. Mass spectrometry was conducted at the Mass Spectrometry Core at the Department of Biological Chemistry, Johns Hopkins School of Medicine, as previously described65

Co-immunoprecipitation

For co-immunoprecipitation of endogenous Sakura protein, wild-type ovaries (w1118) were homogenized in RIPA buffer, centrifuged at 21,000g at 4°C for 10 min, and the clear supernatant protein lysates were used for immunoprecipitation. Four µg of rabbit anti-Sakura and normal rabbit IgG (Cell Signaling, #2729) were incubated with 50 µL of Dynabeads Protein G (ThermoFisher, 10004D) for 20 min at room temperature. The beads were washed once with PBST (1x PBS with 0.1% Tween-20). The ovary lysate supernatant was then incubated with the washed beads at room temperature for 30 min, followed by three washes with PBST. The proteins were eluted with 2x SDS-PAGE loading buffer (80 mM Tris-HCl [pH 6.8], 2% [w/v] SDS, 10% [v/v] glycerol, 0.0006% [w/v] bromophenol blue, 2% [v/v] 2-mercaptoethanol) and heated at 70°C for 10 min. After bead separation using a magnetic tube rack, the eluted proteins in 2x SDS-PAGE loading buffer were heated again at 95°C for 3 min.

Transient protein expression in S2 cells was performed using the pAc5.1/V5-HisB plasmid vector (Invitrogen). A total of 1 µg plasmids were transfected using the Effectene transfection reagent (Qiagen, 301425). Three days after transfection, cells were harvested and lysed with ice-cold lysis buffer supplemented with 1× protease inhibitor cocktail. The cell lysates were then centrifuged at 17,000x g for 10 min at 4°C, and the clear supernatants were collected for immunoprecipitation. For anti-HA immunoprecipitation, supernatants were incubated with 25 µL (0.25 mg) of Pierce™ Anti-HA Magnetic Beads (ThermoFisher, 88837) at room temperature for 30 min. The beads were then washed three times with TBST (1x TBS [0.05 M Tris/HCl and 0.15 M NaCl, pH 7.6] with 0.05% Tween-20). For anti-FLAG immunoprecipitation, 2 µg of mouse anti-FLAG (1/10000, Sigma, F1804) was incubated with 50 µL of Dynabeads Protein G (ThermoFisher, 10004D) for 10 min at room temperature. The beads were washed once with PBST. The S2 cell lysate supernatant was incubated with the washed beads at room temperature for 15 min. The beads were washed three times with PBST. In both anti-HA and anti-FLAG immunoprecipitations, proteins were eluted with 2× SDS-PAGE loading buffer. For anti-HA, the beads in 2x SDS-PAGE loading buffer were heated at 95°C for 7 min. For anti-FLAG, the beads were heated at 70°C for 10 min, then were separated using a magnetic tube rack. The eluted proteins in 2x SDS-PAGE loading buffer were heated again at 95°C for 3 min.

Small RNA and mRNA sequencing

Poly-A+ mRNA purification was performed as previously described71. Small RNA libraries and poly-A+ mRNA libraries were prepared, sequenced on Hiseq2500 (Illumina), and analyzed, as previously reported72,73. SRA accession number for these datasets is PRJNA1156618.

In vitro deubiquitination assay

C-terminally 6xHis-tagged recombinant Sakura protein was expressed in E. coli using a modified pET vector and was purified using Ni-sepharose. N-terminally 3xHA-HRV3Csite-3xFLAG-tagged Otu protein and a negative control, N-terminally 3xHA-HRV3Csite-3xFLAG-tagged firefly luciferase protein, were expressed in S2 cells using pAc5.1/V5-HisB vector. The proteins were immunoprecipitated using Pierce™ Anti-HA Magnetic Beads (ThermoFisher, 88837) as described previously, with a more stringent washing step to effectively remove interacting proteins. The beads were washed six times with a high salt wash buffer (TBST with 800 mM NaCl). The proteins were eluted by cleaving off 3xHA tag with 25 nM GST-HRV3C protease in a cleavage buffer (25mM Tris-HCl pH 7.4, 150 mM NaCl, 5% glycerol, 2mM EDTA), incubated at 4°C with rotation for 6 hours. An equal volume of 100% glycerol was added to the eluted protein.

Purified proteins, including Otu, luciferase, and Sakura, were mixed with Ub-Rhodamine 110 (Ubiquitin-Proteasome Biotechnologies, M3020) in a total volume of 30 µL in reaction buffer (20 mM Tris-HCl, pH 7.5, 200 mM NaCl, 5 mM MgCl2, 2 mM DTT). The mixture was added to a black 384-well low-volume plate. Fluorescence measurements were taken using SpectraMax i3x Multi-Mode Microplate Reader at 37°C, with excitation and emission wavelengths set at 485/20 and 530/20 nm respectively. The fluorescence intensity for each condition was averaged from triplicates and plotted as a function of time.

eLife Assessment

This valuable study reports the first characterization of the CG14545 gene in Drosophila melanogaster, which the authors name "Sakura." Acting during germline stem cell fate and differentiation, Sakura is required for both oogenesis and female fertility. The evidence supporting the claims of the authors is solid, but the manuscript would be strengthened by a more in-depth investigation into the cause-and-effect relationships for the different defects observed.

Reviewer #1 (Public review):

Summary:

In this manuscript, Azlan et al. identified a novel maternal factor called Sakura that is required for proper oogenesis in Drosophila. They showed that Sakura is specifically expressed in the female germline cells. Consistent with its expression pattern, Sakura functioned autonomously in germline cells to ensure proper oogenesis. In Sakura KO flies, germline cells were lost during early oogenesis and often became tumorous before degenerating by apoptosis. In these tumorous germ cells, piRNA production was defective and many transposons were derepressed. Interestingly, Smad signaling, a critical signaling pathway for GSC maintenance, was abolished in sakura KO germline stem cells, resulting in ectopic expression of Bam in whole germline cells in the tumorous germline. A recent study reported that Bam acts together with the deubiquitinase Otu to stabilize Cyc A. In the absence of sakura, Cyc A was upregulated in tumorous germline cells in the germarium. Furthermore, the authors showed that Sakura co-immunoprecipitated Otu in ovarian extracts. A series of in vitro assays suggested that the Otu (1-339 aa) and Sakura (1-49 aa) are sufficient for their direct interaction. Finally, the authors demonstrated that the loss of otu phenocopies the loss of sakura, supporting their idea that Sakura plays a role in germ cell maintenance and differentiation through interaction with Otu during oogenesis.

Strengths:

To my knowledge, this is the first characterization of the role of CG14545 genes. Each experiment seems to be well-designed and adequately controlled.

Weaknesses:

However, the conclusions from each experiment are somewhat separate, and the functional relationships between Sakura's functions are not well established. In other words, although the loss of Sakura in the germline causes pleiotropic effects, the cause-and-effect relationships between the individual defects remain unclear.

Reviewer #2 (Public review):

In this study, the authors identified CG14545 (and named it Sakura), as a key gene essential for Drosophila oogenesis. Genetic analyses revealed that Sakura is vital for both oogenesis progression and ultimate female fertility, playing a central role in the renewal and differentiation of germ stem cells (GSC).

The absence of Sakura disrupts the Dpp/BMP signaling pathway, resulting in abnormal bam gene expression, which impairs GSC differentiation and leads to GSC loss. Additionally, Sakura is critical for maintaining normal levels of piRNAs. Also, the authors convincingly demonstrate that Sakura physically interacts with Otu, identifying the specific domains necessary for this interaction, suggesting a cooperative role in germline regulation. Importantly, the loss of otu produces similar defects to those observed in Sakura mutants, highlighting their functional collaboration.

The authors provide compelling evidence that Sakura is a critical regulator of germ cell fate, maintenance, and differentiation in Drosophila. This regulatory role is mediated through the modulation of pMad and Bam expression. However, the phenotypes observed in the germarium appear to stem from reduced pMad levels, which subsequently trigger premature and ectopic expression of Bam. This aberrant Bam expression could lead to increased CycA levels and altered transcriptional regulation, impacting piRNA expression. Given Sakura's role in pMad expression, it would be insightful to investigate whether overexpression of Mad or pMad could mitigate these phenotypic defects (UAS-Mad line is available at Bloomington Drosophila Stock Center).

A major concern is the overstated role of Sakura in regulating Orb. The data does not reveal mislocalized Orb; rather, a mislocalized oocyte and cytoskeletal breakdown, which may be secondary consequences of defects in oocyte polarity and structure rather than direct misregulation of Orb. The conclusion that Sakura is necessary for Orb localization is not supported by the data. Orb still localizes to the oocyte until about stage 6. In the later stage, it looks like the cytoskeleton is broken down and the oocyte is not positioned properly, however, there is still Orb localization in the ~8-stage egg chamber in the oocyte. This phenotype points towards a defect in the transport of Orb and possibly all other factors that need to localize to the oocyte due to cytoskeletal breakdown, not Orb regulation directly. While this result is very interesting it needs further evaluation on the underlying mechanism. For example, the decrease in E-cadherin levels leads to a similar phenotype and Bam is known to regulate E-cadherin expression. Is Bam expressed in these later knockdowns?

The manuscript would benefit from a more balanced interpretation of the data concerning Sakura's role in Orb regulation. Furthermore, a more expanded discussion on Sakura's potential role in pMad regulation is needed. For example, since Otu and Bam are involved in translational regulation, do the authors think that Mad is not translated and therefore it is the reason for less pMad? Currently the discussion presents just a summary of the results and not an extension of possible interpretation discussed in context of present literature.

Reviewer #3 (Public review):

In this very thorough study, the authors characterize the function of a novel Drosophila gene, which they name Sakura. They start with the observation that sakura expression is predicted to be highly enriched in the ovary and they generate an anti-sakura antibody, a line with a GFP-tagged sakura transgene, and a sakura null allele to investigate sakura localization and function directly. They confirm the prediction that it is primarily expressed in the ovary and, specifically, that it is expressed in germ cells, and find that about 2/3 of the mutants lack germ cells completely and the remaining have tumorous ovaries. Further investigation reveals that Sakura is required for piRNA-mediated repression of transposons in germ cells. They also find evidence that sakura is important for germ cell specification during development and germline stem cell maintenance during adulthood. However, despite the role of sakura in maintaining germline stem cells, they find that sakura mutant germ cells also fail to differentiate properly such that mutant germline stem cell clones have an increased number of "GSC-like" cells. They attribute this phenotype to a failure in the repression of Bam by dpp signaling. Lastly, they demonstrate that sakura physically interacts with otu and that sakura and otu mutants have similar germ cell phenotypes. Overall, this study helps to advance the field by providing a characterization of a novel gene that is required for oogenesis. The data are generally high-quality and the new lines and reagents they generated will be useful for the field. However, there are some weaknesses and I would recommend that they address the comments in the Recommendations for the authors section below.

Author response:

The following is the authors’ response to the original reviews

Public Reviews:

Reviewer #1 (Public review):

Summary:

In this manuscript, Azlan et al. identified a novel maternal factor called Sakura that is required for proper oogenesis in Drosophila. They showed that Sakura is specifically expressed in the female germline cells. Consistent with its expression pattern, Sakura functioned autonomously in germline cells to ensure proper oogenesis. In Sakura KO flies, germline cells were lost during early oogenesis and often became tumorous before degenerating by apoptosis. In these tumorous germ cells, piRNA production was defective and many transposons were derepressed. Interestingly, Smad signaling, a critical signaling pathway for GSC maintenance, was abolished in sakura KO germline stem cells, resulting in ectopic expression of Bam in whole germline cells in the tumorous germline. A recent study reported that Bam acts together with the deubiquitinase Otu to stabilize Cyc A. In the absence of sakura, Cyc A was upregulated in tumorous germline cells in the germarium. Furthermore, the authors showed that Sakura co-immunoprecipitated Otu in ovarian extracts. A series of in vitro assays suggested that the Otu (1-339 aa) and Sakura (1-49 aa) are sufficient for their direct interaction. Finally, the authors demonstrated that the loss of otu phenocopies the loss of sakura, supporting their idea that Sakura plays a role in germ cell maintenance and differentiation through interaction with Otu during oogenesis.

Strengths:

To my knowledge, this is the first characterization of the role of CG14545 genes. Each experiment seems to be well-designed and adequately controlled.

Weaknesses:

However, the conclusions from each experiment are somewhat separate, and the functional relationships between Sakura's functions are not well established. In other words, although the loss of Sakura in the germline causes pleiotropic effects, the cause-and-effect relationships between the individual defects remain unclear.

Reviewer #2 (Public review):

In this study, the authors identified CG14545 (and named it Sakura), as a key gene essential for Drosophila oogenesis. Genetic analyses revealed that Sakura is vital for both oogenesis progression and ultimate female fertility, playing a central role in the renewal and differentiation of germ stem cells (GSC).

The absence of Sakura disrupts the Dpp/BMP signaling pathway, resulting in abnormal bam gene expression, which impairs GSC differentiation and leads to GSC loss. Additionally, Sakura is critical for maintaining normal levels of piRNAs. Also, the authors convincingly demonstrate that Sakura physically interacts with Otu, identifying the specific domains necessary for this interaction, suggesting a cooperative role in germline regulation. Importantly, the loss of otu produces similar defects to those observed in Sakura mutants, highlighting their functional collaboration.

The authors provide compelling evidence that Sakura is a critical regulator of germ cell fate, maintenance, and differentiation in Drosophila. This regulatory role is mediated through the modulation of pMad and Bam expression. However, the phenotypes observed in the germarium appear to stem from reduced pMad levels, which subsequently trigger premature and ectopic expression of Bam. This aberrant Bam expression could lead to increased CycA levels and altered transcriptional regulation, impacting piRNA expression. Given Sakura's role in pMad expression, it would be insightful to investigate whether overexpression of Mad or pMad could mitigate these phenotypic defects (UAS-Mad line is available at Bloomington Drosophila Stock Center).

As suggested reviewer 1, we tested whether overexpression of Mad could rescue or mitigate the loss of sakura phenotypic defects, by using nos-Gal4-VP16 > UASp-Mad-GFP in the background of sakuranull. As shown in Fig S11, we did not observe any mitigation of defects.

Then, we also tested whether expressing a constitutive active form of Tkv, by using UAS-Dcr2, NGT-Gal4 > UASp-tkv.Q235D in the background of sakuraRNAi. As shown in Fig S12, we did not observe any mitigation of defects by this approach either.

A major concern is the overstated role of Sakura in regulating Orb. The data does not reveal mislocalized Orb; rather, a mislocalized oocyte and cytoskeletal breakdown, which may be secondary consequences of defects in oocyte polarity and structure rather than direct misregulation of Orb. The conclusion that Sakura is necessary for Orb localization is not supported by the data. Orb still localizes to the oocyte until about stage 6. In the later stage, it looks like the cytoskeleton is broken down and the oocyte is not positioned properly, however, there is still Orb localization in the ~8-stage egg chamber in the oocyte. This phenotype points towards a defect in the transport of Orb and possibly all other factors that need to localize to the oocyte due to cytoskeletal breakdown, not Orb regulation directly. While this result is very interesting it needs further evaluation on the underlying mechanism. For example, the decrease in E-cadherin levels leads to a similar phenotype and Bam is known to regulate E-cadherin expression. Is Bam expressed in these later knockdowns?

We examined Bam and DE-Cadherin expression in later RNAi knockdowns driven by ToskGal4. As shown in Fig S9, Bam was not expressed in these later knockdowns compared with controls. DE-Cadherin staining suggested a disorganized structure in late-stage egg chambers.

We agree that we overstated a role of Sakura in regulating Orb in the initial manuscript. We changed the text to avoid overstating.

The manuscript would benefit from a more balanced interpretation of the data concerning Sakura's role in Orb regulation. Furthermore, a more expanded discussion on Sakura's potential role in pMad regulation is needed. For example, since Otu and Bam are involved in translational regulation, do the authors think that Mad is not translated and therefore it is the reason for less pMad? Currently the discussion presents just a summary of the results and not an extension of possible interpretation discussed in context of present literature.

We changed the text to avoid overstating a role of Sakura in regulating Orb localization.

Based on our newly added results showing that transgenic overexpression of Mad could not rescue or mitigate the phenotypic defects of sakuranull mutant (Fig S11), we do not think the reason for less pMad is less translation of Mad.

Reviewer #3 (Public review):

In this very thorough study, the authors characterize the function of a novel Drosophila gene, which they name Sakura. They start with the observation that sakura expression is predicted to be highly enriched in the ovary and they generate an anti-sakura antibody, a line with a GFP-tagged sakura transgene, and a sakura null allele to investigate sakura localization and function directly. They confirm the prediction that it is primarily expressed in the ovary and, specifically, that it is expressed in germ cells, and find that about 2/3 of the mutants lack germ cells completely and the remaining have tumorous ovaries. Further investigation reveals that Sakura is required for piRNA-mediated repression of transposons in germ cells. They also find evidence that sakura is important for germ cell specification during development and germline stem cell maintenance during adulthood. However, despite the role of sakura in maintaining germline stem cells, they find that sakura mutant germ cells also fail to differentiate properly such that mutant germline stem cell clones have an increased number of "GSC-like" cells. They attribute this phenotype to a failure in the repression of Bam by dpp signaling. Lastly, they demonstrate that sakura physically interacts with otu and that sakura and otu mutants have similar germ cell phenotypes. Overall, this study helps to advance the field by providing a characterization of a novel gene that is required for oogenesis. The data are generally high-quality and the new lines and reagents they generated will be useful for the field. However, there are some weaknesses and I would recommend that they address the comments in the Recommendations for the authors section below.

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

General Comments:

(1) The gene nomenclature: As mentioned in the text, Sakura means cherry blossom and is one of the national flowers of Japan. I am not sure whether the phenotype of the CG14545 mutant is related to Sakura or not. I would like to suggest the authors reconsider the naming.

The striking phenotype of sakura mutant­ is tumorous and germless ovarioles. The tumorous phenotype, exhibiting lots of round fusome in germarium visualized by anti-Hts staining, looks like cherry blossom blooming to us. Also, the germless phenotype reminds us falling of the cherry blossom, especially considering that the ratio of tumorous phenotype decreases and that of germless decreases over fly age. Furthermore, “Sakura” symbolizes birth and renewal in Japanese culture (the last author of this manuscript is Japanese). Our findings indicated that the gene sakura is involved in regulation of renewal and differentiation of GSCs (which leads to birth). These are the reasons for the naming, which we would like to keep.

(2) In many of the microscopic photographs in the figures, especially for the merged confocal images, the resolution looks low, and the images appear blurred, making it difficult to judge the authors' claims. Also, the Alpha Fold structure in Figure 10A requires higher contrast images. The magnification of the images is often inadequate (e.g. Figures 3A, 3B, 5E, 7A, etc). The authors should take high-magnification images separately for the germarium and several different stages of the egg chambers and lay out the figures.

We are very sorry for the low-resolution images. This was caused when the original PDF file with high-resolution images was compressed in order to meet the small file size limit in the eLife submission portal. In the revised submission, we used high-resolution images.

Specific Comments

(1) How Sakura can cooperate with Otu remains unanswered. Sakura does not regulate deubiquitinase activity in vitro. Both sakura and otu appear to be involved in the Dpp-Smad signaling pathway and in the spatial control of Bam expression in the germarium, whereas Otu has been reported to act in concert with Bam to deubiquitinate and stabilize Cyc A for proper cystoblast differentiation. Therefore, it is plausible that the stabilization of Cyc A in the Sakura mutant is an indirect consequence of Bam misexpression and independent of the Sakura-Otu interaction. The authors may need to provide much deeper insight into the mechanism by which Sakura plays roles in these seemingly separable steps to orchestrate germ cell maintenance and differentiation during early oogenesis.

Yes, it is possible that the stabilization of CycA in the sakura mutant is an indirect consequence of Bam misexpression and independent of the Sakura-Otu interaction. To test the significance and role of the Sakura-Otu interaction, we have attempted to identify Sakura point mutants that lose interaction with Otu. If such point mutants were successfully obtained, we were planning to test if their transgene expression could rescue the phenotypes of sakura mutant as the wild-type transgene did. However, after designing and testing the interaction of over 30 point mutants with Otu, we could not obtain such mutant version of Sakura yet. We will continue making efforts, but it is beyond the scope of the current study. We hope to address this important point in future studies.

(2) Figure 3A and Figure 4: The authors show that piRNA production is abolished in Sakura KO ovaries. It is known that piRNA amplification (the ping-pong cycle) occurs in the Vasa-positive perinuclear nuage in nurse cells. Is the nuage normally formed in the absence of Sakura? The authors provide high-magnification images in the germarium expressing Vas-GFP. How does Sakura, and possibly Out, contribute to piRNA production? Are the defects a direct or indirect consequence of the loss of Sakura?

We provided higher magnification images of germarium expressing Vasa-EGFP in sakura mutant background (Fig 3A and 3B). The nuage formation does not seem to be dysregulated in sakura mutant. Currently, we do not know if the piRNA defects are direct or indirect consequence of the loss of Sakura. This question cannot be answered easily. We hope to address this in future studies.

(3) Figure 7 and Figure 12: The authors showed that Dpp-Smad signaling was abolished in Sakura KO germline cells. The same defects were also observed in otu mutant ovaries (Figure 12B). How does the Sakura-Otu axis contribute to the Dpp-Smad pathway in the germline?

As we mentioned in the response to comment (1), we attempted to test the significance and role of the Sakura-Otu interaction, including in the Dpp-Smad pathway in the germline, but we have not yet been able to obtain loss-of-interaction mutant(s) of Sakura. We hope to address this in future studies.

(4) Figure 9 and Fig 10: The authors raised antibodies against both Sakura and Otu, but their specificities were not provided. For Western blot data, the authors should provide whole gel images as source data files. Also, the authors argue that the Otu band they observed corresponds to the 98-kDa isoform (lines 302-304). The molecular weight on the Western blot alone would be insufficient to support this argument.

When we submitted the initial manuscript, we also submitted original, uncropped, and unmodified whole Western blot images for all gel images to the eLife journal, as requested. We did the same for this revised submission. I believe eLife makes all those files available for downloading to readers.

In the newly added Fig S13B, we used very young 2-5 hours ovaries and 3-7 days ovaries. 2-5 days ovaries contain only mostly pre-differentiated germ cells. Older ovaries (3-7 days in our case here) contain all 14 stages of oogenesis and later stages predominate in whole ovary lysates.

As reported in previous literature (Sass et al. 1995), we detected a higher abundance of the 104 kDa Otu isoform than the 98 kDa isoform in from 2-5 hours ovaries and predominantly the 98 kDa isoform in 3-7 days ovaries (Fig S13B). These results confirmed that the major Otu isoform we detected in Western blot, all of which uses old ovaries except for the 2-5 hours ovaries in Fig S13B, is the 98 kDa isoform.

(5) Otu has been reported to regulate ovo and Sxl in the female germline. Is Sakura involved in their regulation?

We examined sxl alternative splicing pattern in sakura mutant ovaries. As shown in Fig S6, we detected the male-specific isoform of sxl RNA and a reduced level of the female-specific sxl isoform in sakura mutant ovaries. Thus Sakura seems to be involved in sxl splicing in the female germline, while further studies will be needed to understand whether Sakura has a direct or indirect role here.

(6) Lines 443-447: The GSC loss phenotype in piwi mutant ovaries is thought to occur in a somatic cell-autonomous manner: both piwi-mutant germline clones and germline-specific piwi knockdown do not show the GSC-loss phenotype. In contrast, the authors provide compelling evidence that Sakura functions in the germline. Therefore, the Piwi-mediated GSC maintenance pathway is likely to be independent of the Sakura-Otu axis.

We changed the text accordingly.

Reviewer #2 (Recommendations for the authors):

Overall, this is a cleanly written manuscript, with some sentences/sections that are confusing the way they are constructed (i.e. Line 37-38, 334, section on Flp/FRT experiments).

We rewrote those sections to avoid confusion.

Comment for all merged image data: the quality of the merged images is very poor - the individual channels are better but should also be reprocessed for more resolved image data sets. Also, it would be helpful to have boundaries drawn in an individual panel to identify the regions of the germarium, as cartooned in Figure S1A (which should be brought into Figure 1) F-actin or Vsg staining would have helped throughout the manuscript to enhance the visualization of described phenotypes.

We are very sorry for the low-resolution images. This was caused when the original PDF file with high-resolution images was compressed in order to meet the small file size limit in the eLife submission portal. In the revised submission, we used high-resolution images.

We outlined the germarium in Fig 1E.

We brought the former FigS1 into Fig 1A.

We provided Phalloidin (F-Actin) staining images in Fig S7.

All p-values seem off. I recommend running the data through the student t-test again.

We used the student t-test to calculate p-values and confirmed that they are correct. We don’t understand why the reviewer thinks all p-values seem off.

In the original manuscript, as we mentioned in each figure legends, we used asterisk (*) to indicate p-value <0.05, without distinguishing whether it’s <0.001, <0.01< or <0.05.

Probably reviewer 2 is suggesting us to use ***, **, and *, to indicate p-value of <0.001, <0.01, and <0.05, respectively? If so, we now followed reviewer2’s suggestions.

Figure 1

(1) Within the text, C is mentioned before A.

We updated the text and now we mentioned Fig 1A before Fig 1C.

(2) B should be the supplemental figure.

We moved the former Fig 1B to Supplemental Figure 1.

(3) C - How were the different egg chamber stages selected in the WB? Naming them 'oocytes' is deceiving. Recommend labeling them as 'egg chambers', since an oocyte is claimed to be just the one-cell of that cyst.

We changed the labeling to egg chambers.

(4) Is the antibody not detecting Sakura in IF? There is no mention of this anywhere in the manuscript.

While our Sakura antibody detects Sakura in IF, it seems to detect some other proteins as well. Since we have Sakura-EGFP fly strain (which fully rescues sakuranull phenotypes) to examine Sakura expression and localization without such non-specific signal issues, we relied on Sakura-EGFP rather than anti-Sakura antibodies for IF.

(5) Expand on the reliance of the sakura-EGFP fly line. Does this overexpression cause any phenotypes?

sakura-EGFP does not cause any phenotypes in the background of sakura[+/+] and sakura[+/-].

(6) Line 95 "as shown below" is not clear that it's referencing panel D.

We now referenced Fig 1D.

(7) Re: Figures 1 E and F. There is no mention of Hts or Vasa proteins in the text.

"Sakura-EGFP was not expressed in somatic cells such as terminal filament, cap cells, escort cells, or follicle cells (Figure 1E). In the egg chamber, Sakura-EGFP was detected in the cytoplasm of nurse cells and was enriched in developing oocytes (Figure 1F)". Outline these areas or label these structures/sites in the images. The color of Merge labels is confusing as the blue is not easily seen.

We mentioned Hts and Vasa in the text. We labeled the structures/sites in the images and updated the color labeling.

Figure 2

(1) Entire figure is not essential to be a main figure, but rather supplemental.

We don’t agree with the reviewer. We think that the female fertility assay data, where sakura null mutant exhibits strikingly strong phenotype, which was completely rescued by our Sakura-EGFP transgene, is very important data and we would like to present them in a main figure.

(2) 2A- one star (*) significance does not seem correct for the presented values between 0 and 100+.

In the original manuscript, as we mentioned in each figure legends, we used asterisk (*) to indicate p-value <0.05, without distinguishing whether it’s <0.001, <0.01< or <0.05.

Probably reviewer 2 is suggesting us to use ***, **, and *, to indicate p-value of <0.001, <0.01, and <0.05, respectively? If so, we now followed reviewer2’s suggestions.

(3) 2C images are extremely low quality. Should be presented as bigger panels.

We are very sorry for the low-resolution images. This was caused when the original PDF file with high-resolution images was compressed in order to meet the small file size limit in the eLife submission portal. In the revised submission, we used high-resolution images. We also presented as bigger panels.

Figure 3

(1) "We observed that some sakuranull /null ovarioles were devoid of germ cells ("germless"), while others retained germ cells (Fig 3A)" What is described is, that it is hard to see. Must have a zoomed-in panel.

We provided zoomed-in panels in Fig 3B

(2) C - The control doesn't seem to match. Must zoom in.

We provided matched control and also zoomed in.

(3) For clarity, separate the tumorous and germless images.

In the new image, only one tumorous and one germless ovarioles are shown with clear labeling and outline, for clarity.

(4) Use arrows to help clearly indicate the changes that occur. As they are presented, they are difficult to see.

We updated all the panels to enhance clarity.

(5) Line 158 seems like a strong statement since it could be indirect.

We softened the statement.

Figure 4

(1) Line 188-189 - Conclusion is an overstatement.

We softened the statement.

(2) Is the piRNA reduction due to a change in transcription? Or a direct effect by Sakura?

We do not know the answers to these questions. We hope to address these in future studies.

Figure 5

(1) D - It might make more sense if this graph showed % instead of the numbers.

We did not understand the reviewer’s point. We think using numbers, not %, makes more sense.

(2) Line 213 - explain why RNAi 2 was chosen when RNAi 1 looks stronger.

Fly stock of RNAi line 2 is much healthier than RNAi line 1 (without being driven Gal4) for some reasons. We had a concern that the RNAi line 1 might contain an unwanted genetic background. We chose to use the RNAi 2 line to avoid such an issue.

(3) In Line 218 there's an extra parenthesis after the PGC acronym.

We corrected the error.

(4) TOsk-Gal4 fly is not in the Methods section.

We mentioned TOsk-Gal4 in the Methods.

Figure 6:

(1) The FLP-FRT section must be rewritten.

We rewrote the FLP-FRT section.

(2) A - include statistics.

We included statistics using the chi-square test.

(3) B - is not recalled in the Results text.

We referred Fig 6B in the text.

(4) Line 232 references Figure 3, but not a specific panel.

We referred Fig 3A, 3C, 3D, and 3E, in the text.

Figure 7/8 - can go to Supplemental.

We moved Fig 8 to supplemental. However, we think Fig 7 data is important and therefore we would like to present them as a main figure.

(1) There should be CycA expression in the control during the first 4 divisions.

Yes, there is CycA expression observed in the control during the first 4 divisions, while it’s much weaker than in sakuranull clone.

(2) Helpful to add the dotted lines to delineate (A) as well.

We added a dotted outline for germarium in Fig 7A.

(3) Line 263 CycA is miswritten as CyA.

We corrected the typo.

Figure 9

(1) Otu antibody control?

We validated Otu antibody in newly added Fig 10C and Fig S13A.

(2) Which Sakura-EGFP line was used? sakura het. or null background? This isn't mentioned in the text, nor legend.

We used Sakura-EGFP in the background of sakura[+/+]. We added this information in the methods and figure legend.

(3) C - Why the switch to S2 cells? Not able to use the Otu antibody in the IP of ovaries?

We can use the Otu antibody in the IP of ovaries. However, in anti-Sakura Western after anti-Otu IP, antibody light chain bands of the Otu antibodies overlap with the Sakura band. Therefore, we switched to S2 cells to avoid this issue by using an epitope tag.

Figure 10

(1) A- The resolution of images of the ribbon protein structure is poor.

We are very sorry for the low-resolution images. This was caused when the original PDF file with high-resolution images was compressed in order to meet the small file size limit in the eLife submission portal. In the revised submission, we used high-resolution images.

(2) A table summarizing the interactions between domains would help bring clarity to the data presented.

We added a table summarizing the fragment interaction results.

(3) Some images would be nice here to show that the truncations no longer colocalize.

We did not understand the reviewer’s points. In our study, even for the full-length proteins.

We have not shown any colocalization of Sakura and Otu in S2 cells or in ovaries, except that they both are enriched in developing oocytes in egg chambers.

Figure 12

(1) A - control and RNAi lines do not match.

We provided matched images.

(2) In general, since for Sakura, only its binding to Otu was identified and since they phenocopy each other, doesn't most of the characterization of Sakura just look at Otu phenotypes? Does Sakura knockdown affect Otu localization or expression level (and vice versa)?

We tested this by Western (Fig S15) and IF (Fig 12). Sakura knockdown did not decrease Otu protein level, and Otu knockdown did not decrease Sakura protein level (Fig S15). In sakuranull clone, Otu level was not notably affected (Fig 12). In sakuranull clone, Otu lost its localization to the posterior position within egg chambers.

Figure S6

(1) It is Luciferase, not Lucifarase.

We corrected the typo.

Reviewer #3 (Recommendations for the authors):

(1) It is interesting that germless and tumorous phenotypes coexist in the same population of flies. Additional consideration of these essentially opposite phenotypes would significantly strengthen the study. For example, do they co-exist within the same fly and are the tumorous ovarioles present in newly eclosed flies or do they develop with age? The data in Figure 8 show that bam knockdown partially suppresses the germless phenotype. What effect does it have on the tumorous phenotype? Is transposon expression involved in either phenotype? Do Sakura mutant germline stem cell clones overgrow relative to wild-type cells in the same ovariole? Does sakura RNAi driven by NGT-Gal4 only cause germless ovaries or does it also cause tumorous phenotypes? What happens if the knockdown of Sakura is restricted to adulthood with a Gal80ts? It may not be necessary to answer all of these questions, but more insight into how these two phenotypes can be caused by loss of sakura would be helpful.

We performed new experiments to answer these questions.

do they co-exist within the same fly and are the tumorous ovarioles present in newly eclosed flies or do they develop with age?

Tumorous and germless ovarioles coexist in the same fly (in the same ovary). Tumorous ovarioles are present in very young (0-1 day old) flies, including newly eclosed (Fig S5). The ratio of germless ovarioles increases and that of tumorous ovarioles decreases with age (Fig S5).

The data in Figure 8 show that bam knockdown partially suppresses the germless phenotype. What effect does it have on the tumorous phenotype?

bam knockdown effect on tumorous phenotype is shown in Fig S10. bam knockdown increased the ratio of tumorous ovarioles and the number of GSC-like cells.

Is transposon expression involved in either phenotype?

Since our transposon-piRNA reporter uses germline-specific nos promoter, it is expressed only in germ line cells, so we cannot examine in germless ovarioles.

Do Sakura mutant germline stem cell clones overgrow relative to wild-type cells in the same ovariole?

Yes, Sakura mutant GSC clones overgrow. Please compare Fig 6C and Fig S8.

Does sakura RNAi driven by NGT-Gal4 only cause germless ovaries or does it also cause tumorous phenotypes?

Fig S10 and Fig S12 show the ovariole phenotypes of sakura RNAi driven by NGT-Gal4. It causes both germless and tumorous phenotypes.

What happens if the knockdown of Sakura is restricted to adulthood with a Gal80ts?

Our mosaic clone was induced at the adult stage, so we already have data of adulthood-specific loss of function. Gal80ts does not work well with nos-Gal4.

(2) The idea that the excessive bam expression in tumorous ovaries is due to a failure of bam repression by dpp signaling is not well-supported by the data. Dpp signaling is activated in a very narrow region immediately adjacent to the niche but the images in Figure 7A show bam expression in cells that are very far away from the niche. Thus, it seems more likely to be due to a failure to turn bam expression off at the 16-cell stage than to a failure to keep it off in the niche region. To determine whether bam repression in the niche region is impaired, it would be important to examine cells adjacent to the niche directly at a higher magnification than is shown in Figure 7A.

We provided higher magnification images of cells adjacent to the niche in new Fig 7A.

We found that cells adjacent to the niche also express Bam-GFP.

That said, we agree with the reviewer. A failure to turn bam expression off at the 16-cell stage may be an additional or even a main cause of bam misexpression in sakura mutant. We added this in the Discussion.

(3) In addition, several minor comments should be addressed:

a. Does anti-Sakura work for immunofluorescence?

While our Sakura antibody detects Sakura in IF, it seems to detect some other proteins as well. Since we have Sakura-EGFP fly strain to examine Sakura expression and localization without such non-specific signal issues, we relied on Sakura-EGFP rather than anti-Sakura antibodies.

b. Please provide insets to show the phenotypes indicated by the different color stars in Figure 3C more clearly.

We provided new, higher-magnification images to show the phenotypes more clearly.

c. Please indicate the frequency of the expression patterns shown in Figure 4D (do all ovarioles in each genotype show those patterns or is there variable penetrance?).

We indicated the frequency.

d. An image showing TOskGal4 driving a fluorophore should be provided so that readers can see which cells express Gal4 with this driver combination.

It has been already done in the paper ElMaghraby et al, GENETICS, 2022, 220(1), iyab179, so we did not repeat the same experiment.

Version 2

Abstract

Abstract

During oogenesis, self-renewal and differentiation of germline stem cells (GSCs) must be tightly regulated. The Drosophila female germline serves as an excellent model for studying these regulatory mechanisms. Here, we report that a previously uncharacterized gene CG14545, which we named sakura, is essential for oogenesis and female fertility in Drosophila. Sakura is predominantly expressed in the ovaries, particularly in the germline cells, including GSCs. sakura null mutant female flies display rudimentary ovaries with germline-less and tumorous phenotypes, fail to produce eggs, and are completely sterile. The germline-specific depletion of sakura impairs Dpp/BMP signaling, leading to aberrant bag-of-marbles (bam) expression, resulting in faulty differentiation and loss of GSCs. sakura is also necessary for normal levels of piwi-interacting RNAs (piRNAs) levels and for female-specific splicing of sex-lethal (sxl), a master regulator of sex identity determination. We identified Ovarian Tumor (Otu) as a protein binding partner of Sakura and found that loss of otu phenocopies loss of sakura in ovaries. Thus, we identify Sakura as a crucial factor for GSC renewal and differentiation and oogenesis, and propose that Sakura and Otu function together in these processes.

Introduction

Oogenesis is a complex process in which germline stem cells (GSCs) develop into mature female gametes or oocytes. This process involves multiple layers of regulations and numerous genes, some of which remain to be discovered and characterized. Drosophila melanogaster, a genetically tractable model organism, has long been at the forefront of GSC and oogenesis research (Kirilly and Xie 2007).

Adult Drosophila females possess a pair of ovaries, each composed of 12-16 ovarioles. At the anterior tip of each ovariole is a structure called the germarium, which houses two to three GSCs (Fig 1A). GSCs reside in the most anterior region of the germarium, where they directly contact cap cells and escort cells, forming a niche crucial for GSC self-renewal (Lin and Spradling 1993; de Cuevas and Matunis 2011; Losick et al. 2011). GSCs undergo asymmetric mitotic division, producing two distinct cells: one GSC and one cystoblast. The cystoblast, destined to differentiate, undergoes four mitotic divisions with incomplete cytokinesis, resulting in interconnected cysts of 2, 4, 8, and 16 cells. Of the 16 cells, one will differentiate into an oocyte, while the remaining 15 develop into polyploid nurse cells, which synthesize proteins and RNAs for deposition into the oocyte.

Sakura expression pattern and mutant allele

(A) Schematic illustration of Drosophila ovary and germarium. Drosophila female has a pair of ovaries, each consisting of 12-16 ovarioles (cyan). The germarium (outlined with magenta) is located at the anterior tip of the ovarioles and consists of both germ cells and somatic cells. Germ cells include germline stem cells (GSCs), cystoblasts, cysts, and differentiating oocytes. Somatic cells include terminal filament (TF) cells, cap cells (CCc), escort cells (ECs), and follicle cells (FCs). GSCs and cystoblasts have spherical, unbranched spectrosomes, whereas cysts posesss branched fusomes. The distinct regions of the germarium—1, 2a, 2b, and 3—are indicated. (B) Drosophila Sakura protein (Sakura/CG14545) and its null mutant allele generated in this study (C) Western blot of dissected fly tissues. (D) Western blot of ovary lysates. (E) Confocal images of the germarium from the sakura-EGFP transgenic fly. Sakura-EGFP (green), Hts (red), and DAPI (blue). Scale bar: 10 μm. (F) Confocal images of the egg chambers from the sakura-EGFP transgenic fly. Sakura-EGFP (green), Vasa (red), and DAPI (blue). Sakura-EGFP is expressed in nurse cells and enriched in the developing oocyte (yellow arrow). Scale bar: 20 μm.

GSCs must be tightly regulated to maintain their undifferentiated state while dividing and differentiating to produce one GSC and one cystoblast. Failure in GSC self-renewal or inappropriate differentiation leads to stem cell loss, disrupting oocyte production (Lin 1997; Cox et al. 1998). Conversely, uncontrolled GSC division or defective differentiation results in an overabundance of GSC-like cells (a tumorous phenotype) and fewer differentiated cells, ultimately inhibiting oocyte production (Gateff et al. 1996; Ohlstein et al. 2000). Thus, a precise balance between GSC self-renewal and cyst differentiation within the germarium is essential for proper oogenesis and is tightly regulated.

Bone Morphogenetic Protein (BMP) signaling plays a critical role in regulating GSC self-renewal and differentiation during Drosophila oogenesis. In the stem cell niche, GSCs are physically anchored to cap cells via adherens and gap junctions (Song et al. 2002; Gilboa et al. 2003). Cap cells secrete BMP ligands, such as Decapentaplegic (Dpp) and Glass bottom boat (Gbb), which are recognized by transducing receptors like Thickvein (Tkv) and Saxophone (Sax) on the GSCs (Fig S1) (Xie and Spradling 1998; Casanueva and Ferguson 2004). Upon activation, these receptors phosphorylate a transcription factor Mad. The phosphorylated Mad (pMad) translocates into the nucleus, where it represses the transcription of bag-of-marbles (bam), a key differentiation factor (Kai and Spradling 2003; Song et al. 2004).

Bam, a ubiquitin-associated protein, is essential for cystoblast differentiation (McKearin and Spradling 1990; McKearin and Ohlstein 1995). Dpp/BMP signaling in the niche represses bam expression, preventing GSCs from differentiating and enabling self-renew (Song et al. 2004). However, as the daughter cells (cytoblasts) exit the niche, they derepress bam and initiate the differentiation program (Fig S1). Tight regulation of Dpp/BMP signaling and bam expression ensures that only cells within the niche adopt GSC fate, while adjacent daughter cells differentiate into cystoblasts. Loss of bam leads to blocked differentiation and the accumulation of GSC-like cells (tumorous phenotype), whereas ectopic bam expression forces GSCs differentiation, depleting the stem cell pool (McKearin and Spradling 1990; McKearin and Ohlstein 1995; Ohlstein and McKearin 1997).

One of the proteins that interacts with Bam is Ovarian Tumor (Otu) (Ji et al. 2017). Otu forms a deubiquitinase complex with Bam, which deubiquitinates Cyclin A (CycA), stabilizing CycA and promoting GSC differentiation (Ji et al. 2017). Otu also binds RNAs (Ji et al. 2019). Otu is crucial for oogenesis and female fertility. Mutations in the otu gene cause various ovarian defects, including tumorous growths, germ cell loss, abnormalities in oocyte determination and nurse cell dumping, and defects in the sexual identity determination of germ cells, indicating Otu’s importance in multiple stages of oogenesis (Smith and King 1966; Gans et al. 1975; Gollin and King 1981; King and Riley 1982; Storto and King 1988; Pauli et al. 1993; Rodesch et al. 1997; Glenn and Searles 2001). However, the precise molecular mechanisms by which Otu functions in these processes remain largely unknown.

In this study, we aimed to identify a novel gene required for oogenesis. Through a search for uncharacterized genes expressed exclusively in Drosophila ovaries, we discovered CG14545, a gene encoding a 114 amino acids (aa) long protein without any known functional domain (Fig 1B). We found that CG14545 is expressed specifically in female germline cells, including GSCs. Our genetic analysis revealed that CG14545 is required for oogenesis and female fertility and plays an important intrinsic role in regulating GSCs renewal and differentiation. Additionally, CG14545 is important for the piRNA pathway and for the female-specific mRNA splicing pattern of sex-lethal (sxl), a master regulator of sex determination. We identified Otu as a binding partner of the CG14545 protein, suggesting that both proteins may participate in the same molecular pathways to regulate GSC fate and differentiation. We named the CG14545 gene sakura, meaning “cherry blossom” in Japanese, symbolizing birth and renewal.

Results

Sakura is exclusively expressed in the ovaries

Based on high-throughput expression data in Flybase, sakura mRNA is highly and exclusively expressed in adult ovaries (Fig S2). To investigate Sakura protein expression, we generated a polyclonal anti-Sakura antibody against a recombinant full-length Sakura protein. Using this antibody, we examine Sakura protein expression at different stages of Drosophila development and across tissues. We found that Sakura protein is specifically expressed in ovaries, particularly in stage 1-11 egg chambers (Fig 1C).

sakura mutant flies

To explore the biological and molecular functions of Sakura in vivo, we generated a sakura mutant allele (sakuranull) with a one nucleotide (nt) insertion replacing 33 nts within the Sakura coding region using the CRISPR/Cas9 genome editing system (Fig 1B). This mutation causes a frameshift and a premature stop codon, resulting in a 33-amino acid (aa) N-terminal fragment of Sakura followed by a 22-aa segment produced by frameshifted translation (Fig 1B). This truncated fragment is unlikely to be functional, and we were unable to detect stable protein expression, as shown below (Fig 1D). Thus, we consider this allele a null mutation. sakuranull/null flies were viable, revealing that Sakura is not essential for survival.

To validate the sakuranull strain, we performed Western blots using ovary lysates from sakura mutant flies using the anti-Sakura antibody. As expected, full-length Sakura protein was detected in the ovaries of wild-type (sakura+/+) and heterozygous controls (sakuranull/+), but not in those of homozygous mutants (sakuranull/null) (Fig 1D). No smaller protein corresponding to the truncated Sakura fragment was detected in either sakuranull/+ or sakuranull/null ovaries thus, suggesting that the fragment is unstable or not expressed.

Sakura is cytoplasmic and expressed specifically in germ cells

To determine the localization of Sakura protein within the ovaries, we generated transgenic flies carrying a Sakura coding sequence fused with EGFP at the C-terminus (Sakura-EGFP) under the control of the sakura promoter. Western blot analysis of ovary lysates showed that Sakura-EGFP is expressed at levels comparable to endogenous Sakura (Fig 1D).

Oogenesis begins in the germarium, which contains 2-3 GSCs identified by round, unbranched spectrosomes that contact cap cells (Fig 1A) (Kirilly and Xie 2007). In contrast, cysts exhibit branched fusomes. Immunostaining with hu-li tai shao (HTS) antibody marks both spectrosomes and fusomes (Fig 1E), while Vasa is a known germ cell marker (Fig 1F). Confocal imaging revealed that Sakura-EGFP is specifically expressed in the cytoplasm of germ cells, including GSCs, cyst cells, nurse cells, and developing oocytes (Fig 1E and 1F). Sakura-EGFP was not expressed in somatic cells such as terminal filament, cap cells, escort cells, or follicle cells. (Fig 1E). In the egg chamber, Sakura-EGFP was detected in the cytoplasm of nurse cells and was enriched in developing oocytes (Fig 1F).

sakura is essential for female fertility

Given its ovary-specific expression at both mRNA and protein levels (Fig 1C and S2), we hypothesized that Sakura plays a critical role in ovarian function. Fertility assays revealed that sakuranull/null females laid no eggs when mated with wild-type males, while control females (sakura+/+ and sakuranull/+) laid numerous eggs (Fig 2A and 2B). Dissection of sakuranull/null ovaries revealed that they were rudimentary compared to normal ovaries in controls (Fig 2C). This suggests that the inability of sakuranull/null female to lay eggs is due to their underdeveloped ovaries. In contrast, sakuranull/null males were fertile and showed no significant differences compared to controls (sakura+/+ and sakuranull/+) (Fig S3).

sakura mutant flies are female-sterile and have rudimentary ovaries.

(A-B) Female fertility assays. (A) The number of eggs laid by test females crossed with OregonR wild-type males and (B) hatching rates of the eggs. Mean ± SD (n = 3). P-value < 0.001 (Student’s t-test, unpaired, two-tailed) is indicated by ***. (C) Stereomicroscope images of dissected whole ovaries. Scale bar: 200 mm.

To confirm that the observed female sterility was caused by the loss of sakura, we generated sakura rescue flies expressing Sakura-EGFP in the sakuranull/null background (sakuranull/null; sakura-EGFP). Western blotting confirmed that these flies expressed Sakura-EGFP, but not endogenous Sakura (Fig 1D). The sakura-EGFP transgene fully rescued both female fertility and ovary morphology in the sakuranull/null background (Fig 2). sakura-EGFP rescue females showed no significant fertility differences compared to controls (Fig 2A and 2B), and their ovaries were normal in shape and size (Fig 2C). We concluded that Sakura is essential for female fertility and normal ovary morphology but dispensable for male fertility, consistent with the observation that Sakura is predominantly expressed in ovarian germline cells (Fig 1).

sakuranull/null ovaries are germless and tumorous

Given the rudimentary appearance of sakuranull/null ovaries, we questioned whether they contained germ cells. We used flies carrying vasa-EGFP knocked-in reporter in the sakuranull/null background, as Vasa is a known germ cell marker. We observed that some sakuranull/null ovarioles were devoid of germ cells (“germless”, cyan stars), while others retained germ cells (Fig 3A and 3B). Immunostaining with HTS antibody revealed an excess number of GSC-like cells with round spectrosomes in sakuranull/null ovarioles containing germ cells (Fig 3C, orange stars), indicative of a ”tumorous” phenotype as previously described for mutants of other genes including bam, otu, and sxl (Smith and King 1966; Gans et al. 1975; Gollin and King 1981; King and Riley 1982; McKearin and Spradling 1990; Bopp et al. 1993; Eliazer et al. 2011; Jin et al. 2013; Yang et al. 2019). Additionally, we observed an excess number of cyst cells with branched fusomes that persisted throughout the ovarioles, suggesting abnormal cyst cell differentiation and division. Thus, loss of sakura results in both germ cell depletion and overgrowth.

sakuranull ovaries are germless and tumorous

(A, B) Confocal images of the ovaries from control (sakuranull/+) and sakuranull/null expressing Vasa-EGFP. Vasa-EGFP (green) and DAPI (blue). Yellow stars indicate the anterior tip of normal ovarioles, while orange and cyan stars indicate the anterior tips of tumorous and germless ovarioles, respectively, in Figure 1. Higher-magnification images of the germarium regions are shown in (B). Scale bars: 10 μm for (A) and 50 μm for (B). (C) Confocal images of control (sakuranull/+) and sakuranull/null ovaries stained with anti-Hts antibody to label spectrosomes and fusomes. Hts (red) and DAPI (blue). Yellow, orange, and cyan dotted lines mark the normal, tumorous, germless and germaria, respectively. Scale bars: 10 μm. (D) Ratio (%) of normal, germless, and tumorous ovarioles of indicated genotypes (ages 2-5 days; n=33, 25, 74, 29 respectively). (E) Quantification of GSC-like cell number per germarium in the indicated genotypes (ages 2-5 days; n = 33, 30, 35, and 31 respectively). (F) Confocal images of control (sakuranull/+) and sakuranull/null ovaries stained with anti-cleaved Caspase-3 antibody. Cleaved caspase-3 (red) and DAPI (blue). Scale bars: 20 μm.

Within the same ovary, some sakuranull/null ovarioles exhibited the tumorous phenotype, while others were germless, without any spectrosome or fusome staining (Fig 3A-3C, cyan stars). Quantification revealed that 35% of sakuranull/null ovarioles from 2-to 5-day-old flies were tumorous, while the remaining 65% were germless (n=74) (Fig 3D). In contrast, all ovarioles in control (sakura+/+ and sakuranull/+) and sakura-EGFP rescue flies were normal with no tumorous or germless phenotypes observed (Fig 3D). Approximately 95% of sakuranull/null ovarioles containing germ cells had an excess of GSC-like cells (>5) (Fig 3E and Fig S4). The mean number of GSCs or GSC-like cells in 2- to 5-day-old sakura+/+, sakuranull/+, and sakura-EGFP rescue was 2.2 ± 0.6, 2.2 ± 0.6, and 2.1 ± 0.7, respectively while that for sakuranull/null was 26.5 ± 17.5 (p-value <0.05).

We investigated the age-dependent effects on the tumorous and germless phenotypes by examining the ovarioles from 0-1 day old, 7-day old, and 14-day old flies. Ovarioles in control and sakura-EGFP rescue flies remained normal throughout this time course (Fig S5A). In the very young, 0-1 day old sakuranull/null, 78% of ovarioles were tumorous and 22% were germless. Overtime, the ratio of germless ovarioles increased while the ratio of tumorous ovarioles decreased. By 14 days old, only 5% of ovarioles were tumorous, with the remaining 95% being germless. The number of GSC-like cells in sakuranull/null ovarioles was already high in 0-1 day old flies, and it decreased over time (Fig S5B). Tumorous ovarioles, but not germless ones, exhibited significantly higher levels of cleaved Caspase-3 staining, indicative of apoptosis, compared to controls (Fig 3F). These results suggest that tumorous ovarioles undergo apoptosis, and that these ovarioles eventually become germless.

Together, these results suggest that Sakura is essential for regulating the survival, proliferation, and differentiation of germline cells, including GSCs.

Loss of sakura results in reduced germline piRNA

In controls, Vasa-EGFP is enriched in the perinuclear structure known as the nuage within nurse cells (Fig 3A and 3B), which is essential for piwi-interacting RNA (piRNA) biogenesis. In sakuranull/nulltumorous ovarioles, Vasa-EGFP retains its perinuclear localization, indicating that sakura is not required for the proper localization of Vasa to the nuage.

piRNAs are produced in germline cells and somatic follicle cells and transcriptionally and post-transcriptionally silence transposon expression through sequence-complementarity (Huang et al. 2017; Yamashiro and Siomi 2018). Disruption of the piRNA pathway can result in oogenesis arrest, germ cell loss, rudimentary ovaries, and sterility. Loss of piRNAs in germ cells leads to the upregulation (desilencing) of transposons, increased DNA damage, and apoptotic cell death31. We investigated whether the loss of sakura would result in loss of piRNA and transposon upregulation. We performed high-throughput sequencing (small RNA-seq and polyA+ RNA-seq) on ovary RNA samples from sakura+/+, sakuranull/+, sakuranull/null, and sakura-EGFP rescue (sakuranull/null with sakura-EGFP) flies.

piRNA levels in sakuranull/null ovaries were reduced compared to controls (sakura+/+), while sakuranull/+ and sakura-EGFP rescue showed no such reduction (Fig 4A). Consistent with this, transposon RNA levels in sakuranull/null ovaries were upregulated compared to sakura+/+, while sakuranull/+ and sakura-EGFP rescue showed no such upregulation (Fig 4B). Given that Sakura is specifically expressed in germline cells, but not in somatic cells (Fig 1E and 1F), we hypothesized that loss of sakura interrupts the piRNA pathway in germ cells, leading to the desilencing of germline transposons. A well-characterized germline transposon in Drosophila is Burdock (Donertas et al. 2013; Handler et al. 2013). We found that Burdock piRNA levels were significantly lower in sakuranull/null compared to controls and rescue flies (Fig 4C).

Loss of sakura results in lower piRNA levels and loss of piRNA-mediated transposon silencing

(A) Scatter plots of normalized number of transposon piRNA reads from small RNA-seq of indicated genotypes compared with sakura+/+. Means of three biological replicates are plotted. Downregulated (fold-change < 0.5) and upregulated (fold-change > 2) transposon piRNAs are shown in cyan and magenta, respectively. (B) Volcano plots of transposon RNAs from RNA-seq of indicated genotypes compared with sakura+/+. Three biological replicates per genotype were analyzed. Downregulated (adjusted p-value < 0.001 and log2(fold-change) < −2) and upregulated (adjusted p-value < 0.001 and log2(fold-change) > 2) transposons are shown in cyan and magenta, respectively. (C) Normalized number of reads for Burdock piRNAs from small RNA-seq. Mean ± SD (n = 3.) P-value < 0.001 (Student’s t-test, unpaired, two-tailed) is indicated by ***. (D) The Burdock sensor harbors a nanos promoter, a nuclear localization signal (NLS) appended to GFP and β-gal coding sequences, and a target sequence for Burdock piRNAs in the 3’UTR. Confocal images of ovaries from control (yRNAi) and sakuraRNAi flies harboring the Burdock sensor, where RNAi knockdown was specifically driven in the female germline with UAS-Dcr2 and NGT-Gal4. GFP (green), β-gal (red), and DAPI (blue). Scale bars: 50 μm. Three out of 3 tested control samples were negative for both GFP and β-gal, while 9 out of 9 tested sakuraRNAi samples were positive for both markers.

We employed the Burdock sensor, a transposon reporter tool to monitor germline piRNA activity (Handler et al. 2013). The reporter expresses nuclear GFP and β-gal under the control of the nanos promoter for germline expression, with a target sequence for Burdock piRNAs in the 3′ UTR (Fig 4D). Using a sakura RNAi line driven by UAS-Dcr2 and NGT-Gal4 to knock down sakura specifically in the germline, we observed that GFP and β-gal expression were highly elevated in sakura knockdown germlines compared to control RNAi (yRNAi) knockdowns (Fig 4D). This confirmed that loss of sakura leads to a loss of piRNA-mediated transposon silencing in the germline, suggesting that Sakura is essential for proper piRNA levels and piRNA-based transposon silencing in the germline.

Sex-specific sxl mRNA alternative splicing is dysregulated in sakuranull/null ovaries

Sxl is a master regulator of sex determination in Drosophila (Penalva and Sanchez 2003; Salz and Erickson 2010; Grmai et al. 2022). Sex-specific alternative splicing of sxl transcripts produces distinct mRNA isoforms: the female-specific and male-specific mRNA isoforms in respective sex and only the female-specific mRNA isoform encodes the functional Sxl protein, whereas the male-specific isoform does not. In the female germline, loss of Sxl function or disruption of female-specific sxl splicing leads to developmental defects, including germline tumors and sterility. Notably, several mutants with tumorous ovariole phenotypes—including otu mutants—exhibit aberrant sxl mRNA splicing, resulting in the expression of the male-specific isoform in ovaries and defects in germ cell sexual identity (Bopp et al. 1993; Pauli et al. 1993).

We examined sxl splicing in sakura mutants. As expected, ovaries from control and sakura-EGFP rescue flies expressed exclusively the female-specific sxl mRNA isoform and testes from control flies expressed only the male-specific isoform (Fig S6). In contrast, sakuranull/nullovaries exhibited the male-specific isoform and a reduced level of the female-specific isoform. These results indicate that female-specific sxl alternative splicing is disrupted in the absence of sakura.

sakura is important for oogenesis in germline cells beyond GSCs and germline cysts

Both sakuranull/null and sakura germline RNAi knockdown driven by UAS-Dcr-2 and NGT-Gal4, which initiates RNAi in germline from GSCs, resulted in rudimentary ovaries (Fig 2C and 5A). These ovaries lacked later-stage germline cells and egg chambers, making it difficult to assess the role of Sakura beyond the germarium. However, because Sakura-EGFP expression is not limited to GSCs and germline cysts (Fig 1E and 1F), we speculated that Sakura might function in later-stage germline cells as well.

sakura is important for oogenesis in later-stage egg chambers

(A) Stereomicroscope images of dissected whole ovaries. Scale bar: 500 μm. (B) Western blot of dissected Ovary lysates. sakuraRNAi #1 and sakuraRNAi #2 are two independent RNAi lines. (C) Number of eggs laid by sakura RNAi knockdown driven by UAS-Dcr2 and TOsk-Gal4. Mean ± SD (n = 5). P-value < 0.001 (Student’s t-test, unpaired, two-tailed) is indicated by ***. (D) Number of stage 14 oocytes produced by sakura RNAi knockdown flies driven by UAS-Dcr2 and TOsk-Gal4. P-value < 0.001 (Student’s t-test, unpaired, two-tailed) is indicated by ***. (E) Confocal images of the ovaries from sakura RNAi knockdown flies driven by UAS-Dcr2 and TOsk-Gal4, stained with anti-Orb and anti-Vasa antibodies. Orb (green), Vasa (red), and DAPI (blue). Yellow arrows label the normal enrichment of Orb in the developing oocytes. Magenta arrow labels mislocalized developing oocyte. White arrowheads label egg chambers with cytoskeletal disorganization. Scale bars: 100 μm. In A-E, y-RNAi was used as a control.

To explore this possibility, we used TOsk-Gal4 (a combination of osk-Gal4 and αTub67C-Gal4) to drive sakura RNAi knockdown in germline cells after the germline cyst stage (from germarium region 2b onward. Fig 1A), sparing GSCs and germline cysts (ElMaghraby et al. 2022). Unlike NGT-Gal4, TOsk-Gal4-driven sakura RNAi knockdown did not drastically affect ovary morphology (Fig 5A), allowing us to study the effects of sakura loss in egg chambers. We confirmed sakura RNAi (sakura RNAi #1 and #2) knockdown efficiency by Western blot, showing effective depletion of Sakura in ovaries from both NGT-Gal4 and TOsk-Gal4-driven RNAi lines (Fig 5B).

Interestingly, TOsk-Gal4-driven sakura RNAi knockdown severely reduced the numbers of eggs laid (Fig 5C) and stage 14 oocytes in ovaries compared to control RNAi (yRNAi) (Fig 5D), suggesting that sakura is important for oogenesis beyond the germline cyst stage. Additionally, we observed mislocalized and dispersed Oo18 RNA-binding protein (Orb), a marker for oocyte identity, in TOsk-Gal4-driven sakura RNAi ovaries beginning around stage 6 of oogenesis (Fig 5E). In controls, Orb was enriched and properly localized to the posterior end of stage ∼6-8 egg chambers (Fig 5E, yellow arrows). In sakuraRNAi ovaries, although Orb localization appeared normal at stage ∼6 (yellow arrow), mislocalization was evident by stage ∼8 (magenta arrow), and signs of cytoskeletal disorganization were apparent in later-stage egg chambers (white arrowheads). Phalloidin staining to visualize F-actin further supported cytoskeletal defects in sakuraRNAi later-stage egg chambers (Fig S7).

These defects likely contribute to the impaired production of stage 14 oocytes and eggs in sakuraRNAi flies (Fig 5C and 5D). We conclude that Sakura plays an essential role in oogenesis beyond the early germline stages. For all subsequent sakura RNAi experiments, we used the sakura RNAi #2 line.

Sakura is required intrinsically for GSC establishment, maintenance, and differentiation

To investigate whether sakura functions autonomously in the germline, we performed mosaic analysis using the FLP-FRT system with heat shock promoter-driven FLP (hs-flp) (Xu and Rubin 1993; Rubin and Huynh 2015). First, to assesse its role in GSC establishment, we induced sakuranull clones in primordial germ cells (PGCs) by applying heat shock before the early third instar larval stage and tracked their development into adult GSCs, following a previously established protocol (Yang et al. 2007). In control adult flies (FRT82B),12.3% (20/163) of GSCs were marked, wheras only 1.9% (4/213) were marked in sakuranullmutants (FRT82B, sakuranull) (p-value = 4.8 * 10-11, chi-square test), indicating that sakura is autonomously required for GSC establishment.

Next, we assessed GSC maintenance using the same system by generating sakuranull GSC clones in adult flies and measuring clone loss over time, as described previously (Xie and Spradling 1998; Yang et al. 2007). Four days after clone induction, 27.8% of GSCs were marked in controls (FRT82B) and 22.1% in sakuranull (FRT82B, sakuranull), which we defined as initial levels (Fig 6A). In controls, the percentage of marked GSCs declined to 25.8% on day 7 and and 16.8% on 14, resulting in a 39.4% loss rate over 10 days (from day 4 to day 14). In contrast, the proportion of marked sakuranull GSC clones dropped more sharply—to 16.3% on day 7 and and 5.4% on day 14—yielding a higher loss rate of 75.6% over 10 days. These results demonstrate that sakura is intrinsically important for GSC maintenance.

Germline clonal analysis of sakuranull

(A) Percentage of germaria with marked GSC clones indicated by the absence of GFP at 4, 7, and 14 days after clone induction at the adult stage. Arrows indicate the percent decrease of marked GSC clones compared to day 4. The genotype, actual percentage, and total number of germaria examined are shown in the adjacent table. P-value < 0.01 (Chi-squared test) is indicated by **. (B) Confocal images of Germ cell clones. sakuranull and control clones were marked by the absence of GFP. GFP (green), Hts (red), and DAPI (blue). Scale bars: 10 μm. (C) Number of marked (GFP-negative) GSC-like cells in germaria with marked GSCs of the indicated genotypes at 4, 7, and 14 days after clone induction. GSC-like cells containing round spectrosome were identified through immunostaining with anti-Hts antibody. P-value < 0.05 and <0.001 (Student’s t-test, unpaired, two-tailed) is indicated by * and ***, respectively.

We further tested whether sakura is intrinsically required for GSC differentiation, we quantified the number of marked (GFP-negative) and unmarked (GFP-positive) GSC-like cells (germline cells with a round spectrosome) in the germaria containing marked GSC clones at days 4, 7, and 14 post-induction (Fig 6B). We found that marked sakuranull GSC-like cells were significantly more numerous than marked control GSC-like cells at all time points (Fig 6C). The number of marked sakuranullGSC-like cells increased over the 10-day period, while the number of marked control cells did not (Fig 6C). In contrast, unmarked GSC-like cells in germaria containing sakuranullGSC clones did not differ significantly in number compared with those in germaria containing marked control GSC clones, and the number did not increase over time (Fig S8).

These results demonstrate that sakura is intrinsically important in germline cells, including GSCs, to regulate proper division and differentiation. In the absence of intrinsic sakura, germline cells become tumorous and undergo uncontrolled proliferation. Taken together, we conclude that sakura is required intrinsically for GSC establishment, maintenance, and differentiation.

Loss of sakura inhibits Dpp/BMP signaling

The Dpp/BMP signaling pathway plays a central role in regulating bam expression and GSC self-renewal and differentiation (Fig S1) (Kirilly and Xie 2007; Hayashi et al. 2020). We hypothesized that the germless and tumorous phenotypes observed in sakura loss-of-function ovaries could be due to Dpp/BMP signaling misregulation. To test this, we knocked down sakura in the germline driven by UAS-Dcr2 and NGT-Gal4 in the presence of bam-GFP reporter (Chen and McKearin 2003b). In controls (yRNAi), Bam-GFP expression was restricted to 8-cell cysts and disappeared in 16-cell cysts and onward (Fig 7A). In contrast, sakura knockdowns ovaries showed persistent Bam-GFP expression throughout the germarium including the GSC niche region (Fig 7A).

Loss of sakura inhibits Dpp/BMP signaling

(A) Confocal images of ovaries from control (yRNAi) and sakuraRNAi flies harboring the bam-GFP reporter, where RNAi knockdown was specifically driven in the female germline with UAS-Dcr2 and NGT-Gal4. Bam-GFP (green), Hts (red), and DAPI (blue). Germarium are outlined by dotted line. Scale bars: 20 μm. (B, C) Confocal images of germaria with germline clones of sakuranull stained with (B) anti-Bam antibody and (C) anti-pMad antibody. GFP (green), Bam or pMad (red), and DAPI (blue). Scale bars: 10 μm. (D) Mean pMad intensity in the germline clones of the indicated genotypes. Mean ± SD (n = 7). P-value < 0.001 (Student’s t-test, unpaired, two-tailed) is indicated by *** (E) Confocal images of germaria with germline clones of sakuranull stained with anti-CycA antibody. GFP (green), CycA (red), and DAPI (blue). Scale bars: 10 μm. (F) Mean CycA intensity in the germline clones of the indicated genotypes. Mean ± SD (n = 5 and 10 for FRT82B and FRT82B, sakuranull respectively). P-value < 0.01 (Student’s t-test, unpaired, two-tailed) is indicated by **. In B, C, and E, clones were marked with the absence of GFP.

To further confirm this finding, we used the FLP-FRT system to induce sakuranull clones in germaria and performed immunostaining with anti-Bam antibody. In control clones (FRT82B), Bam expression was observed exclusively in 8-cell cysts. However, in all sakuranull clones (FRT82B, sakuranull), Bam was aberrantly expressed throughout the germarium, including in GSCs (Fig 7B). These results suggest that in the absence of sakura, Bam expression is no longer repressed by Dpp/BMP signaling in the GSC niche, leading to GSC loss, and is no longer shut off after the 16-cell cyst stage.

In GSCs, pMad translocates into the nucleus to repress Bam expression (Fig S1) (Kai and Spradling 2003; Song et al. 2004). We stained ovaries with anti-pMad antibodies and found that pMad intensity was significantly reduced in sakuranull GSC clones compared to control GSCs (Fig 7C and D). This indicates that the dysregulated Bam expression observed in sakuranull was due to reduced pMad levels in GSCs.

Previous studies have shown that ectopic expression of a stable form of CycA leads to germ cell loss (Chen et al. 2009). This germ cell loss phenotype is also observed upon ectopic bam expression in GSCs (Xie and Spradling 1998; Chen and McKearin 2003a; Xia et al. 2010). It was reported that Bam associates with Ovarian tumor (Otu) to promote deubiquitination and stabilization of CycA (Ji et al. 2017). Since we observed derepressed bam expression in sakuranullcells, we investigated whether CycA levels were affected. Staining ovaries with anti-CycA antibodies revealed higher CycA intensity in sakuranull clones compared to neighboring wild-type cells in the same germarium (FRT82B, sakuranull) and control clones in control germarium (FRT82B) (Fig 7E). sakuranull clones exhibited significantly higher mean CycA intensity compared to control clones (Fig 7F). The elevated CycA levels in sakuranull clones likely results from Bam misexpression and Bam-mediated stabilization.

To determine whether bam misexpression occurs in egg chambers in sakura knockdown conditions, we examined bam expression in TOsk-Gal4-driven sakura RNAi flies. Bam was appropriately restricted to 8-cell cysts in germaria and not misexpressed in egg chambers compared to yRNAi controls (Fig. S9, cyan arrowheads). Thus, sakura is specifically required in GSCs and cyst cells to regulate bam expression.

We also examined DE-Cadherin (DE-Cad), which plays a role in oocyte positioning within the egg chamber (Godt and Tepass 1998). DE-Cad level and localization appeared normal in both control and sakura RNAi ovaries through approximately stage 8. However, later-stage egg chambers in sakura RNAi flies displayed structural disorganization (Fig. S9, white arrowheads), consistent with defects observed in Fig. 5E and Fig. S7.

Attempts to rescue sakura loss-of-function ovariole phenotypes

To determine whether the phenotypes caused by sakura loss-of-function could be rescued, we first performed genetic interaction experiments by knocking down either bam or cycA in the germline. As expected, germline-specific bam RNAi driven by UAS-Dcr2 and NGT-Gal4 resulted in 100% tumorous ovarioles, while cycA RNAi produced 100% normal ovarioles in 2–5-day-old flies (Fig. S10A). Double knockdown of sakura and bam reduced the proportion of germless ovarioles and increased the proportion of tumorous ovarioles compared to sakura RNAi alone or the double knockdown of sakura and control w RNAi. However, no normal ovarioles were observed (Fig. S10A). sakura and cycA double knockdown had no effect on the phenotype distribution, suggesting that the germless phenotype in sakura mutants is partially attributable to ectopic bam expression but not to cycA misregulation.

The number of GSC-like cells was increased in sakura and bam double knockdowns compared to sakura RNAi or sakura and w double RNAi, whereas sakura and cycA knockdown did not alter GSC-like cell numbers (Fig. S10B),, further supporting a genetic interaction between sakura and bam.

Next, we tested whether the sakura loss-of-function phenotypes could be rescued by overexpressing components of the Dpp/BMP signaling pathway. Germline expression of UASp-Mad-GFP driven by nos-Gal4-VP16 in both sakuranull/+ controls and sakuranull/null did not alter the proportions of ovariole phenotypes (Fig S11A) or the number of GSC-like cells (Fig S11B). Thus, transgenic overexpression of Mad did not rescue the sakura mutant phenotypes.

We then tested whether overexpression of a constitutively active form of the Dpp receptor Thickveins (Tkv.Q253D) (Casanueva and Ferguson 2004) could rescue the phenotypes. Expression of UASp-tkv.Q253D in the germline using a NGT-Gal4 in a control (yRNAi) background resulted in 100% tumorous ovarioles, as expected (Fig. S12A). When co-expressed with sakura RNAi, Tkv.Q253D increased the proportion of tumorous ovarioles and decreased the proportion of germless ovarioles relative to sakuraRNAi alone but did not restore normal ovarioles. The number of GSC-like cells was also increased by Tkv.Q253D expression in sakuraRNAi (Fig S12B), indicating partial phenotypic modulation via BMP pathway activation.

Sakura binds Otu

To explore the molecular function of Sakura in oogenesis, we aimed to identify Sakura-interacting proteins. We conducted a co-immunoprecipitation experiment using anti-GFP magnetic beads on the ovary lysates of flies expressing Sakura-EGFP, followed by mass spectrometry to identify co-immunoprecipitated proteins. Ovary lysates from flies without the sakura-EGFP transgene (w1118) served as negative controls. Mass spectrometric analysis revealed several proteins with peptide signals present in all three biological replicates of the Sakura-EGFP samples without any signals in any of the three biological replicates of the negative control (Table 1). Among these, the most promising candidate, exhibiting the highest unique peptide signal in each of three replicates of the Sakura-EGFP samples was Otu.

To confirm this finding, we generated a polyclonal anti-Otu antibody and validated it by Western blots using ovary lysates from the hypomorphic otu14 mutant (Fig S13A) (King et al. 1986; Storto and King 1987; Steinhauer and Kalfayan 1992a; Pauli et al. 1993). Using this antibody, we detected endogenous Otu protein in the Sakura-EGFP immunoprecipitants from ovaries of 3–7-day-old flies, but not in the negative control (Fig 8A), validating the mass spectrometry results. Next, using wild-type (without Sakura-EGFP) ovary lysates from 3–7-day old-flies, we immunoprecipitated endogenous Sakura with anti-Sakura antibody and detected endogenous Otu in the immunoprecipitant by Western blot (Fig 8B), demonstrating that endogenous Sakura interacts with endogenous Otu. Furthermore, we found that ectopically expressed Flag-tagged Otu co-immunoprecipitated with ectopically expressed HA-tagged Sakura in S2 cells (Fig 8C).

Sakura interacts with Otu.

(A) Co-immunoprecipitation using anti-GFP magnetic beads followed by Western blotting. Ovary lysates expressing Sakura-EGFP in sakura+/+ background and those from w1118 negative control were tested. (B) Co-immunoprecipitation using beads bound with rabbit anti-Sakura followed by Western blotting. Ovary lysates from w1118 flies were used. Rabbit IgG was used as control IP. (C) Co-immunoprecipitation using beads bound with anti-FLAG antibody followed by Western blotting. S2 cell lysates expressing 3xFLAG-Otu and Sakura-mCherry-3xHA or mCherry-3xHA (negative control) were used.

The otu gene encodes two annotated protein isoforms, a predominant 98 kDa isoform and a less abundant 104 kDa isoform generated by alternative splicing (Steinhauer and Kalfayan 1992b). The 98 kDa isoform lacks the Tudor domain, which is encoded by an alternatively spliced 126-nucleotide exon (Fig 9B. Otu(ι1Tudor)) (Van Buskirk and Schupbach 2002). Previous studies have shown that the 104 kDa Otu isoform is more abundant in predifferentiated germline cells and is sufficient to carry all known Otu functions, whereas the 98 kDa Otu isoform becomes prominent during later stages of oogenesis and is implicated in nurse cell regression and oocyte maturation (Sass et al. 1995). Consistent with these findings, we observed that ovaries from very young (2–5 hour old) flies—containing only germaria and previtellogenic egg chambers—predominantly expressed the 104 kDa isoform (Fig S13B). In contrast, ovaries from 3–7 day old flies—containing all 14 stages of oogenesis—predominantly expressed the 98 kDa isoform. Importantly, both the 104 kDa and 98 kDa isoforms co-immunoprecipitated with endogenous Sakura using anti-Sakura antibody, demonstrating that Sakura interacts with both isoforms of Otu and that the Tudor domain is not required for this interaction (Fig. S13B).

N-terminal regions of Sakura and Otu are important for interaction.

(A) Predicted structure of the Sakura and Otu protein complex made by AlphaFold. Full-length Sakura and N-terminal Otu fragment (N1. 1-405aa) were used for prediction. (B) Full-length Otu and Otu fragments tested in co-immunoprecipitation assays. The binding assay results from (D) is summarized. (C) Full-length Sakura and Sakura fragments tested in co-immunoprecipitation assays (N: N-terminal, M: middle, C: C-terminal). The binding assay results from (E) is summarized. (D, E, F) Co-immunoprecipitation using anti-HA magnetic beads followed by Western blotting. S2 cell lysates expressing HA-tagged mCherry were used as negative controls.

Next, we sought to identify which regions of Sakura and Otu are important for their interaction. We predicted the structure of the Sakura-Otu protein complex using AlphaFold (Jumper et al. 2021; Yang et al. 2023). AlphaFold suggested that the N-terminal region of Sakura (1-49aa) and the N-terminal region of Otu (1-405aa) are highly likely to form interactions, while the C-terminal region of Sakura (100-144aa) does not interact with Otu (Fig 9A-C). The Tudor domain of Otu is not directly involved in this interaction, consistent with our findings (Fig S13B). To experimentally validate the predictions made by AlphaFold, we generated a series of truncated Otu fragments, including Otu(ΔTudor) (1-336, 389-853aa, corresponding to the endogenous 98 kDa isoform), Otu(N1) (1-405aa), Otu(N2) (1-339aa), and Otu(C) (405-853aa) (Fig 9B). We then performed co-immunoprecipitation assays to test which fragments could interact with full-length Sakura. All fragments except Otu(C) were associated with full-length Sakura protein in S2 cells (Fig 9D), demonstrating that the C-terminal low-complexity region of Otu is dispensable for the interaction. This assay also showed that the Tudor domain is not critical and the N-terminal fragment (1-339aa) of Otu is sufficient for interaction with Sakura. These findings align well with the predictions made by AlphaFold.

Subsequently, we created several truncated Sakura fragments and conducted co-immunoprecipitation assays in S2 cells to determine which Sakura fragments can interact with full-length Otu. The fragments generated included Sakura(NM) (1-100aa), Sakura(NC) (1-49, 100-144aa), Sakura(MC) (49-144aa), Sakura(N) (1-49aa), Sakura(M) (49-100aa), and Sakura(C) (100-144aa) (Fig 9C). We found that Sakura(NM), Sakura(NC), Sakura(N), and Sakura(M) were associated with Otu (Fig 9E). Compared to Sakura(NM) and Sakura(N), the interaction of Sakura(M) with Otu is relatively weaker, as indicated by a fainter band in the Western blot (Fig 9E). Sakura-NC exhibited a weak interaction with Otu, while Sakura(MC) and Sakura(C) showed no interaction (Fig 9E). These assays demonstrated that the N-terminal region (1-49aa) of Sakura is sufficient for interaction with Otu. Moreover, the results suggest that the C-terminal region of Sakura (100-144aa) does not interact with Otu and that adjoining it with the N-terminal (1-49aa) or the middle (M) region (49-100aa) weakens the interaction with Otu. These findings support the AlphaFold predictions.

Finally, to test whether the interaction between Sakura and Otu can be achieved solely through their N-terminal regions, we performed co-immunoprecipitation assay with Sakura(N) and Otu(N2). We discovered that Sakura(N) and Otu(N2) physically interact (Fig 9F). A reciprocal co-immunoprecipitation assay also confirmed their interaction (Fig S14), revealing that the interaction between Sakura and Otu can be established with just the 1-49aa of Sakura and the 1-339aa of Otu. Therefore, we conclude that Sakura interacts with Otu in vivo, and the N-terminal regions of both proteins are sufficient for their interaction.

Loss of otu phenocopies loss of sakura

Having established the interaction between Sakura and Otu, we were interested in whether they function together during oogenesis. We speculated that for them to interact and function together, Sakura and Otu must exhibit similar expression and localization patterns in ovaries. We generated two transgenic flies carrying full-length otu-EGFP or otu(ΔTudor)-EGFP transgenes; both constructs are C-terminally fused with EGFP and under the control of the otu promoter. We found that both Otu-EGFP and Otu(ΔTudor)-EGFP display similar expression and localization patterns to Sakura-EGFP (Fig 1E, 1F and Fig 10A). Like Sakura-EGFP, Otu-EGFP and Otu(ΔTudor)-EGFP are specifically expressed in germ cells, but not in follicle cells, and are localized to the cytoplasm, being enriched in the developing oocytes (Fig 10A). We did not observe any difference in localization patterns between Otu-EGFP and Otu(ΔTudor)-EGFP; suggesting that the Tudor domain is not crucial for Otu protein localization. Notably, the localization pattern exhibited by the transgenic Otu-EGFP generated in this study is consistent with a previous report (Glenn and Searles 2001).

Loss of otu phenocopies loss of sakura

(A) Confocal images of ovaries from otu-EGFP and otu(ΔTudor)-EGFP transgenic flies. Otu-EGFP and Otu(ΔTudor)-EGFP (green), DAPI (blue). Yellow arrows show the enrichment of Otu-EGFP and Otu(ΔTudor)-EGFP in developing oocytes. Scale bars: 20 μm. (B) Stereomicroscope images of dissected whole ovaries from wRNAi (control) and otuRNAi flies where RNAi knockdown was driven in the female germline with nos-Gal4-VP16 or TOsk-Gal4. Scale bar: 500 μm. (C) Western blot of dissected ovary lysates. (D) Number of eggs laid by TOsk-Gal4 > wRNAi and TOsk-Gal4 > otuRNAi flies. Mean ± SD (n = 5). P-value < 0.001 (Student’s t-test, unpaired, two-tailed) are indicated by ***. (E) Violin plots of the number of stage 14 oocytes produced in TOsk-Gal4 > wRNAi and TOsk-Gal4 > otuRNAi flies. n = 30. P-value < 0.001 (Student’s t-test, unpa©red, two-tailed) is indicated by ***.

Previous studies have shown that mutations in otu lead to defects in germ cell division and differentiation, resulting in phenotypes including tumorous ovarioles (King and Riley 1982; King et al. 1986; Storto and King 1988; Steinhauer and Kalfayan 1992b). To study the role of otu specifically in the germline, we performed RNAi knockdown of otu in the germline driven by nos-Gal4-VP16 (Fig 10B and 10C). Interestingly, we found that germline depletion of otu results in rudimentary ovaries, similar to the loss of sakura (Fig 2C, Fig 5A, and Fig 10B). Furthermore, similar to sakura RNAi (Fig 5), TOsk-Gal4-driven otu RNAi knockdown did not affect ovary morphology (Fig 10B), but it resulted in a reduced number of eggs laid and stage 14 oocytes compared with control RNAi (wRNAi) (Figure 10D and 10E).

Given that sakura loss of function causes tumorous and germless ovarioles, apoptosis, piRNA pathway defects, bam misexpression, and reduced pMad levels (Fig 3, 4 and 7A-D), we asked whether similar defects occur upon otu depletion. Germline otu RNAi knockdown in 2-5 days old flies using UAS-Dcr2 and NGT-Gal4 lead to both germless (∼70%) and tumorous (∼30%) ovarioles, closely resembling the phenotype of sakura RNAi (Fig S10A). Similar germless and tumorous phenotypes were also observed in nos-Gal4-VP16 > otuRNAi flies (Fig 11A). In these otuRNAiovarioles, Bam-GFP expression was no longer restricted to 8-cell cysts, instead persisting throughout the germarium and egg chambers (Fig 11A). Additionally, otu14/14 mutant GSCs showed reduced pMad levels (Fig 11B), and cleaved Caspase-3 staining revealed elevated apoptosis in otu RNAi ovaries (Fig 11C). Germline depletion of otu resulted in elevated expression of GFP and β-Gal produced by the Burdock sensor, indicating loss of piRNA-mediated transposon silencing (Fig 11D).

Loss of otu results in low levels of pMad and derepression of bam in the germaria

(A) Confocal images of the ovaries from control (wRNAi) and otuRNAi flies harboring the bam-GFP reporter, where RNAi knockdown was driven in the female germline with nos-Gal4-VP16. Bam-GFP (green), Hts (red), and DAPI (blue). Scale bars: 20 μm. (B) Confocal images of germaria from control (otu14/+) and otu14/14 mutant flies stained with anti-pMad (green) and anti-Hts (red) antibodies. Scale bars: 10 μm. (C) Confocal images of the ovaries from nos-Gal4-VP16 > wRNAiand nos-Gal4-VP16 > otuRNAi flies stained with anti-cleaved Caspase 3 antibody. Cleaved Caspase-3 (red) and DAPI (blue). Scale bars: 20 μm. (D) Confocal images of the ovaries from wRNAi and otuRNAi flies carrying the Burdock sensor, with RNAi driven in the female germline using UAS-Dcr2, NGT-Gal4 and nos-Gal4-VP16. GFP (green), β-gal (red), and DAPI (blue). Scale bars: 50 μm. Three out of 3 tested control samples were negative for both GFP and β-gal, while 14 out of 14 tested otuRNAi samples were positive for positive for GFP and β-gal.

Together, these data demonstrate that otu loss-of-function closely phenocopies sakura loss-of-function, including defects in germline maintenance and differentiation, bam misexpression, reduced pMad signaling, and compromised piRNA pathway activity. These findings, along with physical interaction between Sakura and Otu (Figs 8, 9, S13, and S14), suggest that they function together in regulating germline maintenance and differentiation.

To further explore their genetic interaction, we compared the ovariole phenotypes caused by single and double knockdowns of sakura and otu in the germline driven by UAS-Dcr2 and NGT-Gal4. Individual knockdown of either gene resulted in ∼70% germless and ∼30% tumorous ovarioles in 2-5 days old flies (Fig S10A). Simultaneous knockdown of both genes increased the propotion of germless ovarioles to 86%, with a corresponding decrease in tumorous phenotypes to 14%, suggesting a synergistic enhancement of the germless phenotype. The number of GSC-like cells were not affected by the double knockdown of sakura and otu compared with their respective single knockdown (Fig S10B).

Sakura and Otu proteins do not depend on each other for their abundance

We investigated whether Sakura and Otu proteins depend on each other for their abundance. We performed Western blot analysis on ovaries with germline-specific knockdown of sakura or otu, driven by TOsk-Gal4. Otu protein levels were not reduced in sakuraRNAi ovaries compared to control yRNAi (Fig S15A), and conversely, Sakura levels were not decreased in otuRNAi ovaries compared with control wRNAi ovaries (Fig S15B). These results indicate that Sakura and Otu do not require each other for their protein expression or stability.

To further test this conclusion and to examine whether Otu enrichment to the posterior of egg chambers is affected by loss of sakura, we generated marked sakuranull germline clones using the FLP-FRT system. Clones were marked by the absence of RFP, and experiments were performed in the flies expressing the otu-EGFP or otu(ΔTudor)-EGFP transgene in an otherwise wild-type (otu+/+) background. Otu-EGFP fluorescence signal was comparable between sakuranullclones (FRT82B, sakuranull) and control clones (FRT82B) (Fig 12A). Similarly, Otu(ΔTudor)-EGFP levels were unaffected in sakuranull clones compared to the control clones (Fig 12B), confirming that Sakura is not required for Otu protein expression or stability.

Otu enrichment to the posterior within egg chambers is lost in sakuranull

Confocal images of egg chambers with germline clones of sakuranull expressing (A) Otu-EGFP or (B) Otu(i1Tudor)-EGFP. Fly genotypes used: hs-flp/w; otu-EGFP/+; FRT82B, ubi-RFP/FRT82B. hs-flp/w; otu-EGFP/+; FRT82B, ubi-RFP/FRT82B, sakuranull. hs-flp/w; otu(i1Tudor)-EGFP/+; FRT82B, ubi-RFP/FRT82B. hs-flp/w; otu(i1Tudor)-EGFP /+; FRT82B, ubi-RFP/FRT82B, sakuranull. RFP (red), Otu-EGFP or Otu(i1Tudor)-EGFP (green), and DAPI (blue). Scale bars: 20 μm. Marked clones (RFP-negative) are outlined with white dotted lines. Yellow arrows indicate normal posterior enrichment of Otu-EGFP and Otu(i1Tudor)-EGFP signal; magenta arrows indicate the loss of this posterior enrichment in sakuranullclones.

Both Otu-EGFP and Otu(ΔTudor)-EGFP exhibited clear enrichment at the posterior of egg chambers in the control clones (FRT82B) —consistent with localization to the developing oocyte (Fig 12A and 12B). This posterior enrichment was absent in sakuranullclones, demonstrating that Sakura is crucial for this process.

Sakura does not affect Otu’s deubiquitinase activity in vitro

A previous study has shown that Otu possesses deubiquitinase activity (Ji et al. 2017). We asked whether Sakura regulates the Otu’s deubiquitinase activity. To test this, we performed an in vitro deubiquitination assay using Ub-Rhodamine 110 as a model substrate. Consistent with the previous study (Ji et al. 2017), we detected a deubiquitinase activity in Otu that was ectopically expressed and purified from S2 cells (Fig S16). The presence of Sakura, purified from E.coli, did not affect Otu’s deubiquitinase activity under these assay conditions. Additionally, Sakura itself did not exhibit deubiquitinase activity.

Discussion

In this study, we identified Sakura, encoded by a previously uncharacterized gene CG14545, as an essential factor for oogenesis and female fertility. We demonstrated that Sakura is specifically expressed in germline cells in the ovary, including GSCs, is localized to the cytoplasm, and is enriched in the developing oocytes. sakura homozygous null mutant flies are viable but completely female-sterile and male-fertile. Loss of sakura, either through null mutation or germline RNAi including in GSCs, results in rudimentary ovaries exhibiting germless and tumorous phenotypes.

The tumorous phenotype associated with the loss of sakura is characterized by an excess of GSC-like cells, which feature round spectrosomes, as well as cyst cells with branched fusomes (Fig 3C). The increased number of GSC-like cells in loss of sakura suggests dysregulation of GSC self-renewal and differentiation. Meanwhile, the presence of excess cyst cells with branched fusomes indicates abnormal differentiation and division of cysts. Differentiation of cystoblasts typically involves four mitotic divisions with incomplete cytokinesis, leading to 16 cyst cells interconnected by branched fusomes. The degree of fusome branching serves as a marker for the stages of cyst cell division, with increased branching from 2-cell to 16-cell cysts (de Cuevas and Spradling 1998). Notably, fusomes begin to degenerate and disappear after 16-cell cysts as the germline cyst enters the meiotic zone or region 2 of the germarium. The persistence of cyst cells with branched fusomes in sakuranull/nullsuggests that sakura is crucial for the proper division and differentiation of cysts, and we speculate that it is required for germline cysts to enter meiotic division.

Mosaic analysis of sakuranull indicates that sakura is intrinsically required for the establishment of GSCs in the ovary. When mutant clones were induced in PGC stage, significantly fewer sakuranull marked GSCs were observed in adult ovaries, suggesting that many sakuranull PGCs fail to survive or differentiate into GSCs. Furthermore, induction of sakuranull clones in adult ovaries led to a more rapid decline in marked sakuranull GSCs compared to controls, indicating that sakura is also intrinsically required for GSC maintenance (Fig 6A). Over time, germaria containing sakuranull GSC clones became increasingly tumorous, with an expanding population of sakuranull GSC-like cells (Fig 6B and 6C). In contrast, the number of GSC-like cells across entire and all ovarioles regardless of whether GSCs in the niche remain, declined over time in sakuranull ovaries (Fig S5B). These results suggest that sakuranull GSCs initially undergo aberrant proliferation, leading to tumor formation, but are ultimately sakuranullGSCs and GSC-like cells are lost through a cell death mechanism, as evidenced by the elevated levels of cleaved Caspase-3 in sakuranull ovaries (Fig 3F).

Dpp/BMP signaling governs GSC self-renewal and cystoblast differentiation by repressing bam in GSCs and de-repressing it in daughter cystoblasts (Fig S1). This process is mediated by the transcription factor Mad, which, when phosphorylated (pMad), translocates into the nucleus to repress bam transcription (Kirilly and Xie 2007; Kahney et al. 2019; Hinnant et al. 2020). Our findings indicate that Bam expression is not limited to 8-cell cysts but continues throughout the germarium in ovaries lacking sakura function (Fig 7A and 7B). The misexpression of Bam in GSCs likely results from reduced levels of pMad in the GSCs (Fig 7C, and 7D). Given the low pMad levels observed upon loss of sakura, we propose that bam misregulation in GSCs occurs primarily at the transcriptional level, although Bam is also known to be subject to post-transcriptional regulation (Pek et al. 2009). The continued expression of Bam throughout the germarium, well beyond the GSC niche, suggests a failure to terminate bam expression at the 16-cell cyst stage —a developmental transition when bam is normally turned off. Thus, sakura mutants likely exhibit two distinct molecular defects contributing bam overexpression: (1) impaired Dpp/BMP signaling within the GSC niche, and (2) a failure to downregulate bam at 16-cell stage. Given the limited space within the GSC niche, the latter defect may be the predominant contributor to the observed bam overexpression. The molecular mechanism that silences bam at the 16-cell cyst stage remains poorly understood, and how sakura loss perturbs this regulation is unclear. The sakura mutant may therefore serve as a valuable model to investigate the mechanism that normally suppresses bam expression at the 16-cell stage.

Transposons are mobile genetic elements that, if not silenced, can generate DNA damage and genomic instability (Levin and Moran 2011). piRNAs derived from transposons and other repeats can target and silence transposon RNAs to preserve genome integrity in germ cells (Siomi et al. 2011). Loss of piRNAs leads to transposon derepression, resulting in increased DNA damage, which subsequently triggers cell death (Kang et al. 2018; Moon et al. 2018). Genetic damage in germ cells can cause developmental defects and diseases that may be inherited by the next generation. Thus, elimination of defective germ cells is crucial for maintaining germline integrity of a species (Chu et al. 2014; Ota and Kobayashi 2020). We found that loss of sakura results in reduced piRNA levels and loss of piRNA-mediated transposon silencing in the germline (Fig 4). The observed apoptosis, indicated by elevated cleaved Caspase-3 levels in sakura mutant ovaries (Fig 3F), suggest that desilencing of transposon due to reduced piRNA levels likely results in increased DNA damage, triggering cell death. We speculate that the germless phenotype in sakura mutants may partly arise from an apoptotic germline elimination program activated to maintain germline integrity.

In addition to its expression in GSCs and cysts in the germarium, Sakura is also expressed in germline cells in later-stage egg chambers (Fig 1E-F) and is required at this stage for proper oogenesis. When Sakura is depleted from region 2b of the germarium onward —leaving GSCs and early cyst cells intact—females have significantly fewer stage 14 oocytes and lay significantly fewer eggs (Fig 5). The failure to produce stage 14 mature oocytes likely stems from cytoskeletal disorganization that disrupt oocyte development, as evidenced by mislocalization of Orb (Fig 5E, S7, and S12). Notably, piRNA pathway mutants also exhibit similar Orb mislocalization (Ohtani et al. 2013). Therefore, the reduced piRNA levels (Fig 4) may contribute to the oocyte development defects, including Orb mislocalization, in sakura mutants. (Fig 5E).

Sakura does not possess any known protein domains. To infer its function, we identified Otu, a protein known to be crucial for oogenesis, as a protein partner of Sakura. Mutations in otu gene lead to a range of ovarian phenotypes, including germ cell loss, tumorous egg chambers filled with undifferentiated germ cells, defects in germlne sexual identity determination, abnormalities in nurse cell chromosome structure, and defects in oocyte determination (King and Riley 1982; Storto and King 1988; Pauli et al. 1993; Glenn and Searles 2001). We showed that germline depletion of Otu via RNAi phenocopies loss of sakura. Similar to Sakura, loss of otu inhibits Dpp/BMP signaling, resulting in low pMad levels in GSCs and Bam de-repression (Fig 11A-B). Additionally, loss of otu also results in the loss of piRNA-mediated silencing, paralleling the effects of sakura loss. Furthermore, germline knockdown of otu yielded a similar ratio of germless ovaries as seen in sakura-RNAi, and simultaneous knockdown of both otu and sakura exacerbated the germless ovary phenotype (Fig S10). These observations raise the possibility that Sakura and Otu function together to regulate germ cell maintenance in the ovaries and are involved in Dpp/BMP signaling to balance stem cell renewal and differentiation.

Otu possesses deubiquitinase activity, catalyzed by its N-terminal Otu domain (Ji et al. 2017; Ji et al. 2019). In germ cells, Otu interacts with Bam, forming a deubiquitinase complex that deubiquitinates and thereby stabilizes CycA, promoting GSC differentiation. The predicted structure of Sakura and Otu complex suggests that the Mid domain of Sakura directly contacts the Otu domain of Otu (Fig 9A). We found that Sakura lacks a deubiquitinase activity and it does not directly affect Otu’s deubiquitinase rate in our in vitro assays using Ub-Rhodamine 110 as a substrate (Fig S16). This does not preclude the possibility that Sakura may still influence Otu’s deubiquitinase activity, potentially guiding Otu to its substrate and determining substrate specificity. Our in vitro assays may not have detected such specificity changes.

Otu is an RNA-binding protein whose deubiquitinase activity is enhanced by RNA binding (Ji et al. 2019). Bam, along with other proteins such as Bgcn, Mei-P26, and Sxl, binds —a key stem cell maintenance factor (Wang and Lin 2004)— and represses its translation in GSCs (Wang and Lin 2004; Li et al. 2009; Chau et al. 2012; Li et al. 2013). Bam and Otu form a protein complex (Ji et al. 2017). Future studies should explore whether Sakura modulates Otu’s RNA-binding properties and its interaction with other proteins. Although Sakura is exclusively expressed in ovaries, particularly in germline cells including GSCs (Fig 1), Otu is broadly expressed in various tissues, including testes and gut (Steinhauer and Kalfayan 1992b). This restricted expression pattern suggests that Sakura may serve as a female germline-specific cofactor that enhances or modifies Otu’s molecular functions. Identifying Otu’s RNA targets and deubiquitinase substrates beyond CycA, as well as determining how Sakura binding influences these activities, will be key to elucidating their roles in oogenesis. One intriguing possibility is that Sakura regulates Otu’s activity toward piRNA pathway components, thereby contributing proper piRNA biogenesis and function. Additionally, the Sakura-Otu complex may directly regulate essential post-transcriptional processes such as sxl alternative splicing and translational control of other oogenic RNAs.

In summary, this study identifies and characterizes the previously unknown gene sakura, which is specifically expressed in female germ cells and is essential for oogenesis and female fertility. Together with its protein partner Otu, Sakura likely regulates germline cell fate, maintenance, and differentiation.

Materials and methods

Fly strains

We generated the sakuranull strain by introducing indels within the Sakura coding region using the CRISPR/Cas9 genome editing system, as we previously reported (Zhu et al. 2018a; Zhu et al. 2018b; Zhu et al. 2019b; Zhu and Fukunaga 2021). The transgenic sakura-EGFP, otu-EGFP, and otu(ΔTudor)-EGFP strains were created following previously published methods (Fukunaga et al. 2012; Kandasamy and Fukunaga 2016; Zhu and Fukunaga 2021). A fragment containing Sakura cDNA flanked by ∼1 kbp upstream and ∼1 kbp downstream genomic sequences was cloned.

Fragments containing Otu and Otu(ΔTudor) cDNAs, flanked by ∼2.5 kbp upstream and ∼1 kbp downstream genomic sequences, were also cloned. The EGFP gene was fused in-frame to the C-terminus of the Sakura and Otu coding sequences within these fragments, which were then inserted into a pattB plasmid vector. The sakura-EGFP plasmid was integrated into the 51C1 site within the fly genome using the BDSC:24482 fly strain and the PhiC31 system, while the otu-EGFP and otu(ΔTudor)-EGFP plasmids were integrated into the 25C6 site of the genome using the attP40 fly strain and the PhiC31 system.

The sakura RNAi #1 (VDRC: v39727) and #2 (VDRC: v103660), y-RNAi (v106068), TOsk-Gal4 (v 314033), Burdock sensor [UAS-Dcr2; NGT-Gal4; nosGal4-VP16, nos>NLS_GF’_lacZ_vas-3’UTR_burdock-target] (v 313217) strains were from Vienna Drosophila Resource Center. w-RNAi (BDSC: 35573), otu-RNAi (BDSC: 34065), bam-RNAi (BDSC: 33631), cycA-RNAi (BDSC: 29313), UAS-Dcr-2; NGT-Gal4 (BDSC: 25751), FRT82B/TM6C, Sb (BDSC: 86313), otu14 (BDSC: 6025), UASp-Mad-GFP (BDSC: 604592), and UASp-tkv.Q253D (BDSC: 604934) were obtained from the Bloomington Stock Center. The bam-GFP reporter (DGRC: 118177) and vasa-EGFP knocked-in fly (DGRC: 118616) were from Kyoto Drosophila Stock Center.

Fertility assay

The female fertility assay was performed as previously described (Zhu et al. 2018b; Liao et al. 2019; Zhu et al. 2019a; Zhu and Fukunaga 2021). Briefly, five test virgin females of each test genotype were mated with three wild-type (OregonR) males in a cage with a 6-cm grape juice agar plate supplemented with wet yeast paste. The agar plates were replaced daily, and the number of eggs laid on the third plate (on day 3) was recorded. After incubation at 25°C for an additional day, the number of hatched eggs was counted. At least three cages per genotype were tested.

For the male fertility assay, a single test male was mated with five wild-type (OregonR) virgin females in each vial, following previously published methods (Zhu et al. 2018b; Zhu et al. 2019a; Zhu and Fukunaga 2021). After 3 days, the females were transferred to a new vial (vial 1). Every two days, they were transferred to a new vial until a total of four vials were obtained. After 2 days in the fourth vial, the females were removed, and the total number of progenies emerging from these four vials was counted. At least five males per genotype were tested.

Sakura and Otu antibodies

We expressed a recombinant full-length Sakura protein as an N-terminal 6xHis-tagged protein in E. coli using a modified pET vector and purified it using Ni-sepharose (GE Healthcare) and HiTrapQ HP (GE Healthcare) columns (Fukunaga and Doudna 2009). Recombinant Otu fragments (145-405aa and 406-853aa) were similarly expressed as N-terminally 6xHis-MBP-fusion proteins in E. coli and purified using Ni-sepharose. These purified proteins were used as antigens to generate polyclonal anti-Sakura and anti-Otu sera in rabbits (Pocono Rabbit Farm & Laboratory, Inc.). The rabbit polyclonal anti-Sakura antibodies were affinity purified using His-MBP-Sakura recombinant protein and Affigel-15 (Bio-rad), following the manufacturer’s instructions. Rabbit anti-Otu sera were first pre-cleared with purified His-MBP protein bound to Affigel-15, and the unbound fraction was affinity purified using purified His-MBP-Otu fragments (145-405aa and 406-853aa) and Affigel-15.

Immunostaining

Stereomicroscope images of dissected ovaries were taken using Leica M125 stereomicrocsope. Ovaries from 2-to 5-day-old, yeast-fed females were hand-dissected in 1X PBS (137 mM NaCl, 2.7 mM KCl, 10 mM Na2HPO4, 1.8 mM KH2PO4, pH 7.4) at room temperature. The dissected ovaries were fixed in a fixative buffer (4% formaldehyde, 15 mM PIPES (pH 7.0), 80 mM KCl, 20 mM NaCl2, 2 mM EDTA, and 0.5 mM EGTA), incubated for 30 minutes at room temperature with gentle rocking. After fixation, the ovaries were rinsed three times with PBX (0.1% Triton X-100 in 1X PBS) and then incubated in a blocking buffer (2% donkey serum in 3% BSA [w/v], 0.02% NAN3 [w/v] in PBX) for 1 hour at room temperature. Then, the ovaries were incubated with primary antibodies diluted in the blocking buffer overnight at 4°C. The following day, the ovaries were rinsed three times with PBX and incubated with Alexa Fluor-conjugated secondary antibodies for 2 hours. The ovaries were rinsed three times with PBX and then mounted in VECTASHIELD® PLUS antifade mounting medium with DAPI (H-2000, Vector lab). Confocal images were acquired on a Zeiss LSM700 confocal microscope at the Johns Hopkins University School of Medicine Microscope Facility.

The primary antibodies used for immunostaining were mouse anti-HTS (1B1) (DSHB, AB_528070, dilution: 1/100), mouse anti-Bam (DSHB, AB_10570327, 1/20), mouse anti-CycA (DSHB, AB_528188, 1/100), rat anti-Vasa (DSHB, AB_760351, 1/100), rat anti-DE Cadherin (DCAD2) (DSHB, AB_528120, 1/20), rabbit anti-pMad (Cell Signaling, Phospho-SMAD1/5 (Ser463/465) mAb #9516, 1/200,), and rabbit anti-cleaved caspase-3 (Cell Signaling, Cleaved Caspase-3 (Asp175) #9661, 1/200). Secondary antibodies used were Alexa Fluor 488 Donkey anti-Mouse Igg (ThermoFisher, A21202, 1/100), Alexa Fluor 594 Donkey anti-Rat Igg (ThermoFisher, A21209, 1/100), Alexa Fluor 594 Donkey anti-Mouse Igg (ThermoFisher, A21203, 1/100), and Alexa Fluor 594 Donkey anti-Rabbit Igg ThermoFisher, A21207, 1/100). Rhodamine phalloidin (ThermoFisher, R415, 1/100) was used to stain F-Actin.

Germline clonal analysis

We generated sakuranull mutant clones using FLP/FRT-mediated recombination (Rubin and Huynh 2015). To induce sakuranull GSC clones, 3-day-old female flies of the genotype hs-flp/w; +; FRT82B, ubi-GFP/FRT82B, sakuranull were heat-shocked at 37°C for 1 hour, twice daily, with an 8-hour interval between heat shocks. Female flies of the genotype hs-flp/w; +; FRT82B, ubi-GFP/FRT82B were used as controls. Ovaries were dissected and stained 4, 7, and 14 days after clone induction. To induce primordial germ cell (PGC) clones, early third-instar larvae were subjected to the same heat shock regime, then allowed to develop to adulthood. Ovaries were dissected and analyzed at 3 to 4 days after eclosion.

To induce sakuranull mutant clones in the presence of otu-EGFP or otu(ΔTudor)-EGFP transgenes, 3-day-old female flies with the genotype hs-flp/w; otu-EGFP/+; FRT82B, ubi-RFP/FRT82B, sakuranull and hs-flp/w; otu(i1Tudor)-EGFP /+; FRT82B, ubi-RFP/FRT82B, sakuranull were heat-shocked under the same conditions. Control genotypes were hs-flp/w; otu-EGFP/+; FRT82B, ubi-RFP/FRT82B and hs-flp/w; otu(i1Tudor)-EGFP/+; FRT82B, ubi-RFP/FRT82B were used as controls. Flies were dissected 3-4 days after clone induction.

Western blot

Lysates of hand-dissected ovaries and tissues were prepared by homogenizing in RIPA buffer (50 mM Tris-HCl [pH 7.4], 150 mM NaCl, 1% [v/v] IGEPAL CA-630, 0.1% [w/v] sodium dodecyl sulfate (SDS), 0.5% [w/v] sodium deoxycholate, 1 mM ethylenediaminetetraacetic acid (EDTA), 5 mM dithiothreitol, and 0.5 mM phenylmethylsulfonyl fluoride (PMSF)) (Kandasamy et al. 2017; Zhu et al. 2018b). The homogenates were centrifuged at 21,000g at 4°C for 10 min, and the protein concentration of the supernatant was determined using the BCA protein assay kit (Pierce) as needed. Fifteen μg of total protein was loaded per lane for Western blot. The sources and dilutions of the primary antibodies were as below. Rabbit anti-Sakura (1/10000, generated in this study), rabbit anti-Otu (1/10000, generated in this study), rabbit anti-alpha-Tubulin [EP1332Y] (1/10000, Abcam, ab52866), mouse anti-alpha-Tubulin [12G10] (1/10000, DSHB, AB_1157911), mouse anti-FLAG (1/10000, Sigma, F1804), mouse anti-HA (1/10000, Sigma, H3663), and mouse anti-GFP [GF28R] (1/3000, Invitrogen, 14-6674-82). IRDye 800CW goat anti-mouse IgG, IRDye 800CW goat anti-rabbit IgG, IRDye 680RD goat anti-mouse, and IgG IRDye 680RD goat anti-rabbit were used as secondary antibodies. The membranes were scanned using the Li-Cor Odyssey CLx Imaging System.

Mass spectrometry

Immunoprecipitation of Sakura-EGFP protein was performed using the GFP-Trap Magnetic Agarose Kit (Proteintech, gtmak-20) on dissected ovaries from flies harboring sakura-EGFP transgene, with w1118 flies as controls. Ovaries were homogenized in 200 μL ice-cold lysis buffer (10 mM Tris-HCl [pH 7.5], 150 mM NaCl, 0.5 mM EDTA, 0.05% [v/v] IGEPAL CA-630) containing 1× protease inhibitor cocktail (100× protease inhibitor cocktail contains 120 mg/ml 1 mM 4-(2-aminoethyl) benzene sulfonyl fluoride hydrochloride (AEBSF), 1 mg/ml aprotinin, 7 mg/ml bestatin, 1.8 mg/ml E-64, and 2.4 mg/ml leupeptin). After homogenization, the tubes were placed on ice for 30 minutes, and the homogenates were extensively pipetted every 10 minutes. The lysates were then centrifuged at 17,000x g for 10 minutes at 4°C. The supernatants were transferred to pre-chilled tubes, and 300 μL dilution buffer (10 mM Tris/Cl pH 7.5, 150 mM NaCl, 0.5 mM EDTA) supplemented with 1x protease inhibitor cocktail were added. The diluted lysates were then added to the GFP-trap magnetic beads in 1.5 mL tubes and rotated for 1 hour at 4°C. After separating the beads with a magnetic tube rack, the beads were washed three times with 500 μL wash buffer (10 mM Tris/Cl pH 7.5, 150 mM NaCl, 0.05 % [v/v] IGEPAL CA-630). Proteins were eluted with 40 μL acidic elution buffer (200 mM glycine pH 2.5) followed by immediate neutralization with 5 uL neutralization buffer (1 M Tris pH 10.4).

As a quality control before mass spectrometry, ∼5 μL of the samples were mixed with an equal volume of 2× SDS-PAGE loading buffer (80 mM Tris-HCl [pH 6.8], 2% [w/v] SDS, 10% [v/v] glycerol, 0.0006% [w/v] bromophenol blue, 2% [v/v] 2-mercaptoethanol), heated at 95 °C for 3 min, and run on 4–20% Mini-PROTEAN® TGX™ Precast Protein Gels (Biorad, #4561094). Silver staining was then performed by using Pierce™ Silver Stain Kit (ThermoFisher, 24612) to assess the quality of the immunoprecipitated protein samples. Mass spectrometry was conducted at the Mass Spectrometry Core at the Department of Biological Chemistry, Johns Hopkins School of Medicine, as previously described (Zhu and Fukunaga 2021).

Co-immunoprecipitation

For co-immunoprecipitation of endogenous Sakura protein, wild-type ovaries (w1118) were homogenized in RIPA buffer, centrifuged at 21,000g at 4°C for 10 min, and the clear supernatant protein lysates were used for immunoprecipitation. Four μg of rabbit anti-Sakura and normal rabbit IgG (Cell Signaling, #2729) were incubated with 50 μL of Dynabeads Protein G (ThermoFisher, 10004D) for 20 min at room temperature. The beads were washed once with PBST (1x PBS with 0.1% Tween-20). The ovary lysate supernatant was then incubated with the washed beads at room temperature for 30 min, followed by three washes with PBST. The proteins were eluted with 2x SDS-PAGE loading buffer (80 mM Tris-HCl [pH 6.8], 2% [w/v] SDS, 10% [v/v] glycerol, 0.0006% [w/v] bromophenol blue, 2% [v/v] 2-mercaptoethanol) and heated at 70°C for 10 min. After bead separation using a magnetic tube rack, the eluted proteins in 2x SDS-PAGE loading buffer were heated again at 95°C for 3 min.

Transient protein expression in S2 cells was performed using the pAc5.1/V5-HisB plasmid vector (Invitrogen). A total of 1 μg plasmids were transfected using the Effectene transfection reagent (Qiagen, 301425). Three days after transfection, cells were harvested and lysed with ice-cold lysis buffer supplemented with 1× protease inhibitor cocktail. The cell lysates were then centrifuged at 17,000x g for 10 min at 4°C, and the clear supernatants were collected for immunoprecipitation. For anti-HA immunoprecipitation, supernatants were incubated with 25 μL (0.25 mg) of Pierce™ Anti-HA Magnetic Beads (ThermoFisher, 88837) at room temperature for 30 min. The beads were then washed three times with TBST (1x TBS [0.05 M Tris/HCl and 0.15 M NaCl, pH 7.6] with 0.05% Tween-20). For anti-FLAG immunoprecipitation, 2 μg of mouse anti-FLAG (1/10000, Sigma, F1804) was incubated with 50 μL of Dynabeads Protein G (ThermoFisher, 10004D) for 10 min at room temperature. The beads were washed once with PBST. The S2 cell lysate supernatant was incubated with the washed beads at room temperature for 15 min. The beads were washed three times with PBST. In both anti-HA and anti-FLAG immunoprecipitations, proteins were eluted with 2× SDS-PAGE loading buffer. For anti-HA, the beads in 2x SDS-PAGE loading buffer were heated at 95°C for 7 min. For anti-FLAG, the beads were heated at 70°C for 10 min, then were separated using a magnetic tube rack. The eluted proteins in 2x SDS-PAGE loading buffer were heated again at 95°C for 3 min.

Small RNA and mRNA sequencing

Poly-A+ mRNA purification was performed as previously described (Fukunaga et al. 2012). Small RNA libraries and poly-A+ mRNA libraries were prepared, sequenced on Hiseq2500 (Illumina), and analyzed, as previously reported (Fukunaga et al. 2014; Kandasamy et al. 2017; Liao et al. 2018; Zhu et al. 2018a; Zhu et al. 2018b; Liao et al. 2019; Zhu et al. 2019a; Zhu et al. 2019b). SRA accession number for these datasets is PRJNA1156618.

RT-PCR

Total RNAs from testes and ovaries were prepared using miRVana (Thermo Fisher Scientific). RNAs were treated with Turbo DNase (Thermo Fisher Scientific) to remove potential genomic DNA contamination. A total of 1 μg of RNA was reverse-transcribed into cDNA using SuperScriptTM VILOTM MasterMix (Thermo Fisher Scientific). To assess sxl alternative splicing, PCR was performed using GoTaq Green Master Mix (Promega) with the primers sxl-F (CTCACCTTCGATCGAGGGTGTA) and sxl-R (GATGGCAGAGAATGGGAC).

In vitro deubiquitination assay

C-terminally 6xHis-tagged recombinant Sakura protein was expressed in E. coli using a modified pET vector and was purified using Ni-sepharose. N-terminally 3xHA-HRV3Csite-3xFLAG-tagged Otu protein and a negative control, N-terminally 3xHA-HRV3Csite-3xFLAG-tagged firefly luciferase protein, were expressed in S2 cells using pAc5.1/V5-HisB vector. The proteins were immunoprecipitated using Pierce™ Anti-HA Magnetic Beads (ThermoFisher, 88837) as described previously, with a more stringent washing step to effectively remove interacting proteins. The beads were washed six times with a high salt wash buffer (TBST with 800 mM NaCl). The proteins were eluted by cleaving off 3xHA tag with 25 nM GST-HRV3C protease in a cleavage buffer (25mM Tris-HCl pH 7.4, 150 mM NaCl, 5% glycerol, 2mM EDTA), incubated at 4°C with rotation for 6 hours. An equal volume of 100% glycerol was added to the eluted protein.

Purified proteins, including Otu, luciferase, and Sakura, were mixed with Ub-Rhodamine 110 (Ubiquitin-Proteasome Biotechnologies, M3020) in a total volume of 30 μL in reaction buffer (20 mM Tris-HCl, pH 7.5, 200 mM NaCl, 5 mM MgCl2, 2 mM DTT). The mixture was added to a black 384-well low-volume plate. Fluorescence measurements were taken using SpectraMax i3x Multi-Mode Microplate Reader at 37°C, with excitation and emission wavelengths set at 485/20 and 530/20 nm respectively. The fluorescence intensity for each condition was averaged from triplicates and plotted as a function of time.

eLife Assessment

This valuable study reports the first characterization of the CG14545 gene in Drosophila melanogaster, which the authors name "Sakura." Acting during germline stem cell fate and differentiation, Sakura is required for both oogenesis and female fertility, although some mechanistic details require further investigation. This solid study presents a wide-ranging and well-controlled characterization of Sakura, and accordingly the findings and associated reagents described will be of use to scientists interested in oogenesis and early development.

Reviewer #1 (Public review):

Summary:

In this manuscript, Azlan et al. identified a novel maternal factor called Sakura that is required for proper oogenesis in Drosophila. They showed that Sakura is specifically expressed in the female germline cells. Consistent with its expression pattern, Sakura functioned autonomously in germline cells to ensure proper oogenesis. In sakura KO flies, germline cells were lost during early oogenesis and often became tumorous before degenerating by apoptosis. In these tumorous germ cells, piRNA production was defective and many transposons were derepressed. Interestingly, Smad signaling, a critical signaling pathway for the GSC maintenance, was abolished in sakura KO germline stem cells, resulting in ectopic expression of Bam in whole germline cells in the tumorous germline. A recent study reported that Bam acts together with the deubiquitinase Otu to stabilize Cyc A. In the absence of sakura, Cyc A was upregulated in tumorous germline cells in the germarium. Furthermore, the authors showed that Sakura co-immunoprecipitated Otu in ovarian extracts. A series of in vitro assays suggested that the Otu (1-339 aa) and Sakura (1-49 aa) are sufficient for their direct interaction. Finally, the authors demonstrated that the loss of otu phenocopies the loss of sakura, supporting their idea that Sakura plays a role in germ cell maintenance and differentiation through interaction with Otu during oogenesis.

Strengths:

To my knowledge, this is the first characterization of the role of CG14545 genes. Each experiment seems to be well-designed and adequately controlled

Weaknesses:

However, the conclusions from each experiment are somewhat separate, and the functional relationships between Sakura's functions are not well established. In other words, although the loss of Sakura in the germline causes pleiotropic effects, the cause-and-effect relationships between the individual defects remain unclear.

Comments on latest version:

The authors have attempted to address my initial concerns with additional experiments and refutations. Unfortunately, my concerns, especially my specific comments 1-3, remain unaddressed. The present manuscript is descriptive and fails to describe the molecular mechanism by which Sakura exerts its function in the germline. Nevertheless, this reviewer acknowledges that the observed defects in sakura mutant ovaries and the possible physiological significance of the Sakura-Out interaction are worth sharing with the research community, as they may lay the groundwork for future research in functional analysis.

Reviewer #1 (Public review):

Summary:

In this manuscript, Azlan et al. identified a novel maternal factor called Sakura that is required for proper oogenesis in Drosophila. They showed that Sakura is specifically expressed in the female germline cells. Consistent with its expression pattern, Sakura functioned autonomously in germline cells to ensure proper oogenesis. In sakura KO flies, germline cells were lost during early oogenesis and often became tumorous before degenerating by apoptosis. In these tumorous germ cells, piRNA production was defective and many transposons were derepressed. Interestingly, Smad signaling, a critical signaling pathway for the GSC maintenance, was abolished in sakura KO germline stem cells, resulting in ectopic expression of Bam in whole germline cells in the tumorous germline. A recent study reported that Bam acts together with the deubiquitinase Otu to stabilize Cyc A. In the absence of sakura, Cyc A was upregulated in tumorous germline cells in the germarium. Furthermore, the authors showed that Sakura co-immunoprecipitated Otu in ovarian extracts. A series of in vitro assays suggested that the Otu (1-339 aa) and Sakura (1-49 aa) are sufficient for their direct interaction. Finally, the authors demonstrated that the loss of otu phenocopies the loss of sakura, supporting their idea that Sakura plays a role in germ cell maintenance and differentiation through interaction with Otu during oogenesis.

Strengths:

To my knowledge, this is the first characterization of the role of CG14545 genes. Each experiment seems to be well-designed and adequately controlled

Weaknesses:

However, the conclusions from each experiment are somewhat separate, and the functional relationships between Sakura's functions are not well established. In other words, although the loss of Sakura in the germline causes pleiotropic effects, the cause-and-effect relationships between the individual defects remain unclear.

Comments on latest version:

The authors have attempted to address my initial concerns with additional experiments and refutations. Unfortunately, my concerns, especially my specific comments 1-3, remain unaddressed. The present manuscript is descriptive and fails to describe the molecular mechanism by which Sakura exerts its function in the germline. Nevertheless, this reviewer acknowledges that the observed defects in sakura mutant ovaries and the possible physiological significance of the Sakura-Out interaction are worth sharing with the research community, as they may lay the groundwork for future research in functional analysis.

Reviewer #3 (Public review):

In this very thorough study, the authors characterize the function of a novel Drosophila gene, which they name Sakura. They start with the observation that sakura expression is predicted to be highly enriched in the ovary and they generate an anti-sakura antibody, a line with a GFP-tagged sakura transgene, and a sakura null allele to investigate sakura localization and function directly. They confirm the prediction that it is primarily expressed in the ovary and, specifically, that it is expressed in germ cells, and find that about 2/3 of the mutants lack germ cells completely and the remaining have tumorous ovaries. Further investigation reveals that Sakura is required for piRNA-mediated repression of transposons in germ cells. They also find evidence that sakura is important for germ cell specification during development and germline stem cell maintenance during adulthood. However, despite the role of sakura in maintaining germline stem cells, they find that sakura mutant germ cells also fail to differentiate properly such that mutant germline stem cell clones have an increased number of "GSC-like" cells. They attribute this phenotype to a failure in the repression of Bam by dpp signaling. Lastly, they demonstrate that sakura physically interacts with otu and that sakura and otu mutants have similar germ cell phenotypes. Overall, this study helps to advance the field by providing a characterization of a novel gene that is required for oogenesis. The data are generally high-quality and the new lines and reagents they generated will be useful for the field.

Comments on latest version:

With these revisions, the authors have addressed my main concerns.

Author response:

The following is the authors’ response to the previous reviews

Public Reviews:

Reviewer #1 (Public review):

Summary:

In this manuscript, Azlan et al. identified a novel maternal factor called Sakura that is required for proper oogenesis in Drosophila. They showed that Sakura is specifically expressed in the female germline cells. Consistent with its expression pattern, Sakura functioned autonomously in germline cells to ensure proper oogenesis. In sakura KO flies, germline cells were lost during early oogenesis and often became tumorous before degenerating by apoptosis. In these tumorous germ cells, piRNA production was defective and many transposons were derepressed. Interestingly, Smad signaling, a critical signaling pathway for the GSC maintenance, was abolished in sakura KO germline stem cells, resulting in ectopic expression of Bam in whole germline cells in the tumorous germline. A recent study reported that Bam acts together with the deubiquitinase Otu to stabilize Cyc A. In the absence of sakura, Cyc A was upregulated in tumorous germline cells in the germarium. Furthermore, the authors showed that Sakura co-immunoprecipitated Otu in ovarian extracts. A series of in vitro assays suggested that the Otu (1-339 aa) and Sakura (1-49 aa) are sufficient for their direct interaction. Finally, the authors demonstrated that the loss of otu phenocopies the loss of sakura, supporting their idea that Sakura plays a role in germ cell maintenance and differentiation through interaction with Otu during oogenesis.

Strengths:

To my knowledge, this is the first characterization of the role of CG14545 genes. Each experiment seems to be well-designed and adequately controlled

Weaknesses:

However, the conclusions from each experiment are somewhat separate, and the functional relationships between Sakura's functions are not well established. In other words, although the loss of Sakura in the germline causes pleiotropic effects, the cause-and-effect relationships between the individual defects remain unclear.

Comments on latest version:

The authors have attempted to address my initial concerns with additional experiments and refutations. Unfortunately, my concerns, especially my specific comments 1-3, remain unaddressed. The present manuscript is descriptive and fails to describe the molecular mechanism by which Sakura exerts its function in the germline. Nevertheless, this reviewer acknowledges that the observed defects in sakura mutant ovaries and the possible physiological significance of the Sakura-Out interaction are worth sharing with the research community, as they may lay the groundwork for future research in functional analysis.

We thank the reviewer for valuable comments. We would like to investigate the molecular mechanism by which Sakura exerts its function in the germline in near future studies.

Reviewer #2 (Public review):

In this study, the authors identified CG14545 (named it sakura), as a key gene essential for Drosophila oogenesis. Genetic analyses revealed that Sakura is vital for both oogenesis progression and ultimate female fertility, playing a central role in the renewal and differentiation of germ stem cells (GSC).

The absence of Sakura disrupts the Dpp/BMP signaling pathway, resulting in abnormal bam gene expression, which impairs GSC differentiation and leads to GSC loss. Additionally, Sakura is critical for maintaining normal levels of piRNAs. Also, the authors convincingly demonstrate that Sakura physically interacts with Otu, identifying the specific domains necessary for this interaction, suggesting a cooperative role in germline regulation. Importantly, the loss of otu produces similar defects to those observed in sakura mutants, highlighting their functional collaboration.

The authors provide compelling evidence that Sakura is a critical regulator of germ cell fate, maintenance, and differentiation in Drosophila. This regulatory role is mediated through modulation of pMad and Bam expression. However, the phenotypes observed in the germarium appear to stem from reduced pMad levels, which subsequently trigger premature and ectopic expression of Bam. This aberrant Bam expression could lead to increased CycA levels and altered transcriptional regulation, impacting piRNA expression. In this revised manuscript, the authors further investigated whether Sakura affects the function of Orb, a binding partner they identified, in deubiquitinase activity when Orb interacts with Bam.

We appreciate the authors' efforts to address all our comments. While these revisions have greatly improved the clarity of certain sections, some of the concerns remain unclear, while details mentioned in the responses about these studies should be incorporated in the manuscript. Specifically, the manuscript still lacks the demonstration that Sakura co-localizes with Orb/Bam despite having the means for staining and visualization. This would bring insight into the selective binding of Orb with Bam vs. Sakura perhaps at different stages of oogenesis. Such analyses would allow for more specific conclusions, further alluding to the underlying mechanism, rather than the general observations currently presented.

This elaborate study will be embraced by both germline-focused scientists and the developmental biology community.

We thank the reviewer for valuable comments. We believe that the author meant Otu, not Orb, for the binding partner of Sakura that we identified. We would like to investigate the colocalization of Sakura with other proteins including Otu and the molecular mechanism by which Sakura exerts its function in the germline in near future studies.

Reviewer #3 (Public review):

In this very thorough study, the authors characterize the function of a novel Drosophila gene, which they name Sakura. They start with the observation that sakura expression is predicted to be highly enriched in the ovary and they generate an anti-sakura antibody, a line with a GFP-tagged sakura transgene, and a sakura null allele to investigate sakura localization and function directly. They confirm the prediction that it is primarily expressed in the ovary and, specifically, that it is expressed in germ cells, and find that about 2/3 of the mutants lack germ cells completely and the remaining have tumorous ovaries. Further investigation reveals that Sakura is required for piRNA-mediated repression of transposons in germ cells. They also find evidence that sakura is important for germ cell specification during development and germline stem cell maintenance during adulthood. However, despite the role of sakura in maintaining germline stem cells, they find that sakura mutant germ cells also fail to differentiate properly such that mutant germline stem cell clones have an increased number of "GSC-like" cells. They attribute this phenotype to a failure in the repression of Bam by dpp signaling. Lastly, they demonstrate that sakura physically interacts with otu and that sakura and otu mutants have similar germ cell phenotypes. Overall, this study helps to advance the field by providing a characterization of a novel gene that is required for oogenesis. The data are generally high-quality and the new lines and reagents they generated will be useful for the field.

Comments on latest version:

With these revisions, the authors have addressed my main concerns.

We thank the reviewer for valuable comments.

Recommendations for the authors:

Reviewer #2 (Recommendations for the authors):

The manuscript is much improved based on the changes made upon recommendations from the reviewers.

Though most of our comments have been addressed, we have a few more we wish to recommend. For previous points we made, we replied with further clarification for the authors.

Figure 1

(1) B should be the supplemental figure.

We moved the former Fig 1B to Supplemental Figure 1.

• Previous Fig1B (sakura mRNA expression level) is now Fig S2, not S1. Please make this data as Fig S1.

We moved Fig S1 to main Fig7A and renumbered Fig S2-S16 to Fig S1-S15.

(2) C - How were the different egg chamber stages selected in the WB? Naming them 'oocytes' is deceiving. Recommend labeling them as 'egg chambers', since an oocyte is claimed to be just the one-cell of that cyst.

We changed the labeling to egg chambers.

• The labels on lanes for Stages 12-13 and Stage 14, still only say "chambers", not "egg chambers". Also there is no Stage 1-3 egg chamber. More accurately, the label should be "Germarium - Stage 11 egg chambers".

We updated the lables on lanes as suggested by the reviewer.

(3) Is the antibody not detecting Sakura in IF? There is no mention of this anywhere in the manuscript.

While our Sakura antibody detects Sakura in IF, it seems to detect some other proteins as well. Since we have Sakura-EGFP fly strain (which fully rescues sakuranull phenotypes) to examine Sakura expression and localization without such non-specific signal issues, we relied on Sakura-EGFP rather than anti-Sakura antibodies for IF.

• Please put this info into the Methods section.

We added this info into the Methods section.

(4) Expand on the reliance of the sakura-EGFP fly line. Does this overexpression cause any phenotypes?

sakura-EGFP does not cause any phenotypes in the background of sakura[+/+] and sakura[+/-].

• Please add this detail into the manuscript.

We added this info into the Methods section.

Figure 5

(1) D - It might make more sense if this graph showed % instead of the numbers.

We did not understand the reviewer's point. We think using numbers, not %, makes more sense.

• Having a different 'n' number for each experiment does not allow one to compare anything except numbers of the egg chambers. This must be normalized.

We still don’t agree with the reviewer. In Fig 5D, we are showing the numbers of stage 14 oocytes per fly (= per a pair of ovaries). ‘n’ is the number of flies (= number of a pair of ovaries) examined. We now clarified this in the figure legend. Different ‘n’ number does not prevent us from comparing the numbers of stage 14 oocytes per fly. Therefore, we would like to show as it is now.

(2) Line 213 - explain why RNAi 2 was chosen when RNAi 1 looks stronger.

Fly stock of RNAi line 2 is much healthier than RNAi line 1 (without being driven Gal4) for some reasons. We had a concern that the RNAi line 1 might contain an unwanted genetic background. We chose to use the RNAi 2 line to avoid such an issue.

• Please add this information to the manuscript.

We added this info into the Methods section.

Figure 7/8 - can go to Supplemental.

We moved Fig 8 to supplemental. However, we think Fig 7 data is important and therefore we would like to present them as a main figure.

• Current Fig S1 should go to Fig 7, to better understand the relationship between pMad and Bam expression.

We moved Fig S1 to main Fig7A and renumbered Fig S2-S16 to Fig S1-S15.

Figure 9C - Why the switch to S2 cells? Not able to use the Otu antibody in the IP of ovaries?

We can use the Otu antibody in the IP of ovaries. However, in anti-Sakura Western after anti Otu IP, antibody light chain bands of the Otu antibodies overlap with the Sakura band. Therefore, we switched to S2 cells to avoid this issue by using an epitope tag.

• Please add this info to the Methods section.

We added this info into the Methods section.

Figure 10- Some images would be nice here to show that the truncations no longer colocalize.

We did not understand the reviewer's points. In our study, even for the full-length proteins. We have not shown any colocalization of Sakura and Otu in S2 cells or in ovaries, except that they both are enriched in developing oocytes in egg chambers.

• Based on your binding studies, we would expect them to colocalize in the egg chamber, and since there are antibodies and a GFP-line available, it would be important to demonstrate that via visualization.

As we wrote in the response and now in the manuscript, our antibodies are not best for immunostaining. We will try to optimize the experimental conditions in the future studies.

Version 3

Abstract

Abstract

During oogenesis, self-renewal and differentiation of germline stem cells (GSCs) must be tightly regulated. The Drosophila female germline serves as an excellent model for studying these regulatory mechanisms. Here, we report that a previously uncharacterized gene CG14545, which we named sakura, is essential for oogenesis and female fertility in Drosophila. Sakura is predominantly expressed in the ovaries, particularly in the germline cells, including GSCs. sakura null mutant female flies display rudimentary ovaries with germline-less and tumorous phenotypes, fail to produce eggs, and are completely sterile. The germline-specific depletion of sakura impairs Dpp/BMP signaling, leading to aberrant bag-of-marbles (bam) expression, resulting in faulty differentiation and loss of GSCs. sakura is also necessary for normal levels of piwi-interacting RNAs (piRNAs) levels and for female-specific splicing of sex-lethal (sxl), a master regulator of sex identity determination. We identified Ovarian Tumor (Otu) as a protein binding partner of Sakura and found that loss of otu phenocopies loss of sakura in ovaries. Thus, we identify Sakura as a crucial factor for GSC renewal and differentiation and oogenesis, and propose that Sakura and Otu function together in these processes.

Introduction

Oogenesis is a complex process in which germline stem cells (GSCs) develop into mature female gametes or oocytes. This process involves multiple layers of regulations and numerous genes, some of which remain to be discovered and characterized. Drosophila melanogaster, a genetically tractable model organism, has long been at the forefront of GSC and oogenesis research (Kirilly and Xie 2007).

Adult Drosophila females possess a pair of ovaries, each composed of 12-16 ovarioles. At the anterior tip of each ovariole is a structure called the germarium, which houses two to three GSCs (Fig 1A). GSCs reside in the most anterior region of the germarium, where they directly contact cap cells and escort cells, forming a niche crucial for GSC self-renewal (Lin and Spradling 1993; de Cuevas and Matunis 2011; Losick et al. 2011). GSCs undergo asymmetric mitotic division, producing two distinct cells: one GSC and one cystoblast. The cystoblast, destined to differentiate, undergoes four mitotic divisions with incomplete cytokinesis, resulting in interconnected cysts of 2, 4, 8, and 16 cells. Of the 16 cells, one will differentiate into an oocyte, while the remaining 15 develop into polyploid nurse cells, which synthesize proteins and RNAs for deposition into the oocyte.

Sakura expression pattern and mutant allele

(A) Schematic illustration of Drosophila ovary and germarium. Drosophila female has a pair of ovaries, each consisting of 12-16 ovarioles (cyan). The germarium (outlined with magenta) is located at the anterior tip of the ovarioles and consists of both germ cells and somatic cells. Germ cells include germline stem cells (GSCs), cystoblasts, cysts, and differentiating oocytes. Somatic cells include terminal filament (TF) cells, cap cells (CCc), escort cells (ECs), and follicle cells (FCs). GSCs and cystoblasts have spherical, unbranched spectrosomes, whereas cysts posesss branched fusomes. The distinct regions of the germarium—1, 2a, 2b, and 3—are indicated. (B) Drosophila Sakura protein (Sakura/CG14545) and its null mutant allele generated in this study. (C) Western blot of dissected fly tissues. (D) Western blot of ovary lysates. (E) Confocal images of the germarium from the sakura-EGFP transgenic fly. Sakura-EGFP (green), Hts (red), and DAPI (blue). Scale bar: 10 μm. (F) Confocal images of the egg chambers from the sakura-EGFP transgenic fly. Sakura-EGFP (green), Vasa (red), and DAPI (blue). Sakura-EGFP is expressed in nurse cells and enriched in the developing oocyte (yellow arrow). Scale bar: 20 μm.

GSCs must be tightly regulated to maintain their undifferentiated state while dividing and differentiating to produce one GSC and one cystoblast. Failure in GSC self-renewal or inappropriate differentiation leads to stem cell loss, disrupting oocyte production (Lin 1997; Cox et al. 1998). Conversely, uncontrolled GSC division or defective differentiation results in an overabundance of GSC-like cells (a tumorous phenotype) and fewer differentiated cells, ultimately inhibiting oocyte production (Gateff et al. 1996; Ohlstein et al. 2000). Thus, a precise balance between GSC self-renewal and cyst differentiation within the germarium is essential for proper oogenesis and is tightly regulated.

Bone Morphogenetic Protein (BMP) signaling plays a critical role in regulating GSC self-renewal and differentiation during Drosophila oogenesis. In the stem cell niche, GSCs are physically anchored to cap cells via adherens and gap junctions (Song et al. 2002; Gilboa et al. 2003). Cap cells secrete BMP ligands, such as Decapentaplegic (Dpp) and Glass bottom boat (Gbb), which are recognized by transducing receptors like Thickvein (Tkv) and Saxophone (Sax) on the GSCs (Xie and Spradling 1998; Casanueva and Ferguson 2004). Upon activation, these receptors phosphorylate a transcription factor Mad. The phosphorylated Mad (pMad) translocates into the nucleus, where it represses the transcription of bag-of-marbles (bam), a key differentiation factor (Kai and Spradling 2003; Song et al. 2004).

Bam, a ubiquitin-associated protein, is essential for cystoblast differentiation (McKearin and Spradling 1990; McKearin and Ohlstein 1995). Dpp/BMP signaling in the niche represses bam expression, preventing GSCs from differentiating and enabling self-renew (Song et al. 2004). However, as the daughter cells (cytoblasts) exit the niche, they derepress bam and initiate the differentiation program. Tight regulation of Dpp/BMP signaling and bam expression ensures that only cells within the niche adopt GSC fate, while adjacent daughter cells differentiate into cystoblasts. Loss of bam leads to blocked differentiation and the accumulation of GSC-like cells (tumorous phenotype), whereas ectopic bam expression forces GSCs differentiation, depleting the stem cell pool (McKearin and Spradling 1990; McKearin and Ohlstein 1995; Ohlstein and McKearin 1997).

One of the proteins that interacts with Bam is Ovarian Tumor (Otu) (Ji et al. 2017). Otu forms a deubiquitinase complex with Bam, which deubiquitinates Cyclin A (CycA), stabilizing CycA and promoting GSC differentiation (Ji et al. 2017). Otu also binds RNAs (Ji et al. 2019). Otu is crucial for oogenesis and female fertility. Mutations in the otu gene cause various ovarian defects, including tumorous growths, germ cell loss, abnormalities in oocyte determination and nurse cell dumping, and defects in the sexual identity determination of germ cells, indicating Otu’s importance in multiple stages of oogenesis (Smith and King 1966; Gans et al. 1975; Gollin and King 1981; King and Riley 1982; Storto and King 1988; Pauli et al. 1993; Rodesch et al. 1997; Glenn and Searles 2001). However, the precise molecular mechanisms by which Otu functions in these processes remain largely unknown.

In this study, we aimed to identify a novel gene required for oogenesis. Through a search for uncharacterized genes expressed exclusively in Drosophila ovaries, we discovered CG14545, a gene encoding a 114 amino acids (aa) long protein without any known functional domain (Fig 1B). We found that CG14545 is expressed specifically in female germline cells, including GSCs. Our genetic analysis revealed that CG14545 is required for oogenesis and female fertility and plays an important intrinsic role in regulating GSCs renewal and differentiation. Additionally, CG14545 is important for the piRNA pathway and for the female-specific mRNA splicing pattern of sex-lethal (sxl), a master regulator of sex determination. We identified Otu as a binding partner of the CG14545 protein, suggesting that both proteins may participate in the same molecular pathways to regulate GSC fate and differentiation. We named the CG14545 gene sakura, meaning “cherry blossom” in Japanese, symbolizing birth and renewal.

Results

Sakura is exclusively expressed in the ovaries

Based on high-throughput expression data in Flybase, sakura mRNA is highly and exclusively expressed in adult ovaries (Fig S1). To investigate Sakura protein expression, we generated a polyclonal anti-Sakura antibody against a recombinant full-length Sakura protein. Using this antibody, we examine Sakura protein expression at different stages of Drosophila development and across tissues. We found that Sakura protein is specifically expressed in ovaries, particularly in germarium - 11 egg chambers (Fig 1C).

sakura mutant flies

To explore the biological and molecular functions of Sakura in vivo, we generated a sakura mutant allele (sakuranull) with a one nucleotide (nt) insertion replacing 33 nts within the Sakura coding region using the CRISPR/Cas9 genome editing system (Fig 1B). This mutation causes a frameshift and a premature stop codon, resulting in a 33-amino acid (aa) N-terminal fragment of Sakura followed by a 22-aa segment produced by frameshifted translation (Fig 1B). This truncated fragment is unlikely to be functional, and we were unable to detect stable protein expression, as shown below (Fig 1D). Thus, we consider this allele a null mutation. sakuranull/null flies were viable, revealing that Sakura is not essential for survival.

To validate the sakuranull strain, we performed Western blots using ovary lysates from sakura mutant flies using the anti-Sakura antibody. As expected, full-length Sakura protein was detected in the ovaries of wild-type (sakura+/+) and heterozygous controls (sakuranull/+), but not in those of homozygous mutants (sakuranull/null) (Fig 1D). No smaller protein corresponding to the truncated Sakura fragment was detected in either sakuranull/+ or sakuranull/null ovaries thus, suggesting that the fragment is unstable or not expressed.

Sakura is cytoplasmic and expressed specifically in germ cells

To determine the localization of Sakura protein within the ovaries, we generated transgenic flies carrying a Sakura coding sequence fused with EGFP at the C-terminus (Sakura-EGFP) under the control of the sakura promoter. Western blot analysis of ovary lysates showed that Sakura-EGFP is expressed at levels comparable to endogenous Sakura (Fig 1D).

Oogenesis begins in the germarium, which contains 2-3 GSCs identified by round, unbranched spectrosomes that contact cap cells (Fig 1A) (Kirilly and Xie 2007). In contrast, cysts exhibit branched fusomes. Immunostaining with hu-li tai shao (HTS) antibody marks both spectrosomes and fusomes (Fig 1E), while Vasa is a known germ cell marker (Fig 1F). Confocal imaging revealed that Sakura-EGFP is specifically expressed in the cytoplasm of germ cells, including GSCs, cyst cells, nurse cells, and developing oocytes (Fig 1E and 1F). Sakura-EGFP was not expressed in somatic cells such as terminal filament, cap cells, escort cells, or follicle cells. (Fig 1E). In the egg chamber, Sakura-EGFP was detected in the cytoplasm of nurse cells and was enriched in developing oocytes (Fig 1F).

sakura is essential for female fertility

Given its ovary-specific expression at both mRNA and protein levels (Fig 1C and S2), we hypothesized that Sakura plays a critical role in ovarian function. Fertility assays revealed that sakuranull/nullfemales laid no eggs when mated with wild-type males, while control females (sakura+/+ and sakuranull/+) laid numerous eggs (Fig 2A and 2B). Dissection of sakuranull/null ovaries revealed that they were rudimentary compared to normal ovaries in controls (Fig 2C). This suggests that the inability of sakuranull/null female to lay eggs is due to their underdeveloped ovaries. In contrast, sakuranull/null males were fertile and showed no significant differences compared to controls (sakura+/+ and sakuranull/+) (Fig S2).

sakura mutant flies are female-sterile and have rudimentary ovaries.

(A-B) Female fertility assays. (A) The number of eggs laid by test females crossed with OregonR wild-type males and (B) hatching rates of the eggs. Mean ± SD (n = 3). P-value < 0.001 (Student’s t-test, unpaired, two-tailed) is indicated by ***. (C) Stereomicroscope images of dissected whole ovaries. Scale bar: 200 μm.

To confirm that the observed female sterility was caused by the loss of sakura, we generated sakura rescue flies expressing Sakura-EGFP in the sakuranull/null background (sakuranull/null; sakura-EGFP). Western blotting confirmed that these flies expressed Sakura-EGFP, but not endogenous Sakura (Fig 1D). The sakura-EGFP transgene fully rescued both female fertility and ovary morphology in the sakuranull/null background (Fig 2). sakura-EGFP rescue females showed no significant fertility differences compared to controls (Fig 2A and 2B), and their ovaries were normal in shape and size (Fig 2C). We concluded that Sakura is essential for female fertility and normal ovary morphology but dispensable for male fertility, consistent with the observation that Sakura is predominantly expressed in ovarian germline cells (Fig 1).

sakuranull/null ovaries are germless and tumorous

Given the rudimentary appearance of sakuranull/null ovaries, we questioned whether they contained germ cells. We used flies carrying vasa-EGFP knocked-in reporter in the sakuranull/null background, as Vasa is a known germ cell marker. We observed that some sakuranull/nullovarioles were devoid of germ cells (“germless”, cyan stars), while others retained germ cells (Fig 3A and 3B). Immunostaining with HTS antibody revealed an excess number of GSC-like cells with round spectrosomes in sakuranull/null ovarioles containing germ cells (Fig 3C, orange stars), indicative of a “tumorous” phenotype as previously described for mutants of other genes including bam, otu, and sxl (Smith and King 1966; Gans et al. 1975; Gollin and King 1981; King and Riley 1982; McKearin and Spradling 1990; Bopp et al. 1993; Eliazer et al. 2011; Jin et al. 2013; Yang et al. 2019). Additionally, we observed an excess number of cyst cells with branched fusomes that persisted throughout the ovarioles, suggesting abnormal cyst cell differentiation and division. Thus, loss of sakura results in both germ cell depletion and overgrowth.

sakuranull ovaries are germless and tumorous

(A, B) Confocal images of the ovaries from control (sakuranull/+) and sakuranull/null expressing Vasa-EGFP. Vasa-EGFP (green) and DAPI (blue). Yellow stars indicate the anterior tip of normal ovarioles, while orange and cyan stars indicate the anterior tips of tumorous and germless ovarioles, respectively, in Figure 1. Higher-magnification images of the germarium regions are shown in (B). Scale bars: 10 μm for (A) and 50 μm for (B). (C) Confocal images of control (sakuranull/+) and sakuranull/null ovaries stained with anti-Hts antibody to label spectrosomes and fusomes. Hts (red) and DAPI (blue). Yellow, orange, and cyan dotted lines mark the normal, tumorous, germless and germaria, respectively. Scale bars: 10 μm. (D) Ratio (%) of normal, germless, and tumorous ovarioles of indicated genotypes (ages 2-5 days; n=33, 25, 74, 29 respectively). (E) Quantification of GSC-like cell number per germarium in the indicated genotypes (ages 2-5 days; n = 33, 30, 35, and 31 respectively). (F) Confocal images of control (sakuranull/+) and sakuranull/null ovaries stained with anti-cleaved Caspase-3 antibody. Cleaved caspase-3 (red) and DAPI (blue). Scale bars: 20 μm.

Within the same ovary, some sakuranull/null ovarioles exhibited the tumorous phenotype, while others were germless, without any spectrosome or fusome staining (Fig 3A-3C, cyan stars). Quantification revealed that 35% of sakuranull/null ovarioles from 2- to 5-day-old flies were tumorous, while the remaining 65% were germless (n=74) (Fig 3D). In contrast, all ovarioles in control (sakura+/+ and sakuranull/+) and sakura-EGFP rescue flies were normal with no tumorous or germless phenotypes observed (Fig 3D). Approximately 95% of sakuranull/null ovarioles containing germ cells had an excess of GSC-like cells (>5) (Fig 3E and Fig S4). The mean number of GSCs or GSC-like cells in 2- to 5-day-old sakura+/+, sakuranull/+, and sakura-EGFP rescue was 2.2 ± 0.6, 2.2 ± 0.6, and 2.1 ± 0.7, respectively while that for sakuranull/null was 26.5 ± 17.5 (p-value <0.05).

We investigated the age-dependent effects on the tumorous and germless phenotypes by examining the ovarioles from 0-1 day old, 7-day old, and 14-day old flies. Ovarioles in control and sakura-EGFP rescue flies remained normal throughout this time course (Fig S4A). In the very young, 0-1 day old sakuranull/null, 78% of ovarioles were tumorous and 22% were germless. Overtime, the ratio of germless ovarioles increased while the ratio of tumorous ovarioles decreased. By 14 days old, only 5% of ovarioles were tumorous, with the remaining 95% being germless. The number of GSC-like cells in sakuranull/null ovarioles was already high in 0-1 day old flies, and it decreased over time (Fig S4B). Tumorous ovarioles, but not germless ones, exhibited significantly higher levels of cleaved Caspase-3 staining, indicative of apoptosis, compared to controls (Fig 3F). These results suggest that tumorous ovarioles undergo apoptosis, and that these ovarioles eventually become germless.

Together, these results suggest that Sakura is essential for regulating the survival, proliferation, and differentiation of germline cells, including GSCs.

Loss of sakura results in reduced germline piRNA

In controls, Vasa-EGFP is enriched in the perinuclear structure known as the nuage within nurse cells (Fig 3A and 3B), which is essential for piwi-interacting RNA (piRNA) biogenesis. In sakuranull/nulltumorous ovarioles, Vasa-EGFP retains its perinuclear localization, indicating that sakura is not required for the proper localization of Vasa to the nuage.

piRNAs are produced in germline cells and somatic follicle cells and transcriptionally and post-transcriptionally silence transposon expression through sequence-complementarity (Huang et al. 2017; Yamashiro and Siomi 2018). Disruption of the piRNA pathway can result in oogenesis arrest, germ cell loss, rudimentary ovaries, and sterility. Loss of piRNAs in germ cells leads to the upregulation (desilencing) of transposons, increased DNA damage, and apoptotic cell death31. We investigated whether the loss of sakura would result in loss of piRNA and transposon upregulation. We performed high-throughput sequencing (small RNA-seq and polyA+ RNA-seq) on ovary RNA samples from sakura+/+, sakuranull/+, sakuranull/null, and sakura-EGFP rescue (sakuranull/null with sakura-EGFP) flies.

piRNA levels in sakuranull/null ovaries were reduced compared to controls (sakura+/+), while sakuranull/+ and sakura-EGFP rescue showed no such reduction (Fig 4A). Consistent with this, transposon RNA levels in sakuranull/null ovaries were upregulated compared to sakura+/+, while sakuranull/+ and sakura-EGFP rescue showed no such upregulation (Fig 4B). Given that Sakura is specifically expressed in germline cells, but not in somatic cells (Fig 1E and 1F), we hypothesized that loss of sakura interrupts the piRNA pathway in germ cells, leading to the desilencing of germline transposons. A well-characterized germline transposon in Drosophila is Burdock (Donertas et al. 2013; Handler et al. 2013). We found that Burdock piRNA levels were significantly lower in sakuranull/null compared to controls and rescue flies (Fig 4C).

Loss of sakura results in lower piRNA levels and loss of piRNA-mediated transposon silencing

(A) Scatter plots of normalized number of transposon piRNA reads from small RNA-seq of indicated genotypes compared with sakura+/+. Means of three biological replicates are plotted. Downregulated (fold-change < 0.5) and upregulated (fold-change > 2) transposon piRNAs are shown in cyan and magenta, respectively. (B) Volcano plots of transposon RNAs from RNA-seq of indicated genotypes compared with sakura+/+. Three biological replicates per genotype were analyzed. Downregulated (adjusted p-value < 0.001 and log2(fold-change) < -2) and upregulated (adjusted p-value < 0.001 and log2(fold-change) > 2) transposons are shown in cyan and magenta, respectively. (C) Normalized number of reads for Burdock piRNAs from small RNA-seq. Mean ± SD (n = 3.) P-value < 0.001 (Student’s t-test, unpaired, two-tailed) is indicated by ***. (D) The Burdock sensor harbors a nanos promoter, a nuclear localization signal (NLS) appended to GFP and β-gal coding sequences, and a target sequence for Burdock piRNAs in the 3’UTR. Confocal images of ovaries from control (yRNAi) and sakuraRNAi flies harboring the Burdock sensor, where RNAi knockdown was specifically driven in the female germline with UAS-Dcr2 and NGT-Gal4. GFP (green), β-gal (red), and DAPI (blue). Scale bars: 50 μm. Three out of 3 tested control samples were negative for both GFP and β-gal, while 9 out of 9 tested sakuraRNAi samples were positive for both markers.

We employed the Burdock sensor, a transposon reporter tool to monitor germline piRNA activity (Handler et al. 2013). The reporter expresses nuclear GFP and β-gal under the control of the nanos promoter for germline expression, with a target sequence for Burdock piRNAs in the 3′ UTR (Fig 4D). Using a sakura RNAi line driven by UAS-Dcr2 and NGT-Gal4 to knock down sakura specifically in the germline, we observed that GFP and β-gal expression were highly elevated in sakura knockdown germlines compared to control RNAi (yRNAi) knockdowns (Fig 4D). This confirmed that loss of sakura leads to a loss of piRNA-mediated transposon silencing in the germline, suggesting that Sakura is essential for proper piRNA levels and piRNA-based transposon silencing in the germline.

Sex-specific sxl mRNA alternative splicing is dysregulated in sakuranull/null ovaries

Sxl is a master regulator of sex determination in Drosophila (Penalva and Sanchez 2003; Salz and Erickson 2010; Grmai et al. 2022). Sex-specific alternative splicing of sxl transcripts produces distinct mRNA isoforms: the female-specific and male-specific mRNA isoforms in respective sex and only the female-specific mRNA isoform encodes the functional Sxl protein, whereas the male-specific isoform does not. In the female germline, loss of Sxl function or disruption of female-specific sxl splicing leads to developmental defects, including germline tumors and sterility. Notably, several mutants with tumorous ovariole phenotypes—including otu mutants—exhibit aberrant sxl mRNA splicing, resulting in the expression of the male-specific isoform in ovaries and defects in germ cell sexual identity (Bopp et al. 1993; Pauli et al. 1993).

We examined sxl splicing in sakura mutants. As expected, ovaries from control and sakura-EGFP rescue flies expressed exclusively the female-specific sxl mRNA isoform and testes from control flies expressed only the male-specific isoform (Fig S5). In contrast, sakuranull/nullovaries exhibited the male-specific isoform and a reduced level of the female-specific isoform. These results indicate that female-specific sxl alternative splicing is disrupted in the absence of sakura.

sakura is important for oogenesis in germline cells beyond GSCs and germline cysts

Both sakuranull/null and sakura germline RNAi knockdown driven by UAS-Dcr-2 and NGT-Gal4, which initiates RNAi in germline from GSCs, resulted in rudimentary ovaries (Fig 2C and 5A). These ovaries lacked later-stage germline cells and egg chambers, making it difficult to assess the role of Sakura beyond the germarium. However, because Sakura-EGFP expression is not limited to GSCs and germline cysts (Fig 1E and 1F), we speculated that Sakura might function in later-stage germline cells as well.

sakura is important for oogenesis in later-stage egg chambers

(A) Stereomicroscope images of dissected whole ovaries. Scale bar: 500 μm. (B) Western blot of dissected Ovary lysates. sakuraRNAi #1 and sakuraRNAi #2 are two independent RNAi lines. (C) Number of eggs laid by sakura RNAi knockdown driven by UAS-Dcr2 and TOsk-Gal4. Mean ± SD (n = 5). P-value < 0.001 (Student’s t-test, unpaired, two-tailed) is indicated by ***. (D) Number of stage 14 oocytes in ovaries of sakura RNAi knockdown flies driven by UAS-Dcr2 and TOsk-Gal4. The numbers of stage 14 oocytes per fly (per a pair of ovaries) are shown. n is the numbers of flies examined. P-value < 0.001 (Student’s t-test, unpaired, two-tailed) is indicated by ***. (E) Confocal images of the ovaries from sakura RNAi knockdown flies driven by UAS-Dcr2 and TOsk-Gal4, stained with anti-Orb and anti-Vasa antibodies. Orb (green), Vasa (red), and DAPI (blue). Yellow arrows label the normal enrichment of Orb in the developing oocytes. Magenta arrow labels mislocalized developing oocyte. White arrowheads label egg chambers with cytoskeletal disorganization. Scale bars: 100 μm. In A-E, y-RNAi was used as a control.

To explore this possibility, we used TOsk-Gal4 (a combination of osk-Gal4 and αTub67C-Gal4) to drive sakura RNAi knockdown in germline cells after the germline cyst stage (from germarium region 2b onward. Fig 1A), sparing GSCs and germline cysts (ElMaghraby et al. 2022). Unlike NGT-Gal4, TOsk-Gal4-driven sakura RNAi knockdown did not drastically affect ovary morphology (Fig 5A), allowing us to study the effects of sakura loss in egg chambers. We confirmed sakura RNAi (sakura RNAi #1 and #2) knockdown efficiency by Western blot, showing effective depletion of Sakura in ovaries from both NGT-Gal4 and TOsk-Gal4-driven RNAi lines (Fig 5B).

Interestingly, TOsk-Gal4-driven sakura RNAi knockdown severely reduced the numbers of eggs laid (Fig 5C) and stage 14 oocytes in ovaries compared to control RNAi (yRNAi) (Fig 5D), suggesting that sakura is important for oogenesis beyond the germline cyst stage. Additionally, we observed mislocalized and dispersed Oo18 RNA-binding protein (Orb), a marker for oocyte identity, in TOsk-Gal4-driven sakura RNAi ovaries beginning around stage 6 of oogenesis (Fig 5E). In controls, Orb was enriched and properly localized to the posterior end of stage ∼6-8 egg chambers (Fig 5E, yellow arrows). In sakuraRNAi ovaries, although Orb localization appeared normal at stage ∼6 (yellow arrow), mislocalization was evident by stage ∼8 (magenta arrow), and signs of cytoskeletal disorganization were apparent in later-stage egg chambers (white arrowheads). Phalloidin staining to visualize F-actin further supported cytoskeletal defects in sakuraRNAi later-stage egg chambers (Fig S6).

These defects likely contribute to the impaired production of stage 14 oocytes and eggs in sakuraRNAi flies (Fig 5C and 5D). We conclude that Sakura plays an essential role in oogenesis beyond the early germline stages. For all subsequent sakura RNAi experiments, we used the sakura RNAi #2 line.

Sakura is required intrinsically for GSC establishment, maintenance, and differentiation

To investigate whether sakura functions autonomously in the germline, we performed mosaic analysis using the FLP-FRT system with heat shock promoter-driven FLP (hs-flp) (Xu and Rubin 1993; Rubin and Huynh 2015). First, to assesse its role in GSC establishment, we induced sakuranull clones in primordial germ cells (PGCs) by applying heat shock before the early third instar larval stage and tracked their development into adult GSCs, following a previously established protocol (Yang et al. 2007). In control adult flies (FRT82B),12.3% (20/163) of GSCs were marked, wheras only 1.9% (4/213) were marked in sakuranullmutants (FRT82B, sakuranull) (p-value = 4.8 * 10-11, chi-square test), indicating that sakura is autonomously required for GSC establishment.

Next, we assessed GSC maintenance using the same system by generating sakuranull GSC clones in adult flies and measuring clone loss over time, as described previously (Xie and Spradling 1998; Yang et al. 2007). Four days after clone induction, 27.8% of GSCs were marked in controls (FRT82B) and 22.1% in sakuranull (FRT82B, sakuranull), which we defined as initial levels (Fig 6A). In controls, the percentage of marked GSCs declined to 25.8% on day 7 and and 16.8% on 14, resulting in a 39.4% loss rate over 10 days (from day 4 to day 14). In contrast, the proportion of marked sakuranull GSC clones dropped more sharply—to 16.3% on day 7 and and 5.4% on day 14—yielding a higher loss rate of 75.6% over 10 days. These results demonstrate that sakura is intrinsically important for GSC maintenance.

Germline clonal analysis of sakuranull

(A) Percentage of germaria with marked GSC clones indicated by the absence of GFP at 4, 7, and 14 days after clone induction at the adult stage. Arrows indicate the percent decrease of marked GSC clones compared to day 4. The genotype, actual percentage, and total number of germaria examined are shown in the adjacent table. P-value < 0.01 (Chi-squared test) is indicated by **. (B) Confocal images of Germ cell clones. sakuranull and control clones were marked by the absence of GFP. GFP (green), Hts (red), and DAPI (blue). Scale bars: 10 μm. (C) Number of marked (GFP-negative) GSC-like cells in germaria with marked GSCs of the indicated genotypes at 4, 7, and 14 days after clone induction. GSC-like cells containing round spectrosome were identified through immunostaining with anti-Hts antibody. P-value < 0.05 and <0.001 (Student’s t-test, unpaired, two-tailed) is indicated by * and ***, respectively.

We further tested whether sakura is intrinsically required for GSC differentiation, we quantified the number of marked (GFP-negative) and unmarked (GFP-positive) GSC-like cells (germline cells with a round spectrosome) in the germaria containing marked GSC clones at days 4, 7, and 14 post-induction (Fig 6B). We found that marked sakuranull GSC-like cells were significantly more numerous than marked control GSC-like cells at all time points (Fig 6C). The number of marked sakuranullGSC-like cells increased over the 10-day period, while the number of marked control cells did not (Fig 6C). In contrast, unmarked GSC-like cells in germaria containing sakuranullGSC clones did not differ significantly in number compared with those in germaria containing marked control GSC clones, and the number did not increase over time (Fig S8).

These results demonstrate that sakura is intrinsically important in germline cells, including GSCs, to regulate proper division and differentiation. In the absence of intrinsic sakura, germline cells become tumorous and undergo uncontrolled proliferation. Taken together, we conclude that sakura is required intrinsically for GSC establishment, maintenance, and differentiation.

Loss of sakura inhibits Dpp/BMP signaling

The Dpp/BMP signaling pathway plays a central role in regulating bam expression and GSC self-renewal and differentiation (Fig 7A) (Kirilly and Xie 2007; Hayashi et al. 2020). We hypothesized that the germless and tumorous phenotypes observed in sakura loss-of-function ovaries could be due to Dpp/BMP signaling misregulation. To test this, we knocked down sakura in the germline driven by UAS-Dcr2 and NGT-Gal4 in the presence of bam-GFP reporter (Chen and McKearin 2003b). In controls (yRNAi), Bam-GFP expression was restricted to 8-cell cysts and disappeared in 16-cell cysts and onward (Fig 7B). In contrast, sakura knockdowns ovaries showed persistent Bam-GFP expression throughout the germarium including the GSC niche region (Fig 7B).

Loss of sakura inhibits Dpp/BMP signaling

(A) Schematic illustration of bam expression regulation. Cap cells secrete diffusible Decapentaplegic (Dpp), which is received by its receptor, a heterodimer of Thick vein (Tkv) and Punt (Put), in GSCs. The activated Dpp signaling eventually phosphorylates Mother-against-dpp (Mad). The phosphorylated Mad (pMad) represses the transcription of bag-of-marbles (bam). The repression of bam in GSCs is crucial for maintaining their stemness. Cystoblasts do not receive Dpp, and Bam expression is crucial for promoting cystoblast differentiation from GSCs. (B) Confocal images of ovaries from control (yRNAi) and sakuraRNAi flies harboring the bam-GFP reporter, where RNAi knockdown was specifically driven in the female germline with UAS-Dcr2 and NGT-Gal4. Bam-GFP (green), Hts (red), and DAPI (blue). Germarium are outlined by dotted line. Scale bars: 20 μm. (C, D) Confocal images of germaria with germline clones of sakuranullstained with (C) anti-Bam antibody and (D) anti-pMad antibody. GFP (green), Bam or pMad (red), and DAPI (blue). Scale bars: 10 μm. (E) Mean pMad intensity in the germline clones of the indicated genotypes. Mean ± SD (n = 7). P-value < 0.001 (Student’s t-test, unpaired, two-tailed) is indicated by *** (F) Confocal images of germaria with germline clones of sakuranull stained with anti-CycA antibody. GFP (green), CycA (red), and DAPI (blue). Scale bars: 10 μm. (G) Mean CycA intensity in the germline clones of the indicated genotypes. Mean ± SD (n = 5 and 10 for FRT82B and FRT82B, sakuranull respectively). P-value < 0.01 (Student’s t-test, unpaired, two-tailed) is indicated by **. In C, D, and F, clones were marked with the absence of GFP.

To further confirm this finding, we used the FLP-FRT system to induce sakuranull clones in germaria and performed immunostaining with anti-Bam antibody. In control clones (FRT82B), Bam expression was observed exclusively in 8-cell cysts. However, in all sakuranull clones (FRT82B, sakuranull), Bam was aberrantly expressed throughout the germarium, including in GSCs (Fig 7C). These results suggest that in the absence of sakura, Bam expression is no longer repressed by Dpp/BMP signaling in the GSC niche, leading to GSC loss, and is no longer shut off after the 16-cell cyst stage.

In GSCs, pMad translocates into the nucleus to repress Bam expression (Fig 7A) (Kai and Spradling 2003; Song et al. 2004). We stained ovaries with anti-pMad antibodies and found that pMad intensity was significantly reduced in sakuranull GSC clones compared to control GSCs (Fig 7D and E). This indicates that the dysregulated Bam expression observed in sakuranull was due to reduced pMad levels in GSCs.

Previous studies have shown that ectopic expression of a stable form of CycA leads to germ cell loss (Chen et al. 2009). This germ cell loss phenotype is also observed upon ectopic bam expression in GSCs (Xie and Spradling 1998; Chen and McKearin 2003a; Xia et al. 2010). It was reported that Bam associates with Ovarian tumor (Otu) to promote deubiquitination and stabilization of CycA (Ji et al. 2017). Since we observed derepressed bam expression in sakuranullcells, we investigated whether CycA levels were affected. Staining ovaries with anti-CycA antibodies revealed higher CycA intensity in sakuranull clones compared to neighboring wild-type cells in the same germarium (FRT82B, sakuranull) and control clones in control germarium (FRT82B) (Fig 7F). sakuranull clones exhibited significantly higher mean CycA intensity compared to control clones (Fig 7G). The elevated CycA levels in sakuranull clones likely results from Bam misexpression and Bam-mediated stabilization.

To determine whether bam misexpression occurs in egg chambers in sakura knockdown conditions, we examined bam expression in TOsk-Gal4-driven sakura RNAi flies. Bam was appropriately restricted to 8-cell cysts in germaria and not misexpressed in egg chambers compared to yRNAi controls (Fig. S8, cyan arrowheads). Thus, sakura is specifically required in GSCs and cyst cells to regulate bam expression.

We also examined DE-Cadherin (DE-Cad), which plays a role in oocyte positioning within the egg chamber (Godt and Tepass 1998). DE-Cad level and localization appeared normal in both control and sakura RNAi ovaries through approximately stage 8. However, later-stage egg chambers in sakura RNAi flies displayed structural disorganization (Fig. S8, white arrowheads), consistent with defects observed in Fig. 5E and Fig. S6.

Attempts to rescue sakura loss-of-function ovariole phenotypes

To determine whether the phenotypes caused by sakura loss-of-function could be rescued, we first performed genetic interaction experiments by knocking down either bam or cycA in the germline. As expected, germline-specific bam RNAi driven by UAS-Dcr2 and NGT-Gal4 resulted in 100% tumorous ovarioles, while cycA RNAi produced 100% normal ovarioles in 2–5-day-old flies (Fig. S9A). Double knockdown of sakura and bam reduced the proportion of germless ovarioles and increased the proportion of tumorous ovarioles compared to sakura RNAi alone or the double knockdown of sakura and control w RNAi. However, no normal ovarioles were observed (Fig. S9A). sakura and cycA double knockdown had no effect on the phenotype distribution, suggesting that the germless phenotype in sakura mutants is partially attributable to ectopic bam expression but not to cycA misregulation.

The number of GSC-like cells was increased in sakura and bam double knockdowns compared to sakura RNAi or sakura and w double RNAi, whereas sakura and cycA knockdown did not alter GSC-like cell numbers (Fig. S9B), further supporting a genetic interaction between sakura and bam.

Next, we tested whether the sakura loss-of-function phenotypes could be rescued by overexpressing components of the Dpp/BMP signaling pathway. Germline expression of UASp-Mad-GFP driven by nos-Gal4-VP16 in both sakuranull/+ controls and sakuranull/null did not alter the proportions of ovariole phenotypes (Fig S10A) or the number of GSC-like cells (Fig S10B). Thus, transgenic overexpression of Mad did not rescue the sakura mutant phenotypes.

We then tested whether overexpression of a constitutively active form of the Dpp receptor Thickveins (Tkv.Q253D) (Casanueva and Ferguson 2004) could rescue the phenotypes. Expression of UASp-tkv.Q253D in the germline using a NGT-Gal4 in a control (yRNAi) background resulted in 100% tumorous ovarioles, as expected (Fig. S11A). When co-expressed with sakura RNAi, Tkv.Q253D increased the proportion of tumorous ovarioles and decreased the proportion of germless ovarioles relative to sakuraRNAi alone but did not restore normal ovarioles. The number of GSC-like cells was also increased by Tkv.Q253D expression in sakuraRNAi (Fig S11B), indicating partial phenotypic modulation via BMP pathway activation.

Sakura binds Otu

To explore the molecular function of Sakura in oogenesis, we aimed to identify Sakura-interacting proteins. We conducted a co-immunoprecipitation experiment using anti-GFP magnetic beads on the ovary lysates of flies expressing Sakura-EGFP, followed by mass spectrometry to identify co-immunoprecipitated proteins. Ovary lysates from flies without the sakura-EGFP transgene (w1118) served as negative controls. Mass spectrometric analysis revealed several proteins with peptide signals present in all three biological replicates of the Sakura-EGFP samples without any signals in any of the three biological replicates of the negative control (Table 1). Among these, the most promising candidate, exhibiting the highest unique peptide signal in each of three replicates of the Sakura-EGFP samples was Otu.

To confirm this finding, we generated a polyclonal anti-Otu antibody and validated it by Western blots using ovary lysates from the hypomorphic otu14 mutant (Fig S12A) (King et al. 1986; Storto and King 1987; Steinhauer and Kalfayan 1992a; Pauli et al. 1993). Using this antibody, we detected endogenous Otu protein in the Sakura-EGFP immunoprecipitants from ovaries of 3–7-day-old flies, but not in the negative control (Fig 8A), validating the mass spectrometry results. Next, using wild-type (without Sakura-EGFP) ovary lysates from 3–7-day old-flies, we immunoprecipitated endogenous Sakura with anti-Sakura antibody and detected endogenous Otu in the immunoprecipitant by Western blot (Fig 8B), demonstrating that endogenous Sakura interacts with endogenous Otu. Furthermore, we found that ectopically expressed Flag-tagged Otu co-immunoprecipitated with ectopically expressed HA-tagged Sakura in S2 cells (Fig 8C).

Sakura interacts with Otu.

(A) Co-immunoprecipitation using anti-GFP magnetic beads followed by Western blotting. Ovary lysates expressing Sakura-EGFP in sakura+/+ background and those from w1118 negative control were tested. (B) Co-immunoprecipitation using beads bound with rabbit anti-Sakura followed by Western blotting. Ovary lysates from w1118 flies were used. Rabbit IgG was used as control IP. (C) Co-immunoprecipitation using beads bound with anti-FLAG antibody followed by Western blotting. S2 cell lysates expressing 3xFLAG-Otu and Sakura-mCherry-3xHA or mCherry-3xHA (negative control) were used.

The otu gene encodes two annotated protein isoforms, a predominant 98 kDa isoform and a less abundant 104 kDa isoform generated by alternative splicing (Steinhauer and Kalfayan 1992b). The 98 kDa isoform lacks the Tudor domain, which is encoded by an alternatively spliced 126-nucleotide exon (Fig 9B. Otu(ΔTudor)) (Van Buskirk and Schupbach 2002). Previous studies have shown that the 104 kDa Otu isoform is more abundant in predifferentiated germline cells and is sufficient to carry all known Otu functions, whereas the 98 kDa Otu isoform becomes prominent during later stages of oogenesis and is implicated in nurse cell regression and oocyte maturation (Sass et al. 1995). Consistent with these findings, we observed that ovaries from very young (2–5 hour old) flies—containing only germaria and previtellogenic egg chambers—predominantly expressed the 104 kDa isoform (Fig S12B). In contrast, ovaries from 3–7 day old flies—containing all 14 stages of oogenesis—predominantly expressed the 98 kDa isoform. Importantly, both the 104 kDa and 98 kDa isoforms co-immunoprecipitated with endogenous Sakura using anti-Sakura antibody, demonstrating that Sakura interacts with both isoforms of Otu and that the Tudor domain is not required for this interaction (Fig. S12B).

N-terminal regions of Sakura and Otu are important for interaction.

(A) Predicted structure of the Sakura and Otu protein complex made by AlphaFold. Full-length Sakura and N-terminal Otu fragment (N1. 1-405aa) were used for prediction. (B) Full-length Otu and Otu fragments tested in co-immunoprecipitation assays. The binding assay results from (D) is summarized. (C) Full-length Sakura and Sakura fragments tested in co-immunoprecipitation assays (N: N-terminal, M: middle, C: C-terminal). The binding assay results from (E) is summarized. (D, E, F) Co-immunoprecipitation using anti-HA magnetic beads followed by Western blotting. S2 cell lysates expressing HA-tagged mCherry were used as negative controls.

Next, we sought to identify which regions of Sakura and Otu are important for their interaction. We predicted the structure of the Sakura-Otu protein complex using AlphaFold (Jumper et al. 2021; Yang et al. 2023). AlphaFold suggested that the N-terminal region of Sakura (1-49aa) and the N-terminal region of Otu (1-405aa) are highly likely to form interactions, while the C-terminal region of Sakura (100-144aa) does not interact with Otu (Fig 9A-C). The Tudor domain of Otu is not directly involved in this interaction, consistent with our findings (Fig S12B). To experimentally validate the predictions made by AlphaFold, we generated a series of truncated Otu fragments, including Otu(ΔTudor) (1-336, 389-853aa, corresponding to the endogenous 98 kDa isoform), Otu(N1) (1-405aa), Otu(N2) (1-339aa), and Otu(C) (405-853aa) (Fig 9B). We then performed co-immunoprecipitation assays to test which fragments could interact with full-length Sakura. All fragments except Otu(C) were associated with full-length Sakura protein in S2 cells (Fig 9D), demonstrating that the C-terminal low-complexity region of Otu is dispensable for the interaction. This assay also showed that the Tudor domain is not critical and the N-terminal fragment (1-339aa) of Otu is sufficient for interaction with Sakura. These findings align well with the predictions made by AlphaFold.

Subsequently, we created several truncated Sakura fragments and conducted co-immunoprecipitation assays in S2 cells to determine which Sakura fragments can interact with full-length Otu. The fragments generated included Sakura(NM) (1-100aa), Sakura(NC) (1-49, 100-144aa), Sakura(MC) (49-144aa), Sakura(N) (1-49aa), Sakura(M) (49-100aa), and Sakura(C) (100-144aa) (Fig 9C). We found that Sakura(NM), Sakura(NC), Sakura(N), and Sakura(M) were associated with Otu (Fig 9E). Compared to Sakura(NM) and Sakura(N), the interaction of Sakura(M) with Otu is relatively weaker, as indicated by a fainter band in the Western blot (Fig 9E). Sakura-NC exhibited a weak interaction with Otu, while Sakura(MC) and Sakura(C) showed no interaction (Fig 9E). These assays demonstrated that the N-terminal region (1-49aa) of Sakura is sufficient for interaction with Otu. Moreover, the results suggest that the C-terminal region of Sakura (100-144aa) does not interact with Otu and that adjoining it with the N-terminal (1-49aa) or the middle (M) region (49-100aa) weakens the interaction with Otu. These findings support the AlphaFold predictions.

Finally, to test whether the interaction between Sakura and Otu can be achieved solely through their N-terminal regions, we performed co-immunoprecipitation assay with Sakura(N) and Otu(N2). We discovered that Sakura(N) and Otu(N2) physically interact (Fig 9F). A reciprocal co-immunoprecipitation assay also confirmed their interaction (Fig S13), revealing that the interaction between Sakura and Otu can be established with just the 1-49aa of Sakura and the 1-339aa of Otu. Therefore, we conclude that Sakura interacts with Otu in vivo, and the N-terminal regions of both proteins are sufficient for their interaction.

Loss of otu phenocopies loss of sakura

Having established the interaction between Sakura and Otu, we were interested in whether they function together during oogenesis. We speculated that for them to interact and function together, Sakura and Otu must exhibit similar expression and localization patterns in ovaries. We generated two transgenic flies carrying full-length otu-EGFP or otu(ΔTudor)-EGFP transgenes; both constructs are C-terminally fused with EGFP and under the control of the otu promoter. We found that both Otu-EGFP and Otu(ΔTudor)-EGFP display similar expression and localization patterns to Sakura-EGFP (Fig 1E, 1F and Fig 10A). Like Sakura-EGFP, Otu-EGFP and Otu(ΔTudor)-EGFP are specifically expressed in germ cells, but not in follicle cells, and are localized to the cytoplasm, being enriched in the developing oocytes (Fig 10A). We did not observe any difference in localization patterns between Otu-EGFP and Otu(ΔTudor)-EGFP; suggesting that the Tudor domain is not crucial for Otu protein localization. Notably, the localization pattern exhibited by the transgenic Otu-EGFP generated in this study is consistent with a previous report (Glenn and Searles 2001).

Loss of otu phenocopies loss of sakura

(A) Confocal images of ovaries from otu-EGFP and otu(ΔTudor)-EGFP transgenic flies. Otu-EGFP and Otu(ΔTudor)-EGFP (green), DAPI (blue). Yellow arrows show the enrichment of Otu-EGFP and Otu(ΔTudor)-EGFP in developing oocytes. Scale bars: 20 μm. (B) Stereomicroscope images of dissected whole ovaries from wRNAi (control) and otuRNAi flies where RNAi knockdown was driven in the female germline with nos-Gal4-VP16 or TOsk-Gal4. Scale bar: 500 μm. (C) Western blot of dissected ovary lysates. (D) Number of eggs laid by TOsk-Gal4 > wRNAi and TOsk-Gal4 > otuRNAi flies. Mean ± SD (n = 5). P-value < 0.001 (Student’s t-test, unpaired, two-tailed) are indicated by ***. (E) Violin plots of the number of stage 14 oocytes produced in TOsk-Gal4 > wRNAi and TOsk-Gal4 > otuRNAi flies. n = 30. P-value < 0.001 (Student’s t-test, unpa©red, two-tailed) is indicated by ***.

Previous studies have shown that mutations in otu lead to defects in germ cell division and differentiation, resulting in phenotypes including tumorous ovarioles (King and Riley 1982; King et al. 1986; Storto and King 1988; Steinhauer and Kalfayan 1992b). To study the role of otu specifically in the germline, we performed RNAi knockdown of otu in the germline driven by nos-Gal4-VP16 (Fig 10B and 10C). Interestingly, we found that germline depletion of otu results in rudimentary ovaries, similar to the loss of sakura (Fig 2C, Fig 5A, and Fig 10B). Furthermore, similar to sakura RNAi (Fig 5), TOsk-Gal4-driven otu RNAi knockdown did not affect ovary morphology (Fig 10B), but it resulted in a reduced number of eggs laid and stage 14 oocytes compared with control RNAi (wRNAi) (Figure 10D and 10E).

Given that sakura loss of function causes tumorous and germless ovarioles, apoptosis, piRNA pathway defects, bam misexpression, and reduced pMad levels (Fig 3, 4 and 7B-E), we asked whether similar defects occur upon otu depletion. Germline otu RNAi knockdown in 2-5 days old flies using UAS-Dcr2 and NGT-Gal4 lead to both germless (∼70%) and tumorous (∼30%) ovarioles, closely resembling the phenotype of sakura RNAi (Fig S9A). Similar germless and tumorous phenotypes were also observed in nos-Gal4-VP16 > otuRNAi flies (Fig 11A). In these otuRNAi ovarioles, Bam-GFP expression was no longer restricted to 8-cell cysts, instead persisting throughout the germarium and egg chambers (Fig 11A). Additionally, otu14/14 mutant GSCs showed reduced pMad levels (Fig 11B), and cleaved Caspase-3 staining revealed elevated apoptosis in otu RNAi ovaries (Fig 11C). Germline depletion of otu resulted in elevated expression of GFP and β-Gal produced by the Burdock sensor, indicating loss of piRNA-mediated transposon silencing (Fig 11D).

Loss of otu results in low levels of pMad and derepression of bam in the germaria

(A) Confocal images of the ovaries from control (wRNAi) and otuRNAi flies harboring the bam-GFP reporter, where RNAi knockdown was driven in the female germline with nos-Gal4-VP16. Bam-GFP (green), Hts (red), and DAPI (blue). Scale bars: 20 μm. (B) Confocal images of germaria from control (otu14/+) and otu14/14mutant flies stained with anti-pMad (green) and anti-Hts (red) antibodies. Scale bars: 10 μm. (C) Confocal images of the ovaries from nos-Gal4-VP16 > wRNAi and nos-Gal4-VP16 > otuRNAi flies stained with anti-cleaved Caspase 3 antibody. Cleaved Caspase-3 (red) and DAPI (blue). Scale bars: 20 μm. (D) Confocal images of the ovaries from wRNAi and otuRNAi flies carrying the Burdock sensor, with RNAi driven in the female germline using UAS-Dcr2, NGT-Gal4 and nos-Gal4-VP16. GFP (green), β-gal (red), and DAPI (blue). Scale bars: 50 μm. Three out of 3 tested control samples were negative for both GFP and β-gal, while 14 out of 14 tested otuRNAi samples were positive for positive for GFP and β-gal.

Together, these data demonstrate that otu loss-of-function closely phenocopies sakura loss-of-function, including defects in germline maintenance and differentiation, bam misexpression, reduced pMad signaling, and compromised piRNA pathway activity. These findings, along with physical interaction between Sakura and Otu (Figs 8, 9, S12, and S13), suggest that they function together in regulating germline maintenance and differentiation.

To further explore their genetic interaction, we compared the ovariole phenotypes caused by single and double knockdowns of sakura and otu in the germline driven by UAS-Dcr2 and NGT-Gal4. Individual knockdown of either gene resulted in ∼70% germless and ∼30% tumorous ovarioles in 2-5 days old flies (Fig S9A). Simultaneous knockdown of both genes increased the propotion of germless ovarioles to 86%, with a corresponding decrease in tumorous phenotypes to 14%, suggesting a synergistic enhancement of the germless phenotype. The number of GSC-like cells were not affected by the double knockdown of sakura and otu compared with their respective single knockdown (Fig S9B).

Sakura and Otu proteins do not depend on each other for their abundance

We investigated whether Sakura and Otu proteins depend on each other for their abundance. We performed Western blot analysis on ovaries with germline-specific knockdown of sakura or otu, driven by TOsk-Gal4. Otu protein levels were not reduced in sakuraRNAi ovaries compared to control yRNAi (Fig S14A), and conversely, Sakura levels were not decreased in otuRNAi ovaries compared with control wRNAi ovaries (Fig S14B). These results indicate that Sakura and Otu do not require each other for their protein expression or stability.

To further test this conclusion and to examine whether Otu enrichment to the posterior of egg chambers is affected by loss of sakura, we generated marked sakuranull germline clones using the FLP-FRT system. Clones were marked by the absence of RFP, and experiments were performed in the flies expressing the otu-EGFP or otu(ΔTudor)-EGFP transgene in an otherwise wild-type (otu+/+) background. Otu-EGFP fluorescence signal was comparable between sakuranullclones (FRT82B, sakuranull) and control clones (FRT82B) (Fig 12A). Similarly, Otu(ΔTudor)-EGFP levels were unaffected in sakuranull clones compared to the control clones (Fig 12B), confirming that Sakura is not required for Otu protein expression or stability.

Otu enrichment to the posterior within egg chambers is lost in sakuranull

Confocal images of egg chambers with germline clones of sakuranull expressing (A) Otu-EGFP or (A) (B) Otu(ΔTudor)-EGFP. Fly genotypes used: hs-flp/w; otu-EGFP/+; FRT82B, ubi-RFP/FRT82B. hs-flp/w; otu-EGFP/+; FRT82B, ubi-RFP/FRT82B, sakuranull. hs-flp/w; otu(ΔTudor)-EGFP/+; FRT82B, ubi-RFP/FRT82B. hs-flp/w; otu(ΔTudor)-EGFP/+; FRT82B, ubi-RFP/FRT82B, sakuranull. RFP (red), Otu-EGFP or Otu(ΔTudor)-EGFP (green), and DAPI (blue). Scale bars: 20 μm. Marked clones (RFP-negative) are outlined with white dotted lines. Yellow arrows indicate normal posterior enrichment of Otu-EGFP and Otu(ΔTudor)-EGFP signal; magenta arrows indicate the loss of this posterior enrichment in sakuranullclones.

Both Otu-EGFP and Otu(ΔTudor)-EGFP exhibited clear enrichment at the posterior of egg chambers in the control clones (FRT82B) —consistent with localization to the developing oocyte (Fig 12A and 12B). This posterior enrichment was absent in sakuranullclones, demonstrating that Sakura is crucial for this process.

Sakura does not affect Otu’s deubiquitinase activity in vitro

A previous study has shown that Otu possesses deubiquitinase activity (Ji et al. 2017). We asked whether Sakura regulates the Otu’s deubiquitinase activity. To test this, we performed an in vitro deubiquitination assay using Ub-Rhodamine 110 as a model substrate. Consistent with the previous study (Ji et al. 2017), we detected a deubiquitinase activity in Otu that was ectopically expressed and purified from S2 cells (Fig S15). The presence of Sakura, purified from E.coli, did not affect Otu’s deubiquitinase activity under these assay conditions. Additionally, Sakura itself did not exhibit deubiquitinase activity.

Discussion

In this study, we identified Sakura, encoded by a previously uncharacterized gene CG14545, as an essential factor for oogenesis and female fertility. We demonstrated that Sakura is specifically expressed in germline cells in the ovary, including GSCs, is localized to the cytoplasm, and is enriched in the developing oocytes. sakura homozygous null mutant flies are viable but completely female-sterile and male-fertile. Loss of sakura, either through null mutation or germline RNAi including in GSCs, results in rudimentary ovaries exhibiting germless and tumorous phenotypes.

The tumorous phenotype associated with the loss of sakura is characterized by an excess of GSC-like cells, which feature round spectrosomes, as well as cyst cells with branched fusomes (Fig 3C). The increased number of GSC-like cells in loss of sakura suggests dysregulation of GSC self-renewal and differentiation. Meanwhile, the presence of excess cyst cells with branched fusomes indicates abnormal differentiation and division of cysts. Differentiation of cystoblasts typically involves four mitotic divisions with incomplete cytokinesis, leading to 16 cyst cells interconnected by branched fusomes. The degree of fusome branching serves as a marker for the stages of cyst cell division, with increased branching from 2-cell to 16-cell cysts (de Cuevas and Spradling 1998). Notably, fusomes begin to degenerate and disappear after 16-cell cysts as the germline cyst enters the meiotic zone or region 2 of the germarium. The persistence of cyst cells with branched fusomes in sakuranull/nullsuggests that sakura is crucial for the proper division and differentiation of cysts, and we speculate that it is required for germline cysts to enter meiotic division.

Mosaic analysis of sakuranull indicates that sakura is intrinsically required for the establishment of GSCs in the ovary. When mutant clones were induced in PGC stage, significantly fewer sakuranull marked GSCs were observed in adult ovaries, suggesting that many sakuranull PGCs fail to survive or differentiate into GSCs. Furthermore, induction of sakuranull clones in adult ovaries led to a more rapid decline in marked sakuranull GSCs compared to controls, indicating that sakura is also intrinsically required for GSC maintenance (Fig 6A). Over time, germaria containing sakuranull GSC clones became increasingly tumorous, with an expanding population of sakuranull GSC-like cells (Fig 6B and 6C). In contrast, the number of GSC-like cells across entire and all ovarioles regardless of whether GSCs in the niche remain, declined over time in sakuranull ovaries (Fig S4B). These results suggest that sakuranull GSCs initially undergo aberrant proliferation, leading to tumor formation, but are ultimately sakuranullGSCs and GSC-like cells are lost through a cell death mechanism, as evidenced by the elevated levels of cleaved Caspase-3 in sakuranull ovaries (Fig 3F).

Dpp/BMP signaling governs GSC self-renewal and cystoblast differentiation by repressing bam in GSCs and de-repressing it in daughter cystoblasts (Fig 7A). This process is mediated by the transcription factor Mad, which, when phosphorylated (pMad), translocates into the nucleus to repress bam transcription (Kirilly and Xie 2007; Kahney et al. 2019; Hinnant et al. 2020). Our findings indicate that Bam expression is not limited to 8-cell cysts but continues throughout the germarium in ovaries lacking sakura function (Fig 7B and 7C). The misexpression of Bam in GSCs likely results from reduced levels of pMad in the GSCs (Fig 7D, and 7E). Given the low pMad levels observed upon loss of sakura, we propose that bam misregulation in GSCs occurs primarily at the transcriptional level, although Bam is also known to be subject to post-transcriptional regulation (Pek et al. 2009). The continued expression of Bam throughout the germarium, well beyond the GSC niche, suggests a failure to terminate bam expression at the 16-cell cyst stage —a developmental transition when bam is normally turned off. Thus, sakura mutants likely exhibit two distinct molecular defects contributing bam overexpression: (1) impaired Dpp/BMP signaling within the GSC niche, and (2) a failure to downregulate bam at 16-cell stage. Given the limited space within the GSC niche, the latter defect may be the predominant contributor to the observed bam overexpression. The molecular mechanism that silences bam at the 16-cell cyst stage remains poorly understood, and how sakura loss perturbs this regulation is unclear. The sakura mutant may therefore serve as a valuable model to investigate the mechanism that normally suppresses bam expression at the 16-cell stage.

Transposons are mobile genetic elements that, if not silenced, can generate DNA damage and genomic instability (Levin and Moran 2011). piRNAs derived from transposons and other repeats can target and silence transposon RNAs to preserve genome integrity in germ cells (Siomi et al. 2011). Loss of piRNAs leads to transposon derepression, resulting in increased DNA damage, which subsequently triggers cell death (Kang et al. 2018; Moon et al. 2018). Genetic damage in germ cells can cause developmental defects and diseases that may be inherited by the next generation. Thus, elimination of defective germ cells is crucial for maintaining germline integrity of a species (Chu et al. 2014; Ota and Kobayashi 2020). We found that loss of sakura results in reduced piRNA levels and loss of piRNA-mediated transposon silencing in the germline (Fig 4). The observed apoptosis, indicated by elevated cleaved Caspase-3 levels in sakura mutant ovaries (Fig 3F), suggest that desilencing of transposon due to reduced piRNA levels likely results in increased DNA damage, triggering cell death. We speculate that the germless phenotype in sakura mutants may partly arise from an apoptotic germline elimination program activated to maintain germline integrity.

In addition to its expression in GSCs and cysts in the germarium, Sakura is also expressed in germline cells in later-stage egg chambers (Fig 1E-F) and is required at this stage for proper oogenesis. When Sakura is depleted from region 2b of the germarium onward —leaving GSCs and early cyst cells intact—females have significantly fewer stage 14 oocytes and lay significantly fewer eggs (Fig 5). The failure to produce stage 14 mature oocytes likely stems from cytoskeletal disorganization that disrupt oocyte development, as evidenced by mislocalization of Orb (Fig 5E, S7, and S12). Notably, piRNA pathway mutants also exhibit similar Orb mislocalization (Ohtani et al. 2013). Therefore, the reduced piRNA levels (Fig 4) may contribute to the oocyte development defects, including Orb mislocalization, in sakura mutants. (Fig 5E).

Sakura does not possess any known protein domains. To infer its function, we identified Otu, a protein known to be crucial for oogenesis, as a protein partner of Sakura. Mutations in otu gene lead to a range of ovarian phenotypes, including germ cell loss, tumorous egg chambers filled with undifferentiated germ cells, defects in germlne sexual identity determination, abnormalities in nurse cell chromosome structure, and defects in oocyte determination (King and Riley 1982; Storto and King 1988; Pauli et al. 1993; Glenn and Searles 2001). We showed that germline depletion of Otu via RNAi phenocopies loss of sakura. Similar to Sakura, loss of otu inhibits Dpp/BMP signaling, resulting in low pMad levels in GSCs and Bam de-repression (Fig 11A-B). Additionally, loss of otu also results in the loss of piRNA-mediated silencing, paralleling the effects of sakura loss. Furthermore, germline knockdown of otu yielded a similar ratio of germless ovaries as seen in sakura-RNAi, and simultaneous knockdown of both otu and sakura exacerbated the germless ovary phenotype (Fig S9). These observations raise the possibility that Sakura and Otu function together to regulate germ cell maintenance in the ovaries and are involved in Dpp/BMP signaling to balance stem cell renewal and differentiation.

Otu possesses deubiquitinase activity, catalyzed by its N-terminal Otu domain (Ji et al. 2017; Ji et al. 2019). In germ cells, Otu interacts with Bam, forming a deubiquitinase complex that deubiquitinates and thereby stabilizes CycA, promoting GSC differentiation. The predicted structure of Sakura and Otu complex suggests that the Mid domain of Sakura directly contacts the Otu domain of Otu (Fig 9A). We found that Sakura lacks a deubiquitinase activity and it does not directly affect Otu’s deubiquitinase rate in our in vitro assays using Ub-Rhodamine 110 as a substrate (Fig S15). This does not preclude the possibility that Sakura may still influence Otu’s deubiquitinase activity, potentially guiding Otu to its substrate and determining substrate specificity. Our in vitro assays may not have detected such specificity changes.

Otu is an RNA-binding protein whose deubiquitinase activity is enhanced by RNA binding (Ji et al. 2019). Bam, along with other proteins such as Bgcn, Mei-P26, and Sxl, binds —a key stem cell maintenance factor (Wang and Lin 2004)— and represses its translation in GSCs (Wang and Lin 2004; Li et al. 2009; Chau et al. 2012; Li et al. 2013). Bam and Otu form a protein complex (Ji et al. 2017). Future studies should explore whether Sakura modulates Otu’s RNA-binding properties and its interaction with other proteins. Although Sakura is exclusively expressed in ovaries, particularly in germline cells including GSCs (Fig 1), Otu is broadly expressed in various tissues, including testes and gut (Steinhauer and Kalfayan 1992b). This restricted expression pattern suggests that Sakura may serve as a female germline-specific cofactor that enhances or modifies Otu’s molecular functions. Identifying Otu’s RNA targets and deubiquitinase substrates beyond CycA, as well as determining how Sakura binding influences these activities, will be key to elucidating their roles in oogenesis. One intriguing possibility is that Sakura regulates Otu’s activity toward piRNA pathway components, thereby contributing proper piRNA biogenesis and function. Additionally, the Sakura-Otu complex may directly regulate essential post-transcriptional processes such as sxl alternative splicing and translational control of other oogenic RNAs.

In summary, this study identifies and characterizes the previously unknown gene sakura, which is specifically expressed in female germ cells and is essential for oogenesis and female fertility. Together with its protein partner Otu, Sakura likely regulates germline cell fate, maintenance, and differentiation.

Materials and Methods

Fly strains

We generated the sakuranull strain by introducing indels within the Sakura coding region using the CRISPR/Cas9 genome editing system, as we previously reported (Zhu et al. 2018a; Zhu et al. 2018b; Zhu et al. 2019b; Zhu and Fukunaga 2021). The transgenic sakura-EGFP, otu-EGFP, and otu(ΔTudor)-EGFP strains were created following previously published methods (Fukunaga et al. 2012; Kandasamy and Fukunaga 2016; Zhu and Fukunaga 2021). A fragment containing Sakura cDNA flanked by ∼1 kbp upstream and ∼1 kbp downstream genomic sequences was cloned. Fragments containing Otu and Otu(ΔTudor) cDNAs, flanked by ∼2.5 kbp upstream and ∼1 kbp downstream genomic sequences, were also cloned. The EGFP gene was fused in-frame to the C-terminus of the Sakura and Otu coding sequences within these fragments, which were then inserted into a pattB plasmid vector. The sakura-EGFP plasmid was integrated into the 51C1 site within the fly genome using the BDSC:24482 fly strain and the PhiC31 system, while the otu-EGFP and otu(ΔTudor)-EGFP plasmids were integrated into the 25C6 site of the genome using the attP40 fly strain and the PhiC31 system.

The sakura RNAi #1 (VDRC: v39727) and #2 (VDRC: v103660), y-RNAi (v106068), TOsk-Gal4 (v 314033), Burdock sensor [UAS-Dcr2; NGT-Gal4; nosGal4-VP16, nos>NLS_GF’_lacZ_vas-3’UTR_burdock-target] (v 313217) strains were from Vienna Drosophila Resource Center. w-RNAi (BDSC: 35573), otu-RNAi (BDSC: 34065), bam-RNAi (BDSC: 33631), cycA-RNAi (BDSC: 29313), UAS-Dcr-2; NGT-Gal4 (BDSC: 25751), FRT82B/TM6C, Sb (BDSC: 86313), otu14 (BDSC: 6025), UASp-Mad-GFP (BDSC: 604592), and UASp-tkv.Q253D (BDSC: 604934) were obtained from the Bloomington Stock Center. The bam-GFP reporter (DGRC: 118177) and vasa-EGFP knocked-in fly (DGRC: 118616) were from Kyoto Drosophila Stock Center. sakura RNAi #2 was much healthier than sakura RNAi #1 without being driven with any Gal4 for some reasons and thus we chose to use the RNAi #2 line in most of this study.

sakura-EGFP does not cause any phenotypes in the background of sakura+/+ and sakura null/+.

Fertility assay

The female fertility assay was performed as previously described (Zhu et al. 2018b; Liao et al. 2019; Zhu et al. 2019a; Zhu and Fukunaga 2021). Briefly, five test virgin females of each test genotype were mated with three wild-type (OregonR) males in a cage with a 6-cm grape juice agar plate supplemented with wet yeast paste. The agar plates were replaced daily, and the number of eggs laid on the third plate (on day 3) was recorded. After incubation at 25°C for an additional day, the number of hatched eggs was counted. At least three cages per genotype were tested.

For the male fertility assay, a single test male was mated with five wild-type (OregonR) virgin females in each vial, following previously published methods (Zhu et al. 2018b; Zhu et al. 2019a; Zhu and Fukunaga 2021). After 3 days, the females were transferred to a new vial (vial 1). Every two days, they were transferred to a new vial until a total of four vials were obtained. After 2 days in the fourth vial, the females were removed, and the total number of progenies emerging from these four vials was counted. At least five males per genotype were tested.

Sakura and Otu antibodies

We expressed a recombinant full-length Sakura protein as an N-terminal 6xHis-tagged protein in E. coli using a modified pET vector and purified it using Ni-sepharose (GE Healthcare) and HiTrapQ HP (GE Healthcare) columns (Fukunaga and Doudna 2009). Recombinant Otu fragments (145-405aa and 406-853aa) were similarly expressed as N-terminally 6xHis-MBP-fusion proteins in E. coli and purified using Ni-sepharose. These purified proteins were used as antigens to generate polyclonal anti-Sakura and anti-Otu sera in rabbits (Pocono Rabbit Farm & Laboratory, Inc.). The rabbit polyclonal anti-Sakura antibodies were affinity purified using His-MBP-Sakura recombinant protein and Affigel-15 (Bio-rad), following the manufacturer’s instructions. Rabbit anti-Otu sera were first pre-cleared with purified His-MBP protein bound to Affigel-15, and the unbound fraction was affinity purified using purified His-MBP-Otu fragments (145-405aa and 406-853aa) and Affigel-15.

Immunostaining

Stereomicroscope images of dissected ovaries were taken using Leica M125 stereomicrocsope. Ovaries from 2- to 5-day-old, yeast-fed females were hand-dissected in 1X PBS (137 mM NaCl, 2.7 mM KCl, 10 mM Na2HPO4, 1.8 mM KH2PO4, pH 7.4) at room temperature. The dissected ovaries were fixed in a fixative buffer (4% formaldehyde, 15 mM PIPES (pH 7.0), 80 mM KCl, 20 mM NaCl2, 2 mM EDTA, and 0.5 mM EGTA), incubated for 30 minutes at room temperature with gentle rocking. After fixation, the ovaries were rinsed three times with PBX (0.1% Triton X-100 in 1X PBS) and then incubated in a blocking buffer (2% donkey serum in 3% BSA [w/v], 0.02% NAN3 [w/v] in PBX) for 1 hour at room temperature. Then, the ovaries were incubated with primary antibodies diluted in the blocking buffer overnight at 4°C. The following day, the ovaries were rinsed three times with PBX and incubated with Alexa Fluor-conjugated secondary antibodies for 2 hours. The ovaries were rinsed three times with PBX and then mounted in VECTASHIELD® PLUS antifade mounting medium with DAPI (H-2000, Vector lab). Confocal images were acquired on a Zeiss LSM700 confocal microscope at the Johns Hopkins University School of Medicine Microscope Facility.

The primary antibodies used for immunostaining were mouse anti-HTS (1B1) (DSHB, AB_528070, dilution: 1/100), mouse anti-Bam (DSHB, AB_10570327, 1/20), mouse anti-CycA (DSHB, AB_528188, 1/100), rat anti-Vasa (DSHB, AB_760351, 1/100), rat anti-DE Cadherin (DCAD2) (DSHB, AB_528120, 1/20), rabbit anti-pMad (Cell Signaling, Phospho-SMAD1/5 (Ser463/465) mAb #9516, 1/200,), and rabbit anti-cleaved caspase-3 (Cell Signaling, Cleaved Caspase-3 (Asp175) #9661, 1/200). Secondary antibodies used were Alexa Fluor 488 Donkey anti-Mouse Igg (ThermoFisher, A21202, 1/100), Alexa Fluor 594 Donkey anti-Rat Igg (ThermoFisher, A21209, 1/100), Alexa Fluor 594 Donkey anti-Mouse Igg (ThermoFisher, A21203, 1/100), and Alexa Fluor 594 Donkey anti-Rabbit Igg ThermoFisher, A21207, 1/100). Rhodamine phalloidin (ThermoFisher, R415, 1/100) was used to stain F-Actin.

While our Sakura antibody detects Sakura in immunostaining, it seems to detect some other proteins as well. Since we have Sakura-EGFP fly strain, which rescues sakuranull phenotypes, we relied on Sakura-EGFP rather than anti-Sakura antibodies immunostaining to examine Sakura expression and localization.

Germline clonal analysis

We generated sakuranull mutant clones using FLP/FRT-mediated recombination (Rubin and Huynh 2015). To induce sakuranull GSC clones, 3-day-old female flies of the genotype hs-flp/w; +; FRT82B, ubi-GFP/FRT82B, sakuranull were heat-shocked at 37°C for 1 hour, twice daily, with an 8-hour interval between heat shocks. Female flies of the genotype hs-flp/w; +; FRT82B, ubi-GFP/FRT82B were used as controls. Ovaries were dissected and stained 4, 7, and 14 days after clone induction. To induce primordial germ cell (PGC) clones, early third-instar larvae were subjected to the same heat shock regime, then allowed to develop to adulthood. Ovaries were dissected and analyzed at 3 to 4 days after eclosion.

To induce sakuranull mutant clones in the presence of otu-EGFP or otu(ΔTudor)-EGFP transgenes, 3-day-old female flies with the genotype hs-flp/w; otu-EGFP/+; FRT82B, ubi-RFP/FRT82B, sakuranull and hs-flp/w; otu(ΔTudor)-EGFP/+; FRT82B, ubi-RFP/FRT82B, sakuranull were heat-shocked under the same conditions. Control genotypes were hs-flp/w; otu-EGFP/+; FRT82B, ubi-RFP/FRT82B and hs-flp/w; otu(ΔTudor)-EGFP/+; FRT82B, ubi-RFP/FRT82B were used as controls. Flies were dissected 3-4 days after clone induction.

Western blot

Lysates of hand-dissected ovaries and tissues were prepared by homogenizing in RIPA buffer (50 mM Tris-HCl [pH 7.4], 150 mM NaCl, 1% [v/v] IGEPAL CA-630, 0.1% [w/v] sodium dodecyl sulfate (SDS), 0.5% [w/v] sodium deoxycholate, 1 mM ethylenediaminetetraacetic acid (EDTA), 5 mM dithiothreitol, and 0.5 mM phenylmethylsulfonyl fluoride (PMSF)) (Kandasamy et al. 2017; Zhu et al. 2018b). The homogenates were centrifuged at 21,000g at 4°C for 10 min, and the protein concentration of the supernatant was determined using the BCA protein assay kit (Pierce) as needed. Fifteen μg of total protein was loaded per lane for Western blot. The sources and dilutions of the primary antibodies were as below. Rabbit anti-Sakura (1/10000, generated in this study), rabbit anti-Otu (1/10000, generated in this study), rabbit anti-alpha-Tubulin [EP1332Y] (1/10000, Abcam, ab52866), mouse anti-alpha-Tubulin [12G10] (1/10000, DSHB, AB_1157911), mouse anti-FLAG (1/10000, Sigma, F1804), mouse anti-HA (1/10000, Sigma, H3663), and mouse anti-GFP [GF28R] (1/3000, Invitrogen, 14-6674-82). IRDye 800CW goat anti-mouse IgG, IRDye 800CW goat anti-rabbit IgG, IRDye 680RD goat anti-mouse, and IgG IRDye 680RD goat anti-rabbit were used as secondary antibodies. The membranes were scanned using the Li-Cor Odyssey CLx Imaging System.

Mass spectrometry

Immunoprecipitation of Sakura-EGFP protein was performed using the GFP-Trap Magnetic Agarose Kit (Proteintech, gtmak-20) on dissected ovaries from flies harboring sakura-EGFP transgene, with w1118 flies as controls. Ovaries were homogenized in 200 μL ice-cold lysis buffer (10 mM Tris-HCl [pH 7.5], 150 mM NaCl, 0.5 mM EDTA, 0.05% [v/v] IGEPAL CA-630) containing 1× protease inhibitor cocktail (100× protease inhibitor cocktail contains 120 mg/ml 1 mM 4-(2-aminoethyl) benzene sulfonyl fluoride hydrochloride (AEBSF), 1 mg/ml aprotinin, 7 mg/ml bestatin, 1.8 mg/ml E-64, and 2.4 mg/ml leupeptin). After homogenization, the tubes were placed on ice for 30 minutes, and the homogenates were extensively pipetted every 10 minutes. The lysates were then centrifuged at 17,000x g for 10 minutes at 4°C. The supernatants were transferred to pre-chilled tubes, and 300 μL dilution buffer (10 mM Tris/Cl pH 7.5, 150 mM NaCl, 0.5 mM EDTA) supplemented with 1x protease inhibitor cocktail were added. The diluted lysates were then added to the GFP-trap magnetic beads in 1.5 mL tubes and rotated for 1 hour at 4°C. After separating the beads with a magnetic tube rack, the beads were washed three times with 500 μL wash buffer (10 mM Tris/Cl pH 7.5, 150 mM NaCl, 0.05 % [v/v] IGEPAL CA-630). Proteins were eluted with 40 μL acidic elution buffer (200 mM glycine pH 2.5) followed by immediate neutralization with 5 uL neutralization buffer (1 M Tris pH 10.4).

As a quality control before mass spectrometry, ∼5 μL of the samples were mixed with an equal volume of 2× SDS-PAGE loading buffer (80 mM Tris-HCl [pH 6.8], 2% [w/v] SDS, 10% [v/v] glycerol, 0.0006% [w/v] bromophenol blue, 2% [v/v] 2-mercaptoethanol), heated at 95 °C for 3 min, and run on 4–20% Mini-PROTEAN® TGX™ Precast Protein Gels (Biorad, #4561094). Silver staining was then performed by using Pierce™ Silver Stain Kit (ThermoFisher, 24612) to assess the quality of the immunoprecipitated protein samples. Mass spectrometry was conducted at the Mass Spectrometry Core at the Department of Biological Chemistry, Johns Hopkins School of Medicine, as previously described (Zhu and Fukunaga 2021).

Co-immunoprecipitation

For co-immunoprecipitation of endogenous Sakura protein, wild-type ovaries (w1118) were homogenized in RIPA buffer, centrifuged at 21,000g at 4°C for 10 min, and the clear supernatant protein lysates were used for immunoprecipitation. Four μg of rabbit anti-Sakura and normal rabbit IgG (Cell Signaling, #2729) were incubated with 50 μL of Dynabeads Protein G (ThermoFisher, 10004D) for 20 min at room temperature. The beads were washed once with PBST (1x PBS with 0.1% Tween-20). The ovary lysate supernatant was then incubated with the washed beads at room temperature for 30 min, followed by three washes with PBST. The proteins were eluted with 2x SDS-PAGE loading buffer (80 mM Tris-HCl [pH 6.8], 2% [w/v] SDS, 10% [v/v] glycerol, 0.0006% [w/v] bromophenol blue, 2% [v/v] 2-mercaptoethanol) and heated at 70°C for 10 min. After bead separation using a magnetic tube rack, the eluted proteins in 2x SDS-PAGE loading buffer were heated again at 95°C for 3 min.

Transient protein expression in S2 cells was performed using the pAc5.1/V5-HisB plasmid vector (Invitrogen). A total of 1 μg plasmids were transfected using the Effectene transfection reagent (Qiagen, 301425). Three days after transfection, cells were harvested and lysed with ice-cold lysis buffer supplemented with 1× protease inhibitor cocktail. The cell lysates were then centrifuged at 17,000x g for 10 min at 4°C, and the clear supernatants were collected for immunoprecipitation. For anti-HA immunoprecipitation, supernatants were incubated with 25 μL (0.25 mg) of Pierce™ Anti-HA Magnetic Beads (ThermoFisher, 88837) at room temperature for 30 min. The beads were then washed three times with TBST (1x TBS [0.05 M Tris/HCl and 0.15 M NaCl, pH 7.6] with 0.05% Tween-20). For anti-FLAG immunoprecipitation, 2 μg of mouse anti-FLAG (1/10000, Sigma, F1804) was incubated with 50 μL of Dynabeads Protein G (ThermoFisher, 10004D) for 10 min at room temperature. The beads were washed once with PBST. The S2 cell lysate supernatant was incubated with the washed beads at room temperature for 15 min. The beads were washed three times with PBST. In both anti-HA and anti-FLAG immunoprecipitations, proteins were eluted with 2× SDS-PAGE loading buffer. For anti-HA, the beads in 2x SDS-PAGE loading buffer were heated at 95°C for 7 min. For anti-FLAG, the beads were heated at 70°C for 10 min, then were separated using a magnetic tube rack. The eluted proteins in 2x SDS-PAGE loading buffer were heated again at 95°C for 3 min.

While the anti-Otu antibody worked in immunoprecipitation from ovaries, in subsequent in anti-Sakura Western blot after anti-Otu IP, Otu antibody light chain bands often overlapped with the Sakura band. Using epitope-tags in S2 cells avoided this issue.

Small RNA and mRNA sequencing

Poly-A+ mRNA purification was performed as previously described (Fukunaga et al. 2012). Small RNA libraries and poly-A+ mRNA libraries were prepared, sequenced on Hiseq2500 (Illumina), and analyzed, as previously reported (Fukunaga et al. 2014; Kandasamy et al. 2017; Liao et al. 2018; Zhu et al. 2018a; Zhu et al. 2018b; Liao et al. 2019; Zhu et al. 2019a; Zhu et al. 2019b). SRA accession number for these datasets is PRJNA1156618.

RT-PCR

Total RNAs from testes and ovaries were prepared using miRVana (Thermo Fisher Scientific). RNAs were treated with Turbo DNase (Thermo Fisher Scientific) to remove potential genomic DNA contamination. A total of 1 μg of RNA was reverse-transcribed into cDNA using SuperScriptTM VILOTM MasterMix (Thermo Fisher Scientific). To assess sxl alternative splicing, PCR was performed using GoTaq Green Master Mix (Promega) with the primers sxl-F (CTCACCTTCGATCGAGGGTGTA) and sxl-R (GATGGCAGAGAATGGGAC).

In vitro deubiquitination assay

C-terminally 6xHis-tagged recombinant Sakura protein was expressed in E. coli using a modified pET vector and was purified using Ni-sepharose. N-terminally 3xHA-HRV3Csite-3xFLAG-tagged Otu protein and a negative control, N-terminally 3xHA-HRV3Csite-3xFLAG-tagged firefly luciferase protein, were expressed in S2 cells using pAc5.1/V5-HisB vector. The proteins were immunoprecipitated using Pierce™ Anti-HA Magnetic Beads (ThermoFisher, 88837) as described previously, with a more stringent washing step to effectively remove interacting proteins. The beads were washed six times with a high salt wash buffer (TBST with 800 mM NaCl). The proteins were eluted by cleaving off 3xHA tag with 25 nM GST-HRV3C protease in a cleavage buffer (25mM Tris-HCl pH 7.4, 150 mM NaCl, 5% glycerol, 2mM EDTA), incubated at 4°C with rotation for 6 hours. An equal volume of 100% glycerol was added to the eluted protein.

Purified proteins, including Otu, luciferase, and Sakura, were mixed with Ub-Rhodamine 110 (Ubiquitin-Proteasome Biotechnologies, M3020) in a total volume of 30 μL in reaction buffer (20 mM Tris-HCl, pH 7.5, 200 mM NaCl, 5 mM MgCl2, 2 mM DTT). The mixture was added to a black 384-well low-volume plate. Fluorescence measurements were taken using SpectraMax i3x Multi-Mode Microplate Reader at 37°C, with excitation and emission wavelengths set at 485/20 and 530/20 nm respectively. The fluorescence intensity for each condition was averaged from triplicates and plotted as a function of time.